---
title: "STARR_linear_regression_March2026"
author: "FL"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE
)

require(data.table)
require(ggplot2)
require(broom)
require(purrr)
require(metafor)
library(dplyr)
library(tidyr)
require(readxl)
require(ggrepel)

setwd("/groups/stark/lorbeer/Burst_params_DG/STARR_March2025_pools/")

```


## Load pre-processed data
1) Table with all derived burst paramters of the STARR-SEQ (ENHANCER) experiments
2) sample-specific expression thresholds from mu_thresholds_v2_Reyna.xlsx 

```{r}
#load data 
# Import motif counts 
param <- readRDS("/groups/stark/lorbeer/Burst_params_DG/STARR_March2025_pools/STARR_annotated_log2FC.rds")
sample_names <- unique(as.factor(param$sample))
sample_names <- sample_names[c(1:3, 7:9, 13:15)]

pooled_samples <- c("libFL010_E24_rg_pooledby2", "libFL013_E39_p309_pooledby4", "libFL014_E39_p319_pooledby6")

mu_thresholds <- read_xlsx("/groups/stark/lorbeer/Burst_params_DG/mu_thresholds_v2_Reyna.xlsx")
param <- merge(param, mu_thresholds, by = c("sample", "experiment", "library"), all.x = TRUE)
param[, high_mu := exp >= mu_threshold_nofilter]

head(param)

param_list <- list()
#read into list object and set filtering criteria 
for (s in sample_names) {
  #param_list[[s]] <- as.data.frame(param[sample == s & (full_prec2 ==TRUE & high_mu == TRUE)])
  #param_list[[s]] <- as.data.frame(param[sample == s & (bfreq.thres == TRUE & bsize.thres == TRUE & KS_pval > 0.01)])
  param_list[[s]] <- as.data.frame(param[sample == s & (high_mu & KS_pval > 0.01)])
}
```
## helper functions
Significance levels, clipping motif occurrence levels 
--> the model is not looking at effect presence/absence but at effect per unit increase in clipped motif count

- motif count predictors are upper-clipped before modeling, the clipping threshold is based on the 95th percentile within each motif column and constrained to lie between 2 and 3
values above the threshold are replaced by the threshold (reduce leverage from rare high-count observations while retaining informaiton)

```{r}
#generate clipped motif counts to account for there being very few observations on the high end of the data 
clip_motifs <- function(dat, cols,
                        q = 0.95,
                        min_cutoff = 2,
                        max_cutoff = 3,
                        suffix = "_clip") {
  dat <- as.data.table(dat)  # safe if it's already a data.table
  
  out_cols <- paste0(cols, suffix)
  
  dat[, (out_cols) := lapply(.SD, function(x) {
    cutoff <- quantile(x, q, na.rm = TRUE)
    
    cutoff <- fcase(
      cutoff < min_cutoff, min_cutoff,
      cutoff > max_cutoff, max_cutoff,
      default = cutoff
    )
    
    fifelse(x > cutoff, cutoff, x)
  }), .SDcols = cols]
  
  dat
}

cols <- c("ETS","GATA","AP1","SREBP","Trl","twist","Ohler6","Ohler7","Ohler1","Dref")

param_list <- lapply(param_list, clip_motifs, cols = cols)

# if you also want the updated column-name vector:
cols_clip <- paste0(cols, "_clip")

sig_stars <- function(p) {
  case_when(
    is.na(p)  ~ NA_character_,
    p < 0.001 ~ "***",
    p < 0.01  ~ "**",
    p < 0.05  ~ "*",
    TRUE      ~ "ns"
  )
}

na_block_for_motifs <- function(motifs, outcome, model_label, n_used = NA_integer_) {
  tibble(
    Motif = motifs,
    Outcome = outcome,
    Model = model_label,
    n_used = n_used,
    n_present = NA_integer_,
    prop_present = NA_real_,
    estimate = NA_real_,
    std.error = NA_real_,
    statistic = NA_real_,
    p.value = NA_real_,
    FoldChange = NA_real_
  )
}

```

## linear model 
(different to function used in STAP-seq is motif counting >0 rather than counting instances of 1 occurrence)
in the enhancer analysis, clipped motif counts are used as numeric predictors, regression coefficients represent the expected change in log10(outcome) per one-unit increase in clipped motif count

```{r}
fit_motif_model_one_df <- function(data, outcome, motifs, include_exp = FALSE) {
  model_label <- ifelse(include_exp, "motifs_plus_log10exp", "motifs_only")
  
  # Basic availability checks
  motifs_present <- intersect(motifs, colnames(data))
  motifs_missing <- setdiff(motifs, motifs_present)
  
  if (!outcome %in% colnames(data)) {
    return(na_block_for_motifs(motifs, outcome, model_label, n_used = NA_integer_))
  }
  if (include_exp && !"exp" %in% colnames(data)) {
    return(na_block_for_motifs(motifs, outcome, model_label, n_used = NA_integer_))
  }
  
  # Fit one model per motif that exists; for missing motifs, return NA rows
  fitted_present <- purrr::map_dfr(motifs_present, function(m) {
    rhs <- m
    if (include_exp) rhs <- paste(rhs, "+ log10(exp)")
    fml <- as.formula(paste0("log10(", outcome, ") ~ ", rhs))
    
    mf <- tryCatch(model.frame(fml, data = data), error = function(e) NULL)
    n_used <- if (is.null(mf)) NA_integer_ else nrow(mf)
    
    if (is.null(mf) || n_used == 0) {
      # No usable rows for this motif's model
      return(
        na_block_for_motifs(m, outcome, model_label, n_used = ifelse(is.null(mf), NA_integer_, 0L)) %>%
          dplyr::mutate(
            n_present = ifelse(is.null(mf), NA_integer_, sum(mf[[m]] > 0, na.rm = TRUE)),
            prop_present = ifelse(is.null(mf), NA_real_, n_present / n_used)
          )
      )
    }
    
    n_present <- sum(mf[[m]] == 1, na.rm = TRUE)
    prop_present <- n_present / n_used
    
    model <- tryCatch(lm(fml, data = data), error = function(e) NULL)
    if (is.null(model)) {
      return(
        na_block_for_motifs(m, outcome, model_label, n_used = n_used) %>%
          dplyr::mutate(n_present = n_present, prop_present = prop_present)
      )
    }
    
    broom::tidy(model) %>%
      dplyr::filter(term == m) %>%
      dplyr::transmute(
        Motif = m,
        Outcome = outcome,
        Model = model_label,
        n_used = n_used,
        n_present = n_present,
        prop_present = prop_present,
        estimate = estimate,
        std.error = std.error,
        statistic = statistic,
        p.value = p.value,
        FoldChange = 10^estimate
      )
  })
  
  missing_part <- na_block_for_motifs(motifs_missing, outcome, model_label, n_used = NA_integer_) %>%
    dplyr::mutate(n_present = NA_integer_, prop_present = NA_real_)
  
  dplyr::bind_rows(fitted_present, missing_part) %>%
    dplyr::mutate(Motif = factor(Motif, levels = motifs)) %>%
    dplyr::arrange(Motif) %>%
    dplyr::mutate(Motif = as.character(Motif))
}

fit_motif_model_multiv_df <- function(data, outcome, motifs, include_exp = FALSE) {
  model_label <- ifelse(include_exp, "motifs_plus_log10exp", "motifs_only")
  
  motifs_present <- intersect(motifs, colnames(data))
  motifs_missing <- setdiff(motifs, motifs_present)
  
  if (!outcome %in% colnames(data)) {
    return(na_block_for_motifs(motifs, outcome, model_label, n_used = NA_integer_))
  }
  if (include_exp && !"exp" %in% colnames(data)) {
    return(na_block_for_motifs(motifs, outcome, model_label, n_used = NA_integer_))
  }
  if (length(motifs_present) == 0) {
    return(na_block_for_motifs(motifs, outcome, model_label, n_used = NA_integer_))
  }
  
  rhs <- paste(motifs_present, collapse = " + ")
  if (include_exp) rhs <- paste(rhs, "+ log10(exp)")
  fml <- as.formula(paste0("log10(", outcome, ") ~ ", rhs))
  
  mf <- tryCatch(model.frame(fml, data = data), error = function(e) NULL)
  n_used <- if (is.null(mf)) NA_integer_ else nrow(mf)
  
  if (is.null(mf) || n_used == 0) {
    # can still compute presence from mf only if mf exists
    out <- na_block_for_motifs(motifs_present, outcome, model_label, n_used = ifelse(is.null(mf), NA_integer_, 0L))
    return(dplyr::bind_rows(
      out,
      na_block_for_motifs(motifs_missing, outcome, model_label, n_used = NA_integer_)
    ))
  }
  
  # presence stats are still per motif, but computed on the multivariable model frame
  presence <- tibble::tibble(
    Motif = motifs_present,
    n_present = purrr::map_int(motifs_present, ~ sum(mf[[.x]] == 1, na.rm = TRUE)),
    prop_present = n_present / n_used
  )
  
  model <- tryCatch(lm(fml, data = data), error = function(e) NULL)
  if (is.null(model)) {
    out <- na_block_for_motifs(motifs_present, outcome, model_label, n_used = n_used)
    out <- dplyr::left_join(out, presence, by = "Motif")
    return(dplyr::bind_rows(
      out,
      na_block_for_motifs(motifs_missing, outcome, model_label, n_used = NA_integer_)
    ))
  }
  
  coefs <- broom::tidy(model) %>%
    dplyr::filter(term %in% motifs_present) %>%
    dplyr::transmute(
      Motif = term,
      Outcome = outcome,
      Model = model_label,
      n_used = n_used,
      estimate = estimate,
      std.error = std.error,
      statistic = statistic,
      p.value = p.value,
      FoldChange = 10^estimate
    ) %>%
    dplyr::left_join(presence, by = "Motif")
  
  missing_part <- na_block_for_motifs(motifs_missing, outcome, model_label, n_used = NA_integer_) %>%
    dplyr::mutate(n_present = NA_integer_, prop_present = NA_real_)
  
  dplyr::bind_rows(coefs, missing_part) %>%
    dplyr::mutate(Motif = factor(Motif, levels = motifs)) %>%
    dplyr::arrange(Motif) %>%
    dplyr::mutate(Motif = as.character(Motif))
}



```

## Usage of the lm with clipped motifs 
Generates a table with the results of lm (univariate, multivariate and also with/without expression included)
meaning of coefficients: association with motif count (clipped at upper bound)
regression asks about dose/count effect

```{r}

# -------------------
# Usage
# -------------------

# pre-specified subset of motifs 

# Define motif vector

motifs_10 <- cols_clip

# Run
# results <- fit_motif_model_one_df(param_list, outcome = "bsize", motifs = motifs_10, include_exp = FALSE)
# print(results)

fit_motif_models_list_both <- function(data_list, outcomes, motifs, include_exp = FALSE) {
  purrr::imap_dfr(data_list, function(df, nm) {
    purrr::map_dfr(outcomes, function(outcome) {
      
      multiv <- fit_motif_model_multiv_df(
        data = df, outcome = outcome, motifs = motifs, include_exp = include_exp
      ) %>% dplyr::mutate(Analysis = "multivariate")
      
      univ <- fit_motif_model_one_df(
        data = df, outcome = outcome, motifs = motifs, include_exp = include_exp
      ) %>% dplyr::mutate(Analysis = "univariate")
      
      dplyr::bind_rows(multiv, univ) %>%
        dplyr::mutate(
          sample_name = nm,
          sig = sig_stars(p.value)
        )
    })
  }) %>%
    dplyr::select(sample_name, Outcome, Analysis, Model, Motif, sig, dplyr::everything())
}

#

outcomes <- c("bsize", "bfreq", "exp")

results_both <- dplyr::bind_rows(
  fit_motif_models_list_both(param_list, outcomes, motifs_10, include_exp = FALSE),
  fit_motif_models_list_both(param_list, outcomes, motifs_10, include_exp = TRUE)
)

print(results_both)



```

## functions for plotting multivariate analysis 
Generates a heatmap 

```{r}
sym_limits <- function(x, midpoint = 0) {
  x <- x[is.finite(x)]
  if (length(x) == 0) return(c(-1, 1))
  max_abs <- max(abs(x - midpoint))
  c(midpoint - max_abs, midpoint + max_abs)
}

make_heatmap_model <- function(results, model_name,
                               motifs_order = NULL,
                               samples_order = NULL,
                               min_occurrence = 0,
                               star_size = 3,
                               analysis = NULL) {
  
  plot_df <- results %>%
    dplyr::filter(Model == model_name, Outcome %in% c("bsize", "bfreq", "exp"))
  
  analysis_label <- if (is.null(analysis)) "both (univariate + multivariable)" else analysis
  
  if (!is.null(analysis)) {
    plot_df <- plot_df %>% dplyr::filter(Analysis == analysis)
  }
  
  # Subset + order samples
  if (!is.null(samples_order)) {
    plot_df <- plot_df %>%
      dplyr::filter(sample_name %in% samples_order) %>%
      dplyr::mutate(sample_name = factor(sample_name, levels = samples_order))
  } else {
    plot_df <- plot_df %>% dplyr::mutate(sample_name = factor(sample_name))
  }
  
  plot_df <- plot_df %>%
    dplyr::mutate(
      value = estimate,
      value = ifelse(!is.na(n_present) & n_present < min_occurrence, NA_real_, value),
      Sig_raw_plot = ifelse(sig %in% c("*","**","***"), sig, "")
    )
  
  # Motif ordering
  if (!is.null(motifs_order)) {
    plot_df <- plot_df %>% dplyr::mutate(Motif = factor(Motif, levels = motifs_order))
  } else {
    plot_df <- plot_df %>% dplyr::mutate(Motif = factor(Motif))
  }
  
  lims <- sym_limits(plot_df$value, midpoint = 0)
  
  ggplot2::ggplot(plot_df, ggplot2::aes(x = sample_name, y = Motif)) +
    ggplot2::geom_tile(ggplot2::aes(fill = value)) +
    ggplot2::geom_text(ggplot2::aes(label = Sig_raw_plot), size = star_size, na.rm = TRUE) +
    ggplot2::facet_grid(. ~ Outcome) +
    ggplot2::scale_fill_gradient2(
      low = "blue", mid = "white", high = "red",
      midpoint = 0,
      limits = lims,
      na.value = "grey80"
    ) +
    ggplot2::labs(
      title = paste0("Motif effects (", analysis_label, "): ", model_name),
      subtitle = paste0(
        "Stars = raw p-value; tiles shown only if n_present ≥ ", min_occurrence,
        if (!is.null(samples_order)) paste0("; samples ordered (n=", length(samples_order), ")") else ""
      ),
      x = "Sample",
      y = "Motif",
      fill = "estimate"
    ) +
    ggplot2::theme_minimal(base_size = 11) +
    ggplot2::theme(
      panel.grid = ggplot2::element_blank(),
      axis.text.x = ggplot2::element_text(angle = 45, hjust = 1, vjust = 1),
      strip.text = ggplot2::element_text(face = "bold")
    )
}



```

## heatmaps following multivariate analysis 
Settings to be set: minimal motif occurrence in groups,
motif order to display, 
which samples to plot, 
multivariate, univariate etc 

```{r}
# Define outside the function
min_occurrence <- 5

motifs_order <- motifs_10
pooled_samples
all_samples <- sample_names

samples_order <- pooled_samples


q_multiv <- make_heatmap_model(results_both, "motifs_only",
                               motifs_order = motifs_order,
                               samples_order = samples_order,
                               min_occurrence = min_occurrence,
                               analysis = "multivariate")

q_univ <- make_heatmap_model(results_both, "motifs_only",
                             motifs_order = motifs_order,
                             samples_order = samples_order,
                             min_occurrence = min_occurrence,
                             analysis = "univariate")

q_multiv
q_univ

#all samples for supplement
p_multiv <- make_heatmap_model(results_both, "motifs_only",
                               motifs_order = motifs_order,
                               samples_order = all_samples,
                               min_occurrence = min_occurrence,
                               analysis = "multivariate")

p_univ <- make_heatmap_model(results_both, "motifs_only",
                             motifs_order = motifs_order,
                             samples_order = all_samples,
                             min_occurrence = min_occurrence,
                             analysis = "univariate")

p_multiv
p_univ


```
## Median log2FC analysis 
median/Wilcoxon analysis: binary presence/absence

```{r}

#median analysis: 
#median difference 
analyze_param_list_log2FC <- function(param_list,
                                      motifs_order,
                                      samples_order = NULL) {
  stopifnot(is.list(param_list))
  
  # ensure names exist (used as sample_name)
  if (is.null(names(param_list)) || any(names(param_list) == "")) {
    names(param_list) <- paste0("sample_", seq_along(param_list))
  }
  
  # optionally restrict + order samples (e.g., pooled_samples)
  if (!is.null(samples_order)) {
    keep <- intersect(samples_order, names(param_list))
    param_list <- param_list[keep]
    # keep the requested order (including if some were missing)
    param_list <- param_list[samples_order[samples_order %in% names(param_list)]]
  }
  
  add_significance <- function(p) {
    dplyr::case_when(
      is.na(p)  ~ "",
      p < 0.001 ~ "***",
      p < 0.01  ~ "**",
      p < 0.05  ~ "*",
      TRUE      ~ ""
    )
  }
  
  sym_limits <- function(x, midpoint = 0) {
    x <- x[is.finite(x)]
    if (length(x) == 0) return(c(-1, 1))
    max_abs <- max(abs(x - midpoint), na.rm = TRUE)
    c(midpoint - max_abs, midpoint + max_abs)
  }
  
  results_df <- purrr::imap_dfr(param_list, function(df, sample_name) {
    df <- data.table::as.data.table(df)
    
    purrr::map_dfr(motifs_order, function(m) {
      if (!m %in% names(df)) {
        return(tibble::tibble(
          sample_name = sample_name,
          motif = m,
          measurement = c("bfreq", "bsize", "exp"),
          value_0 = NA_real_, value_1 = NA_real_, diff = NA_real_, log2FC = NA_real_,
          pvalue = NA_real_, Nobservations_0 = NA_integer_, Nobservations_1 = NA_integer_
        ))
      }
      
      dt_m <- df[, .(bfreq, bsize, exp, motif = get(m))]
      
      n0 <- sum(dt_m$motif == 0, na.rm = TRUE)
      n1 <- sum(dt_m$motif > 0, na.rm = TRUE)
      
      one_measure <- function(y) {
        med0 <- dt_m[motif == 0, median(get(y), na.rm = TRUE)]
        med1 <- dt_m[motif > 0, median(get(y), na.rm = TRUE)]
        wt <- suppressWarnings(stats::wilcox.test(get(y) ~ (motif > 0), data = dt_m))
        
        tibble::tibble(
          sample_name = sample_name,
          motif = m,
          measurement = y,
          value_0 = med0,
          value_1 = med1,
          diff = med1 - med0,
          log2FC = log2(med1 / med0),
          pvalue = wt$p.value,
          Nobservations_0 = n0,
          Nobservations_1 = n1
        )
      }
      
      dplyr::bind_rows(one_measure("bfreq"), one_measure("bsize"), one_measure("exp"))
    })
  }) %>%
    dplyr::mutate(
      significance = add_significance(pvalue),
      motif = factor(motif, levels = motifs_order),
      sample_name = factor(sample_name, levels = names(param_list)),
      measurement = dplyr::recode(
        measurement,
        bfreq = "Burst Frequency",
        bsize = "Burst Size",
        exp   = "Mean expression"
      )
    )
  
  lims <- sym_limits(results_df$log2FC, midpoint = 0)
  
  p <- ggplot2::ggplot(results_df, ggplot2::aes(x = sample_name, y = motif, fill = log2FC)) +
    ggplot2::geom_tile() +
    ggplot2::geom_text(ggplot2::aes(label = significance), size = 3) +
    ggplot2::facet_grid(. ~ measurement) +
    ggplot2::scale_fill_gradient2(
      limits = lims,
      low = "blue", mid = "white", high = "red",
      midpoint = 0,
      oob = scales::squish,
      na.value = "grey90",
      name = "log2FC"
    ) +
    ggplot2::theme_minimal() +
    ggplot2::theme(
      panel.grid = ggplot2::element_blank(),
      axis.text.x = ggplot2::element_text(angle = 45, hjust = 1),
      axis.title.x = ggplot2::element_blank(),
      axis.title.y = ggplot2::element_blank(),
      strip.text = ggplot2::element_text(face = "bold")
    )
  
  list(table = results_df, plot = p)
}


```

## Heatmaps of the log2FC median analysis (motif presence/absence)

```{r}

#pooled samples only
res_pooled <- analyze_param_list_log2FC(
  param_list,
  motifs_order = motifs_10,
  samples_order = pooled_samples
)
print(res_pooled$plot)
pooled_table <- res_pooled$table

#all samples
res_all <- analyze_param_list_log2FC(param_list, motifs_order = motifs_10)
print(res_all$plot)
all_table <- res_all$table

```
## meta analysis 
settings: occurrence in 3 samples, fixed-effect, using multivariate - motifs_only results 

```{r}
# how many datasets/samples must contribute to pool a Motif×Outcome?
min_k <- 3  # set to whatever you want (e.g., 2, 3, ...)

# keep only usable rows for meta-analysis
meta_input <- results_both %>% filter(sample_name %in% pooled_samples[1:3] & Analysis == "multivariate" & Model == "motifs_only") %>% 
  filter(
    is.finite(estimate),
    is.finite(std.error),
    std.error > 0,
    !is.na(sample_name)
  )

# count how many samples contribute per Motif×Outcome
combos <- meta_input %>%
  distinct(sample_name, Motif, Outcome) %>%
  count(Motif, Outcome, name = "k") %>%
  filter(k >= min_k)

pooled_list <- vector("list", nrow(combos))

for (i in seq_len(nrow(combos))) {
  subset_data <- meta_input %>%
    filter(Motif == combos$Motif[i], Outcome == combos$Outcome[i]) %>%
    distinct(sample_name, .keep_all = TRUE)  # safety if duplicates exist
  
  # fixed-effects meta-analysis on log10-scale effect 
  #other option would be RMLE
  meta_obj <- rma(
    yi = subset_data$estimate,
    sei = subset_data$std.error,
    method = "FE"
  )
  
  pooled_list[[i]] <- data.frame(
    Motif = combos$Motif[i],
    Outcome = combos$Outcome[i],
    k = nrow(subset_data),
    
    pooled_estimate = as.numeric(meta_obj$beta[1]),  # log10 effect
    pooled_SE = as.numeric(meta_obj$se),
    
    pooled_pValue = as.numeric(meta_obj$pval),
    pooled_zValue = as.numeric(meta_obj$zval),
    
    # convert pooled estimate and CI back to fold-change on original scale
    pooled_FoldChange = 10^as.numeric(meta_obj$beta[1]),
    pooled_FC_low  = 10^as.numeric(meta_obj$ci.lb),
    pooled_FC_high = 10^as.numeric(meta_obj$ci.ub),
    
    # optional heterogeneity summaries (often useful)
    tau2 = as.numeric(meta_obj$tau2),
    I2 = as.numeric(meta_obj$I2)
  )
}

pooled_results <- bind_rows(pooled_list) %>%
  mutate(
    Sig_raw = case_when(
      pooled_pValue < 0.001 ~ "***",
      pooled_pValue < 0.01  ~ "**",
      pooled_pValue < 0.05  ~ "*",
      TRUE ~ "ns"
    )
  )

print(pooled_results)

#pivot wide for plotting
df_wide <- pooled_results %>%
  filter(Outcome %in% c("bsize", "bfreq")) %>%
  select(Motif, Outcome,
         pooled_estimate, pooled_SE, pooled_pValue,
         pooled_FoldChange, pooled_FC_low, pooled_FC_high,
         k, tau2, I2) %>%
  pivot_wider(
    names_from = Outcome,
    values_from = c(pooled_estimate, pooled_SE, pooled_pValue,
                    pooled_FoldChange, pooled_FC_low, pooled_FC_high,
                    k, tau2, I2),
    names_sep = "_"
  ) %>%
  # keep motifs where you have both pooled outcomes
  filter(!is.na(pooled_FoldChange_bsize), !is.na(pooled_FoldChange_bfreq)) %>%
  mutate(
    Significance = ifelse(
      pooled_pValue_bsize < 0.05 | pooled_pValue_bfreq < 0.05,
      "Significant", "Not significant"
    )
  )

df_wide <- df_wide %>%
  mutate(
    est_low_bsize  = pooled_estimate_bsize - 1.96 * pooled_SE_bsize,
    est_high_bsize = pooled_estimate_bsize + 1.96 * pooled_SE_bsize,
    est_low_bfreq  = pooled_estimate_bfreq - 1.96 * pooled_SE_bfreq,
    est_high_bfreq = pooled_estimate_bfreq + 1.96 * pooled_SE_bfreq
  )


df_wide <- df_wide %>%
  mutate(
    Significance = case_when(
      pooled_pValue_bsize < 0.05 & pooled_pValue_bfreq < 0.05 ~ "Both",
      pooled_pValue_bsize < 0.05 & pooled_pValue_bfreq >= 0.05 ~ "bsize only",
      pooled_pValue_bsize >= 0.05 & pooled_pValue_bfreq < 0.05 ~ "bfreq only",
      TRUE ~ "Neither"
    )
  )

print(df_wide)

```
## meta analysis plot 
Meta-analysis: Association of motif count with burst parameters

```{r}

#plot estimates


p <- ggplot(df_wide,
            aes(x = pooled_estimate_bsize,
                y = pooled_estimate_bfreq,
                color = Significance,
                label = Motif)) +
  geom_errorbarh(aes(xmin = est_low_bsize, xmax = est_high_bsize),
                 height = 0, alpha = 0.4) +
  geom_errorbar(aes(ymin = est_low_bfreq, ymax = est_high_bfreq),
                width = 0, alpha = 0.4) +
  geom_point(size = 3) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray40") +
  geom_hline(yintercept = 0, linetype = "dotted") +
  geom_vline(xintercept = 0, linetype = "dotted") +
  theme_classic(base_size = 14) +
scale_color_manual(values = c(
    "Both" = "deepskyblue",
    "bsize only" = "cyan3",
    "bfreq only" = "royalblue3",
    "Neither" = "gray60"
  )) +
  labs(
    x = expression("Pooled effect on log"[10]*"(bsize)  (beta)"),
    y = expression("Pooled effect on log"[10]*"(bfreq)  (beta)"),
    title = "Meta-analysis: Association of motif count with burst parameters",
    subtitle = paste0("Points = pooled betas; error bars = 95% CI; min_k = ", min_k)
  )

# labels: ggrepel if available; otherwise plain text
if (requireNamespace("ggrepel", quietly = TRUE)) {
  p <- p + ggrepel::geom_text_repel(size = 3, max.overlaps = 30)
} else {
  p <- p + geom_text(size = 3, vjust = -0.5, check_overlap = TRUE)
}
print(p)

# pdf("meta_pooled_analysis_estimates_STARR.pdf", width = 6, height = 5)
# print(p)
# dev.off()



```

