---
title: "promoter-BARe-seq: Figure 1 and Supplemental Figure 2"
author: "FL"
date: "2025-02-13"
output: html_document
---

<!--
================================================================================
OVERVIEW  — which code section makes which figure panel
================================================================================
This document reproduces the data-driven panels of:

  Figure 1
    A  Schematic of the BARe-seq method .......... (illustration, no code)
    B  log2FC BARe-seq vs. STAP-seq promoter activity (scatter)
                                                     -> "FIGURE 1B" chunk below
    C  Observed vs. NB-fit transcript/plasmid histograms for examples 1-3
                                                     -> made in the separate
                                                        counts/NB-fit markdown
                                                        (needs raw UMI counts)
    D  Burst size / frequency / mean-expression histograms
                                                     -> "FIGURE 1D" chunk below
    E  Burst size vs. burst frequency (with log2FC contour)
                                                     -> separate analysis

  Supplemental Figure 2
    A  Two-state model schematic ................. (illustration, no code)
    B  Replicate correlation, UNFILTERED
       (burst size / freq / mean exp) ........... -> "SUPPL. FIG. 2B/2D" chunks
    C  Mean-expression (mu) histograms per sample
       with mean-expression threshold line ...... -> "SUPPL. FIG. 2C" chunk
    D  Replicate correlation, FILTERED
       (KS-pval > 0.01 & mu threshold) .......... -> "SUPPL. FIG. 2B/2D" chunks
    E  log2FC BARe-seq vs. STAP-seq coloured by
       quality filters .......................... -> "SUPPL. FIG. 2E" chunk

Sections marked "QC / exploratory" are diagnostic plots that are not main
figure panels but are kept because they were used to choose thresholds.

All inputs are the published supplemental tables (see the data-loading chunk).
Sample names are the final publication names throughout.
================================================================================
-->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(data.table)
require(dplyr)
require(ggplot2)
require(readxl)
library(ggrepel)
require(patchwork)
require(gridExtra)
require(stringr)
```

## Data loading

R version 4.1.2 and Bioconductor 3.14

```{r, echo = FALSE}
# ------------------------------------------------------------------------------
# DATA LOADING
#
# Everything starts from the published supplemental tables (expected in ./Tables):
#
#   Table3_mu_thresholds.xlsx              per-sample mean-expression (mu)
#                                          thresholds. Only the publication
#                                          sample names (column `new_sample_name`)
#                                          and the threshold are used here.
#
#   Table4_promoter_library_annotations.xlsx  (sheet "STAP")
#                                          promoter annotations, core-promoter
#                                          motifs, gene names, CP_group
#                                          classification, and the STAP-seq
#                                          reference activity used for Fig. 1B
#                                          (column `log2FoldChange_STAPseq_zfh1`).
#
#   Table5_promoter_burst_parameters.xlsx  (sheet "STAP")
#                                          burst parameters (bfreq, bsize, exp,
#                                          log2FoldChange) plus the pre-computed
#                                          quality flags high_mu and KS_pval,
#                                          for every sample.
#
# Raw UMI counts are NOT part of these tables; the count-based panels
# (Fig. 1C) live in a separate markdown and read the counts from the GEO
# data repository accompanying the publication.
# ------------------------------------------------------------------------------

anno          <- read_xlsx("/groups/stark/lorbeer/BARE-seq_data_prep/Tables/Table4_promoter_library_annotations_updated.xlsx", sheet = "STAP")
mu_thresholds <- read_xlsx("/groups/stark/lorbeer/BARE-seq_data_prep/Tables/Table3_mu_thresholds.xlsx")
burst_params  <- read_xlsx("/groups/stark/lorbeer/BARE-seq_data_prep/Tables/Table5_promoter_burst_parameters.xlsx", sheet = "STAP")

# --- STAP_data_annotated: burst parameters + promoter annotations + mu threshold
STAP_data_annotated <- as.data.table(burst_params)

# Split the publication sample name into its components:
#   lib_experiment_condition_rep  e.g. libFL001 / STAP / zfh1 / pool
STAP_data_annotated[, c("lib", "experiment", "condition", "rep") :=
                      tstrsplit(sample, "_", keep = c(1:4))]

# Add promoter annotations (motifs, gene names, classification, STAP-seq ref.)
STAP_data_annotated <- merge(STAP_data_annotated, as.data.table(anno),
                             by = "seq_ID", all.x = TRUE)

# Per-sample mu threshold (used only for the threshold line in Suppl. Fig. 2C).
# Table3 stores log2 thresholds; convert back to linear mu. Use the publication
# sample names only.
mu_thresholds_dt <- as.data.table(mu_thresholds)[
  , .(sample = sample, mu_threshold_nofilter = 2^mu_threshold_log2)]
STAP_data_annotated <- merge(STAP_data_annotated, mu_thresholds_dt,
                             by = "sample", all.x = TRUE)

# high_mu is stored as text in the table; make it a proper logical.
STAP_data_annotated[, high_mu := (high_mu == "TRUE")]

# --- diff_exp_sub2: wide table for the log2FC scatter plots -------------------
# One column of BARe-seq activity per condition (from the pooled samples of
# Table5) plus the STAP-seq reference activity (from Table4).
burst_params_dt <- as.data.table(burst_params)
pool_dt <- burst_params_dt[grepl("_pool$", sample)]
pool_dt[, condition := tstrsplit(sample, "_", keep = 3L)[[1]]]

diff_exp_wide <- dcast(pool_dt, seq_ID ~ condition, value.var = "log2FoldChange")
setnames(diff_exp_wide,
  old = c("zfh1",                "CG4822",                "Pld"),
  new = c("log2FoldChange_zfh1", "log2FoldChange_CG4822", "log2FoldChange_Pld"))

anno_dt <- as.data.table(anno)
diff_exp_sub2 <- merge(
  diff_exp_wide,
  anno_dt[, .(seq_ID,
              log2FoldChange_STAPseq_zfh1,   # STAP-seq reference (x-axis, Fig. 1B)
              classification = CP_group)],
  by = "seq_ID")
setnames(diff_exp_sub2, "seq_ID", "oligo_id")

head(STAP_data_annotated)

# Sample groups used across the panels
sample_reps <- c("libFL001_STAP_zfh1_replicate1",   "libFL001_STAP_zfh1_replicate2",
                 "libFL002_STAP_CG4822_replicate1", "libFL002_STAP_CG4822_replicate2",
                 "libFL003_STAP_Pld_replicate1",    "libFL003_STAP_Pld_replicate2")
pooled_samples <- c("libFL001_STAP_zfh1_pool",
                    "libFL002_STAP_CG4822_pool",
                    "libFL003_STAP_Pld_pool")
```

```{r, echo = FALSE}
out_dir <- "output/Fig1/"
dir.create(out_dir, recursive = TRUE, showWarnings = FALSE)
```

## FIGURE 1D — Burst parameter distributions

Overall distributions of the three inferred burst parameters (burst size, burst
frequency, and mean expression) for the zfh1 library, quality-filtered
(KS-pval > 0.01 & mu threshold).

```{r, echo = FALSE, warning = FALSE}
# --- FIGURE 1D: Histograms of burst size, frequency, and mean expression ------
# Use the zfh1 pool (main experiment) with quality filters applied
fig1d_data <- STAP_data_annotated %>%
  filter(sample == "libFL001_STAP_zfh1_pool",
         KS_pval > 0.01,
         high_mu == TRUE)

# Burst Size histogram
p_bsize <- ggplot(fig1d_data, aes(x = log10(bsize))) +
  geom_histogram(bins = 50, fill = "orange", color = "black", alpha = 0.7) +
  labs(x = "Burst Size (log10)", y = "count") +
  theme_classic(base_size = 12)

# Burst Frequency histogram
p_bfreq <- ggplot(fig1d_data, aes(x = log10(bfreq))) +
  geom_histogram(bins = 50, fill = "orange", color = "black", alpha = 0.7) +
  labs(x = "Burst Freq. (log10)", y = "count") +
  theme_classic(base_size = 12)

# Mean Expression histogram
p_exp <- ggplot(fig1d_data, aes(x = log10(exp))) +
  geom_histogram(bins = 50, fill = "orange", color = "black", alpha = 0.7) +
  labs(x = "Mean Exp. (log10)", y = "count") +
  theme_classic(base_size = 12)

# Stack the three histograms vertically (as shown in Figure 1D)
p_fig1d <- p_bsize / p_bfreq / p_exp
print(p_fig1d)

# Optional manual PDF export:
# pdf(paste0(out_dir, "Fig1D_burst_parameter_histograms.pdf"), width = 4, height = 6)
# print(p_fig1d)
# dev.off()
```

## FIGURE 1B — BARe-seq vs. STAP-seq promoter activity

Scatter of BARe-seq zfh1-pool activity against the published STAP-seq
reference activity (R = 0.94 in the figure).

```{r, echo = FALSE}
# --- FIGURE 1B: log2FC BARe-seq (y) vs. log2FC STAP-seq (x) -------------------
fitlm = lm((log2FoldChange_zfh1) ~ (log2FoldChange_STAPseq_zfh1), data = diff_exp_sub2)
PCC <- cor((diff_exp_sub2$log2FoldChange_STAPseq_zfh1), diff_exp_sub2$log2FoldChange_zfh1,
           method = 'pearson', use = "complete.obs")

ggplot(diff_exp_sub2) +
  geom_point(aes(x = log2FoldChange_STAPseq_zfh1, y = log2FoldChange_zfh1),
             color = 'grey34', alpha = 0.3) +
  geom_abline(aes(intercept = fitlm$coefficients[1], slope = fitlm$coefficients[2]),
              linetype = "dashed") +
  ggtitle("BARe-seq library activity vs. STAP-seq (zfh1)") +
      labs(title = paste0("PCC = ", round(PCC, digits = 2))) +
      ylab("log2FC BARe-seq promoter activity") +
      xlab("log2FC STAP-seq promoter activity") +
  xlim(-3, 15) +
  ylim(-3, 15) +
  coord_fixed() +
  theme_classic(base_size = 15)
```

## QC / exploratory — other enhancer conditions vs. STAP-seq

Not a main figure panel. Same scatter as Fig. 1B but for the CG4822 and Pld
enhancer conditions; both behave as developmental-class promoters.

```{r, echo = FALSE, warning=FALSE}
# --- QC: CG4822 pool vs. STAP-seq reference ----------------------------------
fitlm = lm((log2FoldChange_CG4822) ~ (log2FoldChange_STAPseq_zfh1), data = diff_exp_sub2)
PCC <- cor((diff_exp_sub2$log2FoldChange_STAPseq_zfh1), diff_exp_sub2$log2FoldChange_CG4822,
           method = 'pearson', use = "complete.obs")

p_CG4822 <- diff_exp_sub2 %>% ggplot() +
  geom_point(aes(x = log2FoldChange_STAPseq_zfh1, y = log2FoldChange_CG4822),
             color = 'grey32', alpha = 0.3) +
  geom_abline(aes(intercept = fitlm$coefficients[1], slope = fitlm$coefficients[2]),
              linetype = "dashed") +
  ggtitle("BARe-seq library activity vs. STAP-seq") +
      labs(title = paste0("PCC = ", round(PCC, digits = 2)),
           subtitle = paste0("R^2 = ", round(summary(fitlm)$r.squared, digits = 2))) +
      ylab("log2FC CG4822 enh") +
      xlab("log2FC STAP-seq promoter") +
  xlim(-3, 15) +
  ylim(-3, 15) +
  coord_fixed() +
  theme_classic(base_size = 15)
p_CG4822
# Optional manual PDF export:
# pdf(paste0(out_dir,"log2FC_CG4822.pdf"), width = 5, height = 5)
# print(p_CG4822)
# dev.off()

# --- QC: Pld pool vs. STAP-seq reference -------------------------------------
fitlm = lm((log2FoldChange_Pld) ~ (log2FoldChange_STAPseq_zfh1), data = diff_exp_sub2)
PCC <- cor((diff_exp_sub2$log2FoldChange_STAPseq_zfh1), diff_exp_sub2$log2FoldChange_Pld,
           method = 'pearson', use = "complete.obs")

p_Pld <- diff_exp_sub2 %>% ggplot() +
  geom_point(aes(x = log2FoldChange_STAPseq_zfh1, y = log2FoldChange_Pld),
             color = 'grey32', alpha = 0.3) +
  geom_abline(aes(intercept = fitlm$coefficients[1], slope = fitlm$coefficients[2]),
              linetype = "dashed") +
  ggtitle("BARe-seq library activity vs. STAP-seq") +
      labs(title = paste0("PCC = ", round(PCC, digits = 2)),
           subtitle = paste0("R^2 = ", round(summary(fitlm)$r.squared, digits = 2))) +
      ylab("log2FC Pld enh") +
      xlab("log2FC STAP-seq promoter") +
  xlim(-3, 15) +
  ylim(-3, 15) +
  coord_fixed() +
  theme_classic(base_size = 15)
p_Pld
# Optional manual PDF export:
# pdf(paste0(out_dir,"log2FC_Pld.pdf"), width = 5, height = 5)
# print(p_Pld)
# dev.off()
```

## SUPPL. FIG. 2C — mean-expression (mu) histograms with threshold

Per-sample distribution of inferred mean expression (mu) with the
mean-expression threshold drawn as the red dashed line.

```{r, echo = FALSE, warning=FALSE}
# --- SUPPL. FIG. 2C: mu histograms, one facet per sample ---------------------
p_mu <- STAP_data_annotated %>% filter(sample %in% c(sample_reps, pooled_samples)) %>%
  ggplot() +
  geom_histogram(aes(x = log10(exp)), bins = 50) +
  geom_vline(aes(xintercept = log10(mu_threshold_nofilter)),
             colour = "red", linetype = "dashed") +
  facet_wrap(~sample, ncol = 3) +
  theme_classic(base_size = 15)

# Recompute high_mu from the per-sample mu threshold (matches Table5's flag);
# kept here so the threshold line and the flag are guaranteed consistent.
STAP_data_annotated[, high_mu := exp >= mu_threshold_nofilter]

# Optional manual PDF export:
# pdf(paste0(out_dir,"STAP_samples_mu_threshold.pdf"), width = 6, height = 6)
# print(p_mu)
# dev.off()

# QC variant: same histograms after removing sequences that fail the KS fit test
p_mu_KS <- STAP_data_annotated %>% filter(sample %in% c(sample_reps, pooled_samples)) %>%
  filter(KS_pval > 0.01) %>%
  ggplot() +
  geom_histogram(aes(x = log10(exp)), bins = 50) +
  geom_vline(aes(xintercept = log10(mu_threshold_nofilter)),
             colour = "red", linetype = "dashed") +
  facet_wrap(~sample, ncol = 3) +
  theme_classic(base_size = 15)
p_mu_KS
```

```{r, echo = FALSE, warning = FALSE}
# --- Build STAP_annotated: burst parameters joined to per-condition log2FC ----
# Melt the wide log2FC table to long format so each (condition, seq_ID) carries
# its log2FC, then join onto the burst-parameter table.
diff_exp_long <- melt(as.data.table(diff_exp_sub2),
                    id.vars = c("oligo_id", "classification"),
                    measure.vars = patterns("^log2FoldChange_"),
                    variable.name = "condition",
                    value.name = "log2FoldChange")

# Strip the "log2FoldChange_" prefix so condition matches STAP_data_annotated
# (zfh1 / CG4822 / Pld). Note the STAP-seq reference column becomes
# "STAPseq_zfh1" here and simply does not match any burst-parameter condition.
diff_exp_long[, condition := gsub("log2FoldChange_", "", condition)]
colnames(diff_exp_long)[1] <- "seq_ID"

STAP_annotated <- merge(STAP_data_annotated, diff_exp_long,
                        by = c("condition", "seq_ID"), all.x = TRUE)
```

## QC / exploratory — log2FC vs. mu coloured by each filter

Diagnostic plots showing how the mu threshold and the KS-fit test partition the
pooled samples. Used to set the filters shown in Suppl. Fig. 2E.

```{r, echo = FALSE}
# --- QC: coloured by mu threshold --------------------------------------------
p_mu <- STAP_annotated %>% filter(sample %in% pooled_samples) %>%
  ggplot() +
  geom_point(aes(y = log2FoldChange.x, x = log10(exp), color = high_mu), alpha = 0.5) +
  scale_color_manual(values = c("grey", "red")) +
  facet_wrap(~sample) +
  ggtitle("log2FC vs. log10(mu) coloured by mu threshold") +
  theme_classic(base_size = 15)
p_mu

# --- QC: coloured by KS-fit test ---------------------------------------------
p_KS <- STAP_annotated %>% filter(sample %in% pooled_samples) %>%
  mutate(KS_cutoff = if_else(KS_pval > 0.01, TRUE, FALSE)) %>%
  ggplot() +
  geom_point(aes(y = log2FoldChange.x, x = log10(exp), color = KS_cutoff), alpha = 0.5) +
  scale_color_manual(values = c("grey", "red")) +
  facet_wrap(~sample) +
  ggtitle("log2FC vs. log10(mu) coloured by KS.pval") +
  theme_classic(base_size = 15)
p_KS

# --- QC: coloured by both filters combined -----------------------------------
p_mu_KS <- STAP_annotated %>% filter(sample %in% pooled_samples) %>%
  mutate(KS_cutoff = if_else(KS_pval > 0.01 & high_mu, TRUE, FALSE)) %>%
  ggplot() +
  geom_point(aes(y = log2FoldChange.x, x = log10(exp), color = KS_cutoff), alpha = 0.5) +
  scale_color_manual(values = c("grey", "red")) +
  facet_wrap(~sample) +
  ggtitle("log2FC vs. log10(mu) coloured by KS.pval & mu threshold") +
  theme_classic(base_size = 15)
p_mu_KS

# Quick counts of sequences passing the KS test in each zfh1 replicate
STAP_annotated %>% filter(rep == "replicate1" & KS_pval > 0.01 & condition == "zfh1")
STAP_annotated %>% filter(rep == "replicate2" & KS_pval > 0.01 & condition == "zfh1")
```

## SUPPL. FIG. 2E — log2FC scatter coloured by quality filters

Same scatter as Fig. 1B, coloured by which sequences pass the combined
KS-fit + mu-threshold filter (the right-hand panel of Suppl. Fig. 2E).

```{r, echo = FALSE}
# --- SUPPL. FIG. 2E: BARe-seq vs. STAP-seq coloured by KS & mu filters --------
fitlm = lm((log2FoldChange_zfh1) ~ (log2FoldChange_STAPseq_zfh1), data = diff_exp_sub2)
PCC <- cor((diff_exp_sub2$log2FoldChange_STAPseq_zfh1), diff_exp_sub2$log2FoldChange_zfh1,
           method = 'pearson', use = "complete.obs")

# Attach the quality flags from the zfh1 pool (the experiment shown here)
diff_exp_sub2 <- merge(
  diff_exp_sub2,
  STAP_annotated[condition == "zfh1" & rep == "pool", .(seq_ID, KS_pval, high_mu)],
  by.x = "oligo_id", by.y = "seq_ID", all.x = TRUE)

p_color <- diff_exp_sub2 %>%
  mutate(KS_cutoff = if_else(KS_pval > 0.01 & high_mu == TRUE, TRUE, FALSE)) %>%
  filter(
    !is.na(KS_cutoff),
    !is.na(log2FoldChange_STAPseq_zfh1),
    !is.na(log2FoldChange_zfh1)
  ) %>%
  ggplot() +
  geom_point(aes(x = log2FoldChange_STAPseq_zfh1, y = log2FoldChange_zfh1, color = KS_cutoff),
             alpha = 0.3) +
  geom_abline(aes(intercept = fitlm$coefficients[1], slope = fitlm$coefficients[2]),
              linetype = "dashed") +
  scale_color_manual(values = c("grey", "red", "grey")) +
  ggtitle("BARe-seq library activity vs. STAP-seq (zfh1)") +
      labs(title = paste0("PCC = ", round(PCC, digits = 2))) +
      ylab("log2FC BARe-seq promoter activity") +
      xlab("log2FC STAP-seq promoter activity") +
  coord_fixed(xlim = c(-3, 15), ylim = c(-3, 15)) +
  theme_classic(base_size = 15)
p_color

# Optional manual PDF export:
# pdf(paste0(out_dir,"log2FC_zfh1_KS_pval_mu.pdf"), width = 5, height = 5)
# print(p_color)
# dev.off()
```

## SUPPL. FIG. 2B / 2D — replicate correlations

Pairwise replicate correlations of the burst parameters for the zfh1 library.
The same comparison is drawn for three filtering levels:

  * "unfiltered"    -> Suppl. Fig. 2B (all sequences)
  * "mu_and_KSpval" -> Suppl. Fig. 2D (KS-pval > 0.01 & mu threshold)

("KS_pval_0.01" is an intermediate QC level kept for reference.)

```{r}
# Datasets at the three filtering levels
data_list <- list(
  "unfiltered"    = STAP_annotated,
  "KS_pval_0.01"  = STAP_annotated %>% filter(KS_pval > 0.01),
  "mu_and_KSpval" = STAP_annotated %>% filter(KS_pval > 0.01 & high_mu == TRUE)
)

# Replicate pair to compare (zfh1 replicate 1 vs. replicate 2)
comparisons <- list(
  zfh1_reps = c("libFL001_STAP_zfh1_replicate1", "libFL001_STAP_zfh1_replicate2")
)
```

```{r, echo = FALSE}
# Compare two samples for burst size, burst frequency and expression.
# Sample name layout used for parsing: lib_experiment_condition_rep
replicate_comparison <- function(comp, data, dataset_name = "") {
  l1 <- strsplit(comp[1], "_")[[1]][1]  # library
  e1 <- strsplit(comp[1], "_")[[1]][2]  # experiment
  r1 <- strsplit(comp[1], "_")[[1]][4]  # replicate

  l2 <- strsplit(comp[2], "_")[[1]][1]
  e2 <- strsplit(comp[2], "_")[[1]][2]
  r2 <- strsplit(comp[2], "_")[[1]][4]

  d1 <- as.data.table(data)[experiment == e1 & lib == l1 & rep == r1,
                    c("bsize", "bfreq", "exp", "seq_ID")]
  colnames(d1)[1:3] <- paste(colnames(d1)[1:3], e1, l1, r1, sep = "_")

  d2 <- as.data.table(data)[experiment == e2 & lib == l2 & rep == r2,
                    c("bsize", "bfreq", "exp", "seq_ID")]
  colnames(d2)[1:3] <- paste(colnames(d2)[1:3], e2, l2, r2, sep = "_")

  cast <- merge(d1, d2, by = "seq_ID", allow.cartesian = TRUE)

  if (nrow(cast) == 0) {
    warning(paste("No data overlap for comparison:", comp[1], "vs", comp[2],
                  "in dataset:", dataset_name))
    return(NULL)
  }

  # --- burst size ---
  x.size <- paste("bsize", e1, l1, r1, sep = "_")
  y.size <- paste("bsize", e2, l2, r2, sep = "_")
  df_size <- data.frame(x = log10(cast[[x.size]]), y = log10(cast[[y.size]]))
  df_size <- df_size[is.finite(df_size$x) & is.finite(df_size$y), ]
  if (nrow(df_size) < 2) return(NULL)
  fitlm <- lm(y ~ x, data = df_size)
  PCC <- cor(df_size$x, df_size$y, method = "pearson", use = "complete.obs")

  size <- ggplot(df_size, aes(x = x, y = y)) +
    geom_point(alpha = 0.2) +
    geom_smooth(method = "lm", formula = y ~ x, se = TRUE, color = "red", linetype = "dashed") +
    geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "darkgrey") +
    coord_cartesian(xlim = c(-1.5, 1.5), ylim = c(-1.5, 1.5)) +
    labs(
      title = paste0("Burst size (Dataset: ", dataset_name, ")"),
      subtitle = paste0(comp[1], " vs ", comp[2],
                    "\nPCC = ", round(PCC, 2),
                    "   R² = ", round(summary(fitlm)$r.squared, 2)),
      x = paste0("log10(burst_size_", comp[1], ")"),
      y = paste0("log10(burst_size_", comp[2], ")")
    ) +
    theme_classic()

  # --- burst frequency ---
  x.freq <- paste("bfreq", e1, l1, r1, sep = "_")
  y.freq <- paste("bfreq", e2, l2, r2, sep = "_")
  df_freq <- data.frame(x = log10(cast[[x.freq]]), y = log10(cast[[y.freq]]))
  df_freq <- df_freq[is.finite(df_freq$x) & is.finite(df_freq$y), ]
  if (nrow(df_freq) < 2) return(NULL)
  fitlm_freq <- lm(y ~ x, data = df_freq)
  PCC_freq <- cor(df_freq$x, df_freq$y, method = "pearson", use = "complete.obs")

  freq <- ggplot(df_freq, aes(x = x, y = y)) +
    geom_point(alpha = 0.2) +
    geom_smooth(method = "lm", formula = y ~ x, se = TRUE, color = "red", linetype = "dashed") +
    geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "darkgrey") +
    coord_cartesian(xlim = c(-2, 0.5), ylim = c(-2, 0.5)) +
    labs(
      title = paste0("Burst frequency (Dataset: ", dataset_name, ")"),
      subtitle = paste0(comp[1], " vs ", comp[2],
                    "\nPCC = ", round(PCC_freq, 2),
                    "   R² = ", round(summary(fitlm_freq)$r.squared, 2)),
      x = paste0("log10(burst_freq_", comp[1], ")"),
      y = paste0("log10(burst_freq_", comp[2], ")")
    ) +
    theme_classic()

  # --- expression ---
  x.exp <- paste("exp", e1, l1, r1, sep = "_")
  y.exp <- paste("exp", e2, l2, r2, sep = "_")
  df_exp <- data.frame(x = log10(cast[[x.exp]]), y = log10(cast[[y.exp]]))
  df_exp <- df_exp[is.finite(df_exp$x) & is.finite(df_exp$y), ]
  if (nrow(df_exp) < 2) return(NULL)
  fitlm_exp <- lm(y ~ x, data = df_exp)
  PCC_exp <- cor(df_exp$x, df_exp$y, method = "pearson", use = "complete.obs")

  exp_plot <- ggplot(df_exp, aes(x = x, y = y)) +
    geom_point(alpha = 0.2) +
    geom_smooth(method = "lm", formula = y ~ x, se = TRUE, color = "red", linetype = "dashed") +
    geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "darkgrey") +
    coord_cartesian(xlim = c(-1.5, 1.5), ylim = c(-1.5, 1.5)) +
    labs(
      title = paste0("Expression (Dataset: ", dataset_name, ")"),
      subtitle = paste0(comp[1], " vs ", comp[2],
                    "\nPCC = ", round(PCC_exp, 2),
                    "   R² = ", round(summary(fitlm_exp)$r.squared, 2)),
      x = paste0("log10(exp_", comp[1], ")"),
      y = paste0("log10(exp_", comp[2], ")")
    ) +
    theme_classic()

  list(freq = freq, size = size, exp = exp_plot)
}

# Render all comparisons for one dataset and also write them to a PDF.
process_and_save_plots <- function(data, name, comparisons) {
  test <- lapply(comparisons, function(comp) {
    replicate_comparison(comp, data, dataset_name = name)
  })

  pdf_filename <- paste0(out_dir, "Comparisons_", name, ".pdf")
  pdf(pdf_filename, width = 4, height = 4)
  for (res in test) {
    if (!is.null(res)) {
      print(res$freq)
      print(res$size)
      print(res$exp)
    }
  }
  dev.off()
  return(test)
}
```

```{r}
# Suppl. Fig. 2B  = "unfiltered" panels
# Suppl. Fig. 2D  = "mu_and_KSpval" panels
results <- Map(function(data, name) {
  message(paste("Processing dataset:", name))
  process_and_save_plots(data, name, comparisons)
},
data_list,
names(data_list))

for (dname in names(data_list)) {
  cat("## Dataset:", dname, "\n\n")
  for (cname in names(comparisons)) {
    cat("### Comparison:", cname, "\n\n")
    plots <- replicate_comparison(
      comparisons[[cname]],
      data_list[[dname]],
      dataset_name = dname
    )
    if (!is.null(plots)) {
      print(plots$freq)
      print(plots$size)
      print(plots$exp)
    }
  }
}
```

## QC / exploratory — replicate correlation heatmaps

Not a main figure panel. Pairwise Pearson correlation of each burst parameter
across the replicate samples, for each filtering level.

```{r, echo = FALSE}
make_correlation_heatmap <- function(dt, feature, dataset_name, sample_reps = NULL) {
  dt <- as.data.table(dt)
  if (!"sample" %in% names(dt)) stop("No 'sample' column found in data")

  wide_dt <- dt[, .(
    seq_ID,
    sample_label = sample,
    value = as.numeric(get(feature))
  )]

  # Keep only the requested samples
  if (!is.null(sample_reps)) {
    wide_dt <- wide_dt[sample_label %in% sample_reps]
  }
  if (nrow(wide_dt) == 0) {
    warning(paste("No data left after filtering for sample_reps in dataset:", dataset_name))
    return(NULL)
  }

  wide_dt <- dcast(wide_dt, seq_ID ~ sample_label, value.var = "value")

  mat <- as.matrix(wide_dt[, -1, with = FALSE])
  storage.mode(mat) <- "numeric"
  mat[mat <= 0] <- NA
  mat <- log10(mat)
  mat <- mat[, colSums(is.finite(mat)) >= 2, drop = FALSE]
  if (ncol(mat) < 2) return(NULL)

  cormat <- cor(mat, use = "pairwise.complete.obs", method = "pearson")
  cormat_tbl <- as.data.frame(as.table(cormat))
  colnames(cormat_tbl) <- c("Sample1", "Sample2", "Correlation")

  ggplot(cormat_tbl, aes(Sample1, Sample2, fill = Correlation)) +
    geom_tile(color = "white") +
    scale_fill_gradient2(
      low = "#4575b4", mid = "white", high = "#d73027",
      midpoint = 0.5, limits = c(0, 1), na.value = "grey90"
    ) +
    geom_text(aes(label = ifelse(is.na(Correlation), "NA", sprintf("%.2f", Correlation))), size = 3) +
    labs(title = paste0("Dataset: ", dataset_name, " | Feature: ", feature), x = "", y = "") +
    theme_minimal(base_size = 10) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          panel.grid = element_blank())
}
```

```{r}
features <- c("exp", "bfreq", "bsize")
all_heatmaps <- list()

for (dname in names(data_list)) {
  cat("Processing dataset:", dname, "\n")
  dt <- data_list[[dname]]

  for (f in features) {
    p <- make_correlation_heatmap(dt, f, dname, sample_reps = sample_reps)
    if (!is.null(p)) {
      heatmap_name <- paste(dname, f, sep = "_")
      all_heatmaps[[heatmap_name]] <- p
      file_path <- paste0(out_dir, "Heatmap_", heatmap_name, ".pdf")
      ggsave(filename = file_path, plot = p, width = 5, height = 5, units = "in")
    }
  }
}

for (nm in names(all_heatmaps)) {
  print(all_heatmaps[[nm]])
}
```
