---
title: "promoter-BARe-seq: Figure 1C and Supplemental Figure 1 — count-based analyses"
author: "FL"
date: "2025-02-13"
output: html_document
---

<!--
================================================================================
OVERVIEW  — Figure 1C and Supplemental Figure 1
================================================================================
This document reproduces the count-based panels of Figure 1 and Supplemental
Figure 1:

  Figure 1C: For individual promoters, shows the observed distribution of
  reporter transcripts per plasmid (the "transcript/plasmid" histogram)
  together with the fitted 0-truncated negative-binomial model. The fitted
  burst parameters (burst size, burst frequency, mean expression) are printed
  on each panel.

  Supplemental Figure 1B: Library-level pUMI distribution showing the number
  of times each plasmid UMI was observed.

  Supplemental Figure 1D: Replicate comparison for two example promoters
  (TATA and DPE), showing consistency of transcript/plasmid distributions
  across biological replicates.

Worked example: library libFL001 (STAP, zfh1). Only this one library is
included here as a self-contained example; the same code runs unchanged on any
other library.

Inputs
------
1. Raw UMI counts (from the GEO repository accompanying the publication),
   one file per replicate. Columns:
       sample_name        publication sample name
       seq_ID             promoter identifier
       plasmid_UMI_main   unique plasmid UMI
       N                  number of reporter transcripts on that plasmid
   Expected in ./data :
       libFL001_STAP_zfh1_replicate1_batch1_counts.txt
       libFL001_STAP_zfh1_replicate2_batch2_counts.txt

2. Table5_promoter_burst_parameters.xlsx (sheet "STAP"), expected in ./Tables .
   Provides the inferred burst parameters (bsize, bfreq, exp) per sample; these
   define the NB fit line drawn over each histogram.

Model
-----
Transcript-per-plasmid counts are modelled with a 0-truncated negative
binomial. In this parametrisation:
    mean expression  (exp)   = mu     (the NB mean)
    burst frequency  (bfreq) = size   (the NB size / dispersion parameter)
    burst size       (bsize) = derived (mu / size)
================================================================================
-->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(data.table)
require(dplyr)
require(ggplot2)
require(readxl)
require(gridExtra)
require(patchwork)
```

## Data loading and pooling

```{r, echo = FALSE}
# ------------------------------------------------------------------------------
# 1. Read the raw UMI counts for the two replicates of libFL001.
#    (Obtain these files from the GEO data repository; place them in ./data .)
# ------------------------------------------------------------------------------
rep1 <- fread("/groups/stark/lorbeer/BARe-seq_GEO_submission_v2/libFL001_STAP_zfh1_replicate1_batch1_counts.txt")
rep2 <- fread("/groups/stark/lorbeer/BARe-seq_GEO_submission_v2/libFL001_STAP_zfh1_replicate2_batch2_counts.txt")

# Use the publication sample naming and a consistent column name.
setnames(rep1, "sample_name", "sample")
setnames(rep2, "sample_name", "sample")
rep1[, sample := "libFL001_STAP_zfh1_replicate1"]
rep2[, sample := "libFL001_STAP_zfh1_replicate2"]

# ------------------------------------------------------------------------------
# 2. Build the pooled sample.
#    The pooled burst parameters in Table5 are inferred from the two
#    replicates' UMIs combined, so the pool is simply the union of the two
#    replicates' count rows, relabelled with the pooled sample name.
# ------------------------------------------------------------------------------
pool <- rbindlist(list(rep1, rep2), use.names = TRUE)
pool[, sample := "libFL001_STAP_zfh1_pool"]

# ------------------------------------------------------------------------------
# 3. Stack replicates + pool into a single long count table and split the
#    sample name into its components (lib / experiment / condition / rep).
# ------------------------------------------------------------------------------
count_data <- rbindlist(list(rep1, rep2, pool), use.names = TRUE)
count_data[, c("library", "experiment", "condition", "rep") :=
             tstrsplit(sample, "_", keep = 1:4)]

# ------------------------------------------------------------------------------
# 4. Burst parameters for these samples (the NB fit lines come from here).
# ------------------------------------------------------------------------------
burst_params <- as.data.table(
  read_xlsx("/groups/stark/lorbeer/BARE-seq_data_prep/Tables/Table5_promoter_burst_parameters.xlsx", sheet = "STAP"))
burst_params <- burst_params[grepl("^libFL001_", sample)]

out_dir <- "output/Fig1/"
dir.create(out_dir, recursive = TRUE, showWarnings = FALSE)
```

## NB-fit plotting function

```{r, echo = FALSE}
# 0-truncated negative binomial density.
# Parametrisation: mu = mean expression (exp), size = burst frequency (bfreq).
d0T_NB <- function(x, mu, size) {
  dnbinom(x, mu = mu, size = size) / (1 - dnbinom(0, mu = mu, size = size))
}

# For each requested promoter (seq_id) in a chosen sample, draw the observed
# transcript/plasmid histogram (relative frequency) with the fitted 0-truncated
# NB density on top, annotated with burst size / frequency / mean expression.
# Promoters are shown as stacked facets (as in Figure 1C).
plot_nb_fit <- function(count_data, burst_params, seq_ids, sample_sel,
                        x_max = 25) {

  count_data   <- as.data.table(count_data)
  burst_params <- as.data.table(burst_params)

  obs_list <- list(); fit_list <- list(); lab_list <- list()

  for (sid in seq_ids) {
    # observed transcript/plasmid counts for this promoter in the chosen sample
    x   <- count_data[sample == sample_sel & seq_ID == sid, N]
    par <- burst_params[sample == sample_sel & seq_ID == sid]

    if (length(x) == 0 || nrow(par) == 0) {
      warning(paste("No counts/parameters for", sid, "in", sample_sel))
      next
    }

    mu <- par$exp[1]     # NB mean            = mean expression
    sz <- par$bfreq[1]   # NB size            = burst frequency
    bs <- par$bsize[1]   # derived burst size

    # keep promoter order stable across the facets
    ord <- match(sid, seq_ids)

    obs_list[[sid]] <- data.frame(N = x, seq_ID = sid, ord = ord)

    xr <- 1:max(x)
    fit_list[[sid]] <- data.frame(x = xr,
                                  density = d0T_NB(xr, mu = mu, size = sz),
                                  seq_ID = sid, ord = ord)

    lab_list[[sid]] <- data.frame(
      seq_ID = sid, ord = ord,
      label  = sprintf("size = %.1f\nfreq = %.2f\nexp = %.2f", bs, sz, mu))
  }

  obs <- bind_rows(obs_list)
  fit <- bind_rows(fit_list)
  lab <- bind_rows(lab_list)

  # order facets as requested
  lev <- unique(obs$seq_ID[order(obs$ord)])
  obs$seq_ID <- factor(obs$seq_ID, levels = lev)
  fit$seq_ID <- factor(fit$seq_ID, levels = lev)
  lab$seq_ID <- factor(lab$seq_ID, levels = lev)

  ggplot(obs, aes(x = N)) +
    geom_histogram(aes(y = after_stat(density)), binwidth = 1,
                   fill = "#9ecae1", colour = "white", alpha = 0.9) +
    geom_line(data = fit, aes(x = x, y = density),
              colour = "orange", linewidth = 1) +
    geom_text(data = lab, aes(label = label),
              x = Inf, y = Inf, hjust = 1.1, vjust = 1.1, size = 3.5) +
    facet_wrap(~ seq_ID, ncol = 1, scales = "free_y") +
    coord_cartesian(xlim = c(0, x_max)) +
    labs(x = "Transcript/plasmid", y = "Relative frequency") +
    theme_classic(base_size = 13)
}
```

## SUPPLEMENTAL FIGURE 1B — library pUMI distribution

Distribution of the number of times each plasmid UMI (pUMI) is observed in the
sequenced input (plasmid) library. The vast majority of pUMIs are observed
exactly once (98.3% in the figure), confirming that each pUMI marks a distinct
plasmid molecule rather than a PCR/sequencing duplicate.

Note: this panel is built from the INPUT library counts, not the RNA counts.
The required file is the input (plasmid) count table for one replicate:
    libFL001_input_zfh1_replicate1_counts.txt   (expected in ./data)
It is assumed to share the column layout of the RNA count files
(sample_name, seq_ID, plasmid_UMI_main, N), where N is the number of
observations (reads) supporting each pUMI.

```{r, echo = FALSE, warning = FALSE, message = FALSE, fig.height = 3, fig.width = 4}
# ------------------------------------------------------------------------------
# Read the input (plasmid) library counts. Each row is one plasmid UMI with
# N = number of observations of that pUMI in the sequenced input library.
# ------------------------------------------------------------------------------
input_counts <- fread("/groups/stark/lorbeer/BARe-seq_GEO_submission_v2/libFL001_input_zfh1_replicate1_counts.txt")
setnames(input_counts, "sample_name", "sample")

# Expected plasmid-UMI structure for this library, taken from the libFL library
# specification table. libFL001 (STAP, zfh1) uses the "fix" 30-nt pUMI design:
#   GC NNNN AC NNNN GT NNNN TA NNNN AC NNNN     (N = any of T/G/C/A)
umi_pattern <- "GC[TGCA]{4}AC[TGCA]{4}GT[TGCA]{4}TA[TGCA]{4}AC[TGCA]{4}"

# ------------------------------------------------------------------------------
# 2. Filter: remove unmapped promoters and poly-G plasmid UMIs, then flag
#    whether each pUMI matches the expected structured pattern.
# ------------------------------------------------------------------------------
input_counts <- input_counts[seq_ID != "*" & !grepl("GGGGGG", plasmid_UMI)]
input_counts[, UMI_match := grepl(umi_pattern, plasmid_UMI)]

# ------------------------------------------------------------------------------
# 3. Bin the per-pUMI observation count (column N) into 1, 2, 3, 4+ and plot.
#    (To restrict to structurally valid pUMIs only, set the filter below to
#     input_counts[UMI_match == TRUE].)
# ------------------------------------------------------------------------------
pumi_dist <- input_counts %>%
  mutate(n_obs_capped = if_else(N > 4, 4L, as.integer(N))) %>%
  count(n_obs_capped, name = "count") %>%
  mutate(pct = 100 * count / sum(count))

# Percentage labels on each bar
labels <- pumi_dist %>%
  mutate(label = if_else(pct >= 1,
                         sprintf("%.1f%%", pct),
                         sprintf("%.2f%%", pct)))

p_pumi <- ggplot(pumi_dist, aes(x = factor(n_obs_capped), y = count)) +
  geom_bar(stat = "identity", fill = "grey30", width = 0.7) +
  geom_text(data = labels, aes(label = label), vjust = -0.3, size = 3.5) +
  scale_y_continuous(labels = function(x) x / 1e6,
                     expand = expansion(mult = c(0, 0.15))) +
  labs(x = "No. observations",
       y = expression("count " * x10^6),
       title = "Library pUMIs") +
  theme_classic(base_size = 13)
print(p_pumi)

# Optional manual PDF export:
# pdf(paste0(out_dir, "SuppFig1B_pUMI_distribution.pdf"), width = 4, height = 3)
# print(p_pumi)
# dev.off()
```

## SUPPLEMENTAL FIGURE 1D — replicate / promoter comparison histograms

Two example promoters are compared: the high-burst-size TATA promoter
chrX_9556722_9556855_- and the high-burst-frequency DPE promoter
chr2R_8592612_8592745_+. Figure 1D shows the transcript/plasmid distributions
from two complementary views:

  * LEFT  (facet per promoter, colored by replicate): the two biological
    replicates overlaid on the same promoter, showing replicate reproducibility.
  * RIGHT (facet per replicate, colored by promoter): the two promoters overlaid
    within the same replicate, showing that the promoters have distinct shapes.

```{r, echo = FALSE, warning = FALSE, message = FALSE, fig.height = 5, fig.width = 10}
# The two example promoters highlighted in Suppl. Fig. 1D
suppfig1d_promoters <- c(
  "chr2R_8592612_8592745_+",   # DPE  (orange in the figure)
  "chrX_9556722_9556855_-"     # TATA (cyan in the figure)
)

# Extract counts for these promoters in both replicates
rep_comp <- count_data %>%
  filter(rep %in% c("replicate1", "replicate2"),
         seq_ID %in% suppfig1d_promoters)

# Nicer replicate labels for legends / facets
rep_labels <- c(replicate1 = "Replicate 1", replicate2 = "Replicate 2")
rep_comp <- rep_comp %>%
  mutate(rep_label = rep_labels[rep])

# ------------------------------------------------------------------------------
# LEFT column: overlapping replicates, one facet per promoter.
# Each promoter gets its own two-tone palette (light = rep1, dark = rep2),
# matching the gold (DPE) and cyan (TATA) shading in the figure.
# ------------------------------------------------------------------------------
rep_shades <- c("Replicate 1" = "goldenrod2", "Replicate 2" = "goldenrod4")  # DPE panel
rep_shades_tata <- c("Replicate 1" = "cyan3",  "Replicate 2" = "darkcyan")   # TATA panel

p_dpe <- rep_comp %>%
  filter(seq_ID == "chr2R_8592612_8592745_+") %>%
  ggplot(aes(x = N, fill = rep_label, color = rep_label)) +
  geom_histogram(aes(y = after_stat(density)), binwidth = 1,
                 position = "identity", alpha = 0.5) +
  scale_fill_manual(values = rep_shades, name = "sample") +
  scale_color_manual(values = rep_shades, name = "sample") +
  coord_cartesian(xlim = c(0, 50)) +
  labs(x = "transcript/plasmid", y = "Relative frequency",
       title = "chr2R_8592612_8592745_+") +
  theme_classic(base_size = 13)

p_tata <- rep_comp %>%
  filter(seq_ID == "chrX_9556722_9556855_-") %>%
  ggplot(aes(x = N, fill = rep_label, color = rep_label)) +
  geom_histogram(aes(y = after_stat(density)), binwidth = 1,
                 position = "identity", alpha = 0.5) +
  scale_fill_manual(values = rep_shades_tata, name = "sample") +
  scale_color_manual(values = rep_shades_tata, name = "sample") +
  coord_cartesian(xlim = c(0, 50)) +
  labs(x = "transcript/plasmid", y = "Relative frequency",
       title = "chrX_9556722_9556855_-") +
  theme_classic(base_size = 13)

# ------------------------------------------------------------------------------
# RIGHT column: overlapping promoters, one facet per replicate.
# ------------------------------------------------------------------------------
prom_colors <- c("chr2R_8592612_8592745_+" = "orange",
                 "chrX_9556722_9556855_-"  = "darkcyan")

make_prom_panel <- function(rep_sel, title) {
  rep_comp %>%
    filter(rep == rep_sel) %>%
    ggplot(aes(x = N, fill = seq_ID, color = seq_ID)) +
    geom_histogram(aes(y = after_stat(density)), binwidth = 1,
                   position = "identity", alpha = 0.5) +
    scale_fill_manual(values = prom_colors, name = "promoter") +
    scale_color_manual(values = prom_colors, name = "promoter") +
    coord_cartesian(xlim = c(0, 50)) +
    labs(x = "transcript/plasmid", y = "Relative frequency", title = title) +
    theme_classic(base_size = 13)
}

p_rep1 <- make_prom_panel("replicate1", "Replicate 1")
p_rep2 <- make_prom_panel("replicate2", "Replicate 2")

# ------------------------------------------------------------------------------
# Assemble the 2x2 layout: promoter facets (left) + replicate facets (right).
# ------------------------------------------------------------------------------
p_suppfig1d <- (p_dpe / p_tata) | (p_rep1 / p_rep2)
print(p_suppfig1d)

# Optional manual PDF export:
# pdf(paste0(out_dir, "SuppFig1D_replicate_comparison.pdf"), width = 10, height = 5)
# print(p_suppfig1d)
# dev.off()
```

## FIGURE 1C — example promoters

The three example promoters highlighted in Figure 1B/1C, shown for the pooled
libFL001 sample. Each panel shows the observed transcript/plasmid distribution
(blue) and the fitted 0-truncated NB model (orange), annotated with the
inferred burst parameters.

```{r, echo = FALSE, warning = FALSE, message = FALSE, fig.height = 8, fig.width = 4}
# The three Figure 1C examples (order 1-2-3 as in the figure):
#   1  size ~ 4.2,  freq ~ 0.8,  exp ~ 3.4
#   2  size ~ 16.3, freq ~ 0.05, exp ~ 0.8
#   3  size ~ 1.8,  freq ~ 1.8,  exp ~ 3.2
fig1c_examples <- c(
  "chr2L_21629861_21629994_-",
  "chrX_9556722_9556855_-",
  "chr2R_8592612_8592745_+"
)

p_fig1c <- plot_nb_fit(count_data, burst_params,
                       seq_ids    = fig1c_examples,
                       sample_sel = "libFL001_STAP_zfh1_pool")
print(p_fig1c)

# Optional manual PDF export:
# pdf(paste0(out_dir, "Fig1C_transcript_plasmid_NBfit.pdf"), width = 4, height = 8)
# print(p_fig1c)
# dev.off()
```
