---
title: "promoter-BARe-seq for Figure 1"
author: "FL"
date: "2025-02-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(data.table)
require(dplyr)
require(ggplot2)
require(readxl)
library(ggrepel)
require(patchwork)
require(gridExtra)
require(stringr)
#require(scales)
#library("cowplot")

```

## Data sets for Figure 1 
R version 4.1.2 and Bioconductor 3.14
```{r, echo = FALSE}
#read in the different kinds of data sets required
#annotation file used 
anno <- read_xlsx("/groups/stark/lorbeer/BARE-seq_data_prep/Tables/Table4_promoter_library_annotations.xlsx")
diff_exp_sub2$log2FoldChange_zfh1

new_STAP <- fread("/groups/stark/lorbeer/STAP-TWIST/Franzi_STAP_TWIST_metadata-v2.txt")

#differential expression data set 
# code for generating STAP_diff_exp_E38_E28_combine.R
diff_exp_sub2 <- readRDS("/groups/stark/lorbeer/Burst_params_DG/STAP_March2025_pools/diff_exp_STAPexperiments.rds")
#counts data sets
#load all the counts for E28 and E38
#counts were combined into tables and converted in script STAP_counts_E38_E28_combine.R
count_data <- readRDS("/groups/stark/lorbeer/Burst_params_DG/STAP_March2025_pools/raw_counts_E38_E28_aug.rds")
final_result <- readRDS("/groups/stark/lorbeer/Burst_params_DG/STAP_March2025_pools/counts_data_combined_E38_E28_annotated_augmented.rds")
# raw_counts <- readRDS("/groups/stark/lorbeer/Burst_params_DG/STAP_March2025_pools/raw_counts_E38_E28.rds")
# final_result <- readRDS("/groups/stark/lorbeer/Burst_params_DG/STAP_March2025_pools/counts_data_combined_E38_E28_annotated.rds")

#burst parameter data sets 
#script: STAP_pooled_data_prep_publication.R
STAP_data_annotated <- readRDS("/groups/stark/lorbeer/Burst_params_DG/STAP_March2025_pools/STAP_data_annotated.rds")
STAP_data_annotated[, c("lib","experiment","condition", "rep"):= tstrsplit(sample, "_", keep= c(1:4))]
STAP_data_annotated$rep <- gsub("^r([1-3])$", "rep\\1", STAP_data_annotated$rep)

#some samples have rep4, 5, 6 because of experimental reasons
STAP_data_annotated[rep=="rep4"& condition == "e339", rep:= "rep1"]
STAP_data_annotated[rep=="rep5"& condition == "e339", rep:= "rep2"]
STAP_data_annotated[rep=="rep6"& condition == "e339", rep:= "rep3"]
STAP_data_annotated[rep=="r1"& condition == "RT2", rep:= "rep1"]
STAP_data_annotated[rep=="r2"& condition == "RT2", rep:= "rep2"]
STAP_data_annotated[rep=="r3"& condition == "RT2", rep:= "rep3"]
STAP_data_annotated[sample=="libFL008_E28_RT2_r1", sample:= "libFL008_E28_RT2_rep1"]
STAP_data_annotated[sample=="libFL008_E28_RT2_r2", sample:= "libFL008_E28_RT2_rep2"]
STAP_data_annotated[sample=="libFL008_E28_RT2_r3", sample:= "libFL008_E28_RT2_rep3"]

count_data[sample=="libFL008_E28_RT2_r1", sample:= "libFL008_E28_RT2_rep1"]
count_data[sample=="libFL008_E28_RT2_r2", sample:= "libFL008_E28_RT2_rep2"]
count_data[sample=="libFL008_E28_RT2_r3", sample:= "libFL008_E28_RT2_rep3"]


#read in mu thresholds 
mu_thresholds <- read_xlsx("/groups/stark/lorbeer/Burst_params_DG/mu_thresholds_v2_Reyna.xlsx")
STAP_data_annotated <- merge(STAP_data_annotated, mu_thresholds, by = c("sample", "experiment"), all.x = TRUE)

head(STAP_data_annotated)
sample_reps <- c("libFL008_E28_RT2_rep2", "libFL008_E28_RT2_rep3", #"libFL008_E38_dRT2_pair1", "libFL008_E38_dRT2_pair2", 
                 "libFL016_E38_e325_rep1", "libFL016_E38_e325_rep2", "libFL017_E38_e339_rep5", "libFL017_E38_e339_rep6")
pooled_samples <- c("libFL008_E28_RT2_pooledby2", #"libFL008_E38_dRT2_pool4", 
                    "libFL016_E38_e325_pooledby2", "libFL017_E38_e339_pooledby2")

```

```{r, echo = FALSE}

out_dir <- "/groups/stark/lorbeer/Burst_params_DG/STAP_March2025_pools/Fig1_April26/"

```

## Comparing the data with published STAP-seq data set 
#plots for E28
```{r, echo = FALSE}
#calculate Rsquare and PCC
fitlm = lm((log2FoldChange_RT2)~(log2FoldChange_zfh1) , data = diff_exp_sub2)
PCC <- cor((diff_exp_sub2$log2FoldChange_zfh1), diff_exp_sub2$log2FoldChange_RT2, method = 'pearson', use = "complete.obs")


ggplot(diff_exp_sub2) + 
  geom_point(aes(x = log2FoldChange_zfh1, y = log2FoldChange_RT2), color = 'grey34', alpha = 0.3) + 
  geom_abline(aes(intercept = fitlm$coefficients[1], slope = fitlm$coefficients[2]), linetype = "dashed") + 
  #geom_abline(aes(intercept = 0, slope = 1), linetype = "dashed", color = "red") + 
  ggtitle("BARe-seq library activity vs. STAP-seq E28") + 
      labs(title = paste0("PCC = ", round(PCC, digits = 2)), 
           #subtitle = paste0("R^2 = ", round(summary(fitlm)$r.squared, digits = 2))
           ) + 
      ylab("log2FC BARe-seq promoter activity") + 
      xlab("log2FC STAP-seq promoter activity") + 
  xlim(-3, 15) + 
  ylim(-3,15) + 
  coord_fixed() + 
  theme_classic(base_size = 15)


```


## comparison of the different enhancers used with E28 data 
all turn out to be developmental class by previous definition 
```{r, echo = FALSE, warning=FALSE}

fitlm = lm((log2FoldChange_e325)~(log2FoldChange_zfh1) , data = diff_exp_sub2)
PCC <- cor((diff_exp_sub2$log2FoldChange_zfh1), diff_exp_sub2$log2FoldChange_e325, method = 'pearson', use = "complete.obs")

p_e325 <- diff_exp_sub2 %>% ggplot() + 
  geom_point(aes(x = log2FoldChange_zfh1, y = log2FoldChange_e325), color = 'grey32', alpha = 0.3) + 
  #scale_color_manual(values = c("#1f77b4", "darkorange", "midnightblue", "darkgrey","darkgrey")) + 
  geom_abline(aes(intercept = fitlm$coefficients[1], slope = fitlm$coefficients[2]), linetype = "dashed") + 
  ggtitle("BARe-seq library activity vs. STAP-seq") + 
      labs(title = paste0("PCC = ", round(PCC, digits = 2)), 
           subtitle = paste0("R^2 = ", round(summary(fitlm)$r.squared, digits = 2))) + 
      ylab("log2FC e325") + 
      xlab("log2FC STAP-seq promoter") + 
  xlim(-3, 15) + 
  ylim(-3,15) + 
  coord_fixed() + 
  theme_classic(base_size = 15)
p_e325
#Optional manual PDF export:
# pdf(paste0(out_dir,"log2FC_e325.pdf"), width = 5, height = 5)
# print(p_e325)
# dev.off()

fitlm = lm((log2FoldChange_e339)~(log2FoldChange_zfh1) , data = diff_exp_sub2)
PCC <- cor((diff_exp_sub2$log2FoldChange_zfh1), diff_exp_sub2$log2FoldChange_e339, method = 'pearson', use = "complete.obs")

p_e339 <- diff_exp_sub2 %>% ggplot() + 
  geom_point(aes(x = log2FoldChange_zfh1, y = log2FoldChange_e339), color = 'grey32', alpha = 0.3) + 
  #scale_color_manual(values = c("#1f77b4", "darkorange", "midnightblue", "darkgrey","darkgrey")) + 
  geom_abline(aes(intercept = fitlm$coefficients[1], slope = fitlm$coefficients[2]), linetype = "dashed") + 
  ggtitle("BARe-seq library activity vs. STAP-seq") + 
      labs(title = paste0("PCC = ", round(PCC, digits = 2)), 
           subtitle = paste0("R^2 = ", round(summary(fitlm)$r.squared, digits = 2))) + 
      ylab("log2FC e339") + 
      xlab("log2FC STAP-seq promoter") + 
  xlim(-3, 15) + 
  ylim(-3,15) + 
  coord_fixed() + 
  theme_classic(base_size = 15)
p_e339
# Optional manual PDF export:
# pdf(paste0(out_dir,"log2FC_e339.pdf"), width = 5, height = 5)
# print(p_e339)
# dev.off()
```

##Burst parameters overall assessment 

```{r, echo = FALSE, warning=FALSE}

p_mu <- STAP_data_annotated %>% filter(sample %in% c(sample_reps, pooled_samples)) %>% 
  ggplot() + 
  geom_histogram(aes(x = log10(exp)), bins = 50) + 
  geom_vline(aes(xintercept = log10(mu_threshold_nofilter)), 
             colour = "red", linetype = "dashed") +
  facet_wrap(~sample, ncol = 3) + 
  theme_classic(base_size = 15)

STAP_data_annotated[, high_mu := exp >= mu_threshold_nofilter]

# Optional manual PDF export:
# pdf(paste0(out_dir,"STAP_samples_mu_threshold.pdf"), width = 6, height = 6)
# print(p_mu)
# dev.off()

p_mu_KS <- STAP_data_annotated %>% filter(sample %in% c(sample_reps, pooled_samples)) %>% filter(KS_pval > 0.01) %>%
  ggplot() + 
  geom_histogram(aes(x = log10(exp)), bins = 50) + 
  geom_vline(aes(xintercept = log10(mu_threshold_nofilter)), 
             colour = "red", linetype = "dashed") +
  facet_wrap(~sample, ncol = 3) + 
  theme_classic(base_size = 15)
p_mu_KS

```

```{r, echo = FALSE, warning = FALSE}

# Melt the log2FoldChange columns to long format
diff_exp_long <- melt(as.data.table(diff_exp_sub2), 
                      id.vars = c("oligo_id", "classification"),
                      measure.vars = patterns("^log2FoldChange_"),
                      variable.name = "condition", 
                      value.name = "log2FoldChange")

# Clean up the condition names (remove "log2FoldChange_" prefix)
diff_exp_long[, condition := gsub("log2FoldChange_", "", condition)]
colnames(diff_exp_long)[1] <- "seq_ID"

STAP_annotated <- merge(STAP_data_annotated, diff_exp_long, by = c("condition", "seq_ID"), all.x = TRUE)
#saveRDS(STAP_annotated, "/groups/stark/lorbeer/Burst_params_DG/STAP_March2025_pools/STAP_annotated_Fig1_compiled_v2.rds")

```

```{r, echo = FALSE}

p_mu <- STAP_annotated %>% filter (sample %in% pooled_samples) %>% 
  ggplot() + 
  geom_point(aes( y= log2FoldChange, x = log10(exp), color = high_mu) , alpha = 0.5) + 
  scale_color_manual(values = c("grey", "red")) + 
  facet_wrap(~sample) + 
    ggtitle("log2FC vs. log10(mu) colored by mu threshold") + 
  theme_classic(base_size = 15)
p_mu
# Optional manual PDF export:
# pdf(paste0(out_dir,"STAP_muvlog2FC_mu_threshold.pdf"), width = 6, height = 3)
# print(p_mu)
# dev.off()

p_KS <- STAP_annotated %>% filter (sample %in% pooled_samples) %>% 
  mutate(KS_cutoff = if_else(KS_pval >0.01 ,TRUE,FALSE)) %>% 
  ggplot() + 
  geom_point(aes( y= log2FoldChange, x = log10(exp), color = KS_cutoff) , alpha = 0.5) + 
    scale_color_manual(values = c("grey", "red")) + 
  facet_wrap(~sample) + 
    ggtitle("log2FC vs. log10(mu) colored by KS.pval ") + 
  theme_classic(base_size = 15)
p_KS 
# Optional manual PDF export:
# pdf(paste0(out_dir,"STAP_muvlog2FC_KS_threshold.pdf"), width = 6, height = 3)
# print(p_KS)
# dev.off()

p_mu_KS <- STAP_annotated %>% filter (sample %in% pooled_samples) %>% 
  mutate(KS_cutoff = if_else(KS_pval >0.01 & high_mu ,TRUE,FALSE)) %>% 
  ggplot() + 
  geom_point(aes( y= log2FoldChange, x = log10(exp), color = KS_cutoff) , alpha = 0.5) + 
    scale_color_manual(values = c("grey", "red")) + 
  facet_wrap(~sample) + 
    ggtitle("log2FC vs. log10(mu) colored by KS.pval & mu threshold") + 
  theme_classic(base_size = 15)
p_mu_KS 
# Optional manual PDF export:
# pdf(paste0(out_dir,"STAP_muvlog2FC_muKS_threshold.pdf"), width = 6, height = 3)
# print(p_mu_KS)
# dev.off()

#counting values 
STAP_annotated %>% filter(rep == "rep2" & KS_pval >0.01 & condition == "RT2")
STAP_annotated %>% filter(rep == "rep3" & KS_pval >0.01 & condition == "RT2")

```
```{r, echo = FALSE}

summary_data <- count_data[, .(
  mean_expression = mean(N),
  total_expression = sum(N)
), by = .(seq_ID, sample, library, experiment, enhancer, rep)]

# Ensure STAP_annotated is a data.table
setDT(STAP_annotated)

# Define the columns you want to merge on. 
# Note: I omitted 'enhancer' here assuming it might not be in STAP_annotated based on the output. 
# If 'enhancer' is in STAP_annotated, add it to the 'by' vector!
merge_cols <- c("seq_ID", "sample", "library", "experiment", "rep")

# Merge the datasets (all.x = TRUE performs a left join)
STAP_merged <- merge(
  STAP_annotated, 
  summary_data, 
  by = merge_cols, 
  all.x = TRUE 
)

# View the result
head(STAP_merged)

```

```{r, echo = FALSE}

# -------------------------------------------------------------
# 1) Total expression vs. Mean expression (Faceted by sample)
# -------------------------------------------------------------
plot_expr <- STAP_merged %>% filter(sample %in% c("libFL008_E28_RT2_pooledby2")) %>% mutate(KS_cutoff = if_else(KS_pval >0.01 ,TRUE,FALSE)) %>% 
  ggplot( aes(x = log2(mean_expression), y = log2(total_expression), color = KS_cutoff)) +
  geom_point(alpha = 0.5, size = 1) +
  facet_wrap(~ sample) +
  theme_classic() +
  labs(title = "Total vs. Mean Expression",
       x = "log2( calculated Mean Expression)",
       y = "log2(calcualted total Expression)")

print(plot_expr)
# pdf(paste0(out_dir,"mu_vs_log2FC_regression_KS_color.pdf"), width = 5, height = 5)
# print(plot_expr)
# dev.off()

# -------------------------------------------------------------
# 2) Mean expression vs. mu (Faceted by sample)
# -------------------------------------------------------------
# Note: If your actual mean value is stored in 'exp', change `x = mu` to `x = exp`
plot_mu <- STAP_merged %>% filter(sample %in% c("libFL008_E28_RT2_pooledby2")) %>% mutate(KS_cutoff = if_else(KS_pval >0.01 ,TRUE,FALSE)) %>% 
  ggplot( aes(x = log10(exp), y = log10(mean_expression), color = KS_cutoff)) +
  geom_point(alpha = 0.5, size = 1) +
  facet_wrap(~ sample) +
  theme_classic() +
  labs(title = "Mean Expression vs. Mu",
       x = "Mu log10 - inferred",
       y = "Mean Expression log10 - calculated")

print(plot_mu)
# pdf(paste0(out_dir,"mu_vs_inferred_mu_regression_KScolor.pdf"), width = 5, height = 5)
# print(plot_mu)
# dev.off()


# -------------------------------------------------------------
# 3) Total expression vs. log2FC (Faceted by sample)
# -------------------------------------------------------------
plot_lfc <- STAP_merged %>% filter(sample %in% c("libFL008_E28_RT2_pooledby2")) %>% mutate(KS_cutoff = if_else(KS_pval >0.01 ,TRUE,FALSE)) %>% 
  ggplot(aes(x = log2FoldChange, y = log2(total_expression), color = KS_cutoff)) +
  geom_point(alpha = 0.5, size = 1) +
  facet_wrap(~ sample) +
  theme_classic() +
  labs(title = "Total Expression vs. log2FoldChange",
       x = "log2FoldChange",
       y = "log2(Total Expression)")

print(plot_lfc)

#Optional manual PDF export:
# pdf(paste0(out_dir,"log2FC_total_exp_correlation_KScolor.pdf"), width = 5, height = 5)
# print(plot_lfc)
# dev.off()

```


```{r, echo = FALSE}
library(dplyr)
library(ggplot2)
library(ggpubr)

# Prepare the data once
plot_data <- STAP_merged %>% 
  filter(sample %in% c("libFL008_E28_RT2_pooledby2")) %>% 
  mutate(
    KS_cutoff = if_else(KS_pval > 0.01, TRUE, FALSE),
    Highlight = if_else(KS_cutoff == TRUE & high_mu == TRUE, TRUE, FALSE)
  )

# 1) Total expression vs. Mean expression
plot_expr <- ggplot(plot_data, aes(x = log2(mean_expression), y = log2(total_expression), color = Highlight)) +
  geom_point(alpha = 0.5, size = 1) +
  scale_color_manual(values = c("FALSE" = "grey70", "TRUE" = "firebrick")) +
  facet_wrap(~ sample) +
  theme_classic() +
  labs(title = "Total vs. Mean Expression", x = "log2( calculated Mean Expression )", y = "log2( calculated total Expression )")
print(plot_expr)

# 2) Mean expression vs. mu
plot_mu <- ggplot(plot_data, aes(x = log10(exp), y = log10(mean_expression), color = Highlight)) +
  geom_point(alpha = 0.5, size = 1) +
  scale_color_manual(values = c("FALSE" = "grey70", "TRUE" = "firebrick")) +
  facet_wrap(~ sample) +
  theme_classic() +
  labs(title = "Mean Expression vs. Mu", x = "Mu log10 - inferred", y = "Mean Expression log10 - calculated")
print(plot_mu)

# 3) Total expression vs. log2FC
plot_lfc <- ggplot(plot_data, aes(x = log2FoldChange, y = log2(total_expression), color = Highlight)) +
  geom_point(alpha = 0.5, size = 1) +
  scale_color_manual(values = c("FALSE" = "grey70", "TRUE" = "firebrick")) +
  facet_wrap(~ sample) +
  theme_classic() +
  labs(title = "Total Expression vs. log2FoldChange", x = "log2FoldChange", y = "log2(Total Expression)")
print(plot_lfc)

```

```{r, echo = FALSE}
# 1) Total expression vs. Mean expression (with Regression & Pearson's R)
plot_expr_reg <- ggplot(plot_data, aes(x = log2(mean_expression), y = log2(total_expression), color = Highlight)) +
  geom_point(alpha = 0.5, size = 1) +
  scale_color_manual(values = c("FALSE" = "grey70", "TRUE" = "firebrick")) +
  geom_smooth(data = filter(plot_data, Highlight == TRUE), method = "lm", color = "firebrick") +
  stat_cor(data = filter(plot_data, Highlight == TRUE), aes(label = after_stat(r.label)), color = "black", label.x.npc = "left", label.y.npc = "top") +
  facet_wrap(~ sample) +
  theme_classic() +
  labs(title = "Total vs. Mean Expression (Target Regression)", 
       x = "log2( calculated Mean Expression )", 
       y = "log2( calculated total Expression )")

print(plot_expr_reg)
# 
# pdf(paste0(out_dir,"mu_vs_log2FC_regression.pdf"), width = 5, height = 5)
# print(plot_expr_reg)
# dev.off()

# 2) Mean expression vs. mu (with Regression & Pearson's R)
plot_mu_reg <- ggplot(plot_data, aes(x = log10(exp), y = log10(mean_expression), color = Highlight)) +
  geom_point(alpha = 0.5, size = 1) +
  scale_color_manual(values = c("FALSE" = "grey70", "TRUE" = "firebrick")) +
  geom_smooth(data = filter(plot_data, Highlight == TRUE), method = "lm", color = "firebrick") +
  stat_cor(data = filter(plot_data, Highlight == TRUE), aes(label = after_stat(r.label)), color = "black", label.x.npc = "left", label.y.npc = "top") +
  facet_wrap(~ sample) +
  theme_classic() +
  labs(title = "Mean Expression vs. Mu (Target Regression)", 
       x = "Mu log10 - inferred", 
       y = "Mean Expression log10 - calculated")

print(plot_mu_reg)

# pdf(paste0(out_dir,"mu_vs_inferred_mu_regression.pdf"), width = 5, height = 5)
# print(plot_mu_reg)
# dev.off()

# 3) Total expression vs. log2FC (with Regression & Pearson's R)
plot_lfc_reg <- ggplot(plot_data, aes(x = log2FoldChange, y = log2(total_expression), color = Highlight)) +
  geom_point(alpha = 0.5, size = 1) +
  scale_color_manual(values = c("FALSE" = "grey70", "TRUE" = "firebrick")) +
  geom_smooth(data = filter(plot_data, Highlight == TRUE), method = "lm", color = "firebrick") +
  stat_cor(data = filter(plot_data, Highlight == TRUE), aes(label = after_stat(r.label)), color = "black", label.x.npc = "left", label.y.npc = "top") +
  facet_wrap(~ sample) +
  theme_classic() +
  labs(title = "Total Expression vs. log2FoldChange (Target Regression)", 
       x = "log2FoldChange", 
       y = "log2(Total Expression)")

print(plot_lfc_reg)

#Optional manual PDF export:
# pdf(paste0(out_dir,"log2FC_total_exp_correlation.pdf"), width = 5, height = 5)
# print(plot_lfc_reg)
# dev.off()


```



```{r, echo = FALSE}
#calculate Rsquare and PCC
fitlm = lm((log2FoldChange_RT2)~(log2FoldChange_zfh1) , data = diff_exp_sub2)
PCC <- cor((diff_exp_sub2$log2FoldChange_zfh1), diff_exp_sub2$log2FoldChange_RT2, method = 'pearson', use = "complete.obs")

diff_exp_sub2 <- merge(diff_exp_sub2, STAP_annotated[, c("condition", "rep","seq_ID", "KS_pval", "high_mu"), with = FALSE], all.x = TRUE, by.x = "oligo_id", by.y = "seq_ID")

p_color <- diff_exp_sub2 %>% 
  mutate(KS_cutoff = if_else(KS_pval >0.01 & high_mu == TRUE,TRUE,FALSE)) %>% 
filter(
    !is.na(KS_cutoff),
    !is.na(log2FoldChange_zfh1),
    !is.na(log2FoldChange_RT2)
  ) %>%
  ggplot() + 
  geom_point(aes(x = log2FoldChange_zfh1, y = log2FoldChange_RT2, color = KS_cutoff),alpha = 0.3) + 
  geom_abline(aes(intercept = fitlm$coefficients[1], slope = fitlm$coefficients[2]), linetype = "dashed") + 
      scale_color_manual(values = c("grey", "red", "grey")) + 
  #geom_abline(aes(intercept = 0, slope = 1), linetype = "dashed", color = "red") + 
  ggtitle("BARe-seq library activity vs. STAP-seq E28") + 
      labs(title = paste0("PCC = ", round(PCC, digits = 2)), 
           #subtitle = paste0("R^2 = ", round(summary(fitlm)$r.squared, digits = 2))
           ) + 
      ylab("log2FC BARe-seq promoter activity") + 
      xlab("log2FC STAP-seq promoter activity") + 
  #xlim(-3, 15) + 
  #ylim(-3,15) + 
  #coord_fixed() + 
  coord_fixed(xlim = c(-3, 15), ylim = c(-3, 15)) +
  theme_classic(base_size = 15)
p_color

#Optional manual PDF export:
# pdf(paste0(out_dir,"log2FC_RT2_KS_pval_mu.pdf"), width = 5, height = 5)
# print(p_color)
# dev.off()

```


## pairwise comparisons 
```{r}
# List of filtered datasets and their labels
data_list <- list(
  "unfiltered" = STAP_annotated ,
  "KS_pval_0.01" = STAP_annotated %>% filter(KS_pval > 0.01 ),
  "mu_and_KSpval" = STAP_annotated %>% filter(KS_pval > 0.01 & high_mu == TRUE )
)

#list of comparisons to make
comparisons <- list(
                    h5 = c("libFL008_E28_RT2_rep2", "libFL008_E28_RT2_rep3")#, 
                    )

```

```{r, echo = FALSE}
replicate_comparison <- function(comp, data, dataset_name = "") {
  e1 <- strsplit(comp[1], "_")[[1]][2]
  l1 <- strsplit(comp[1], "_")[[1]][1]
  r1 <- strsplit(comp[1], "_")[[1]][4]

  e2 <- strsplit(comp[2], "_")[[1]][2]
  l2 <- strsplit(comp[2], "_")[[1]][1]
  r2 <- strsplit(comp[2], "_")[[1]][4]

  d1 <- as.data.table(data)[experiment == e1 & lib == l1 & rep == r1,
                            c("bsize", "bfreq", "exp", "seq_ID")]
  colnames(d1)[1:3] <- paste(colnames(d1)[1:3], e1, l1, r1, sep = "_")

  d2 <- as.data.table(data)[experiment == e2 & lib == l2 & rep == r2,
                            c("bsize", "bfreq", "exp", "seq_ID")]
  colnames(d2)[1:3] <- paste(colnames(d2)[1:3], e2, l2, r2, sep = "_")

  cast <- merge(d1, d2, by = "seq_ID", allow.cartesian = TRUE)

  if (nrow(cast) == 0) {
    warning(paste("No data overlap for comparison:", comp[1], "vs", comp[2], "in dataset:", dataset_name))
    return(NULL)
  }

  x.size <- paste("bsize", e1, l1, r1, sep = "_")
  y.size <- paste("bsize", e2, l2, r2, sep = "_")

  df_size <- data.frame(
    x = log10(cast[[x.size]]),
    y = log10(cast[[y.size]])
  )
  df_size <- df_size[is.finite(df_size$x) & is.finite(df_size$y), ]
  if (nrow(df_size) < 2) return(NULL)

  fitlm <- lm(y ~ x, data = df_size)
  PCC <- cor(df_size$x, df_size$y, method = "pearson", use = "complete.obs")

  size <- ggplot(df_size, aes(x = x, y = y)) +
    geom_point(alpha = 0.2) +
    #geom_abline(intercept = coef(fitlm)[1], slope = coef(fitlm)[2],
    #            linetype = "dashed", color = "red") +
    geom_smooth(method = "lm", formula = y ~ x, se = TRUE, color = "red", linetype = "dashed") + 
    geom_abline(intercept = 0, slope = 1,
                linetype = "dashed", color = "darkgrey") +
    coord_cartesian(xlim = c(-1.5, 1.5), ylim = c(-1.5, 1.5)) +
    labs(
      title = paste0("Burst size (Dataset: ", dataset_name, ")"),
      subtitle = paste0(comp[1], " vs ", comp[2],
                        "\nPCC = ", round(PCC, 2),
                        "   R² = ", round(summary(fitlm)$r.squared, 2)),
      x = paste0("log10(burst_size_", comp[1], ")"),
      y = paste0("log10(burst_size_", comp[2], ")")
    ) +
    theme_classic()

  x.freq <- paste("bfreq", e1, l1, r1, sep = "_")
  y.freq <- paste("bfreq", e2, l2, r2, sep = "_")

  df_freq <- data.frame(
    x = log10(cast[[x.freq]]),
    y = log10(cast[[y.freq]])
  )
  df_freq <- df_freq[is.finite(df_freq$x) & is.finite(df_freq$y), ]
  if (nrow(df_freq) < 2) return(NULL)

  fitlm_freq <- lm(y ~ x, data = df_freq)
  PCC_freq <- cor(df_freq$x, df_freq$y, method = "pearson", use = "complete.obs")

  freq <- ggplot(df_freq, aes(x = x, y = y)) +
    geom_point(alpha = 0.2) +
    #geom_abline(intercept = coef(fitlm_freq)[1], slope = coef(fitlm_freq)[2],
    #            linetype = "dashed", color = "red") +
    geom_smooth(method = "lm", formula = y ~ x, se = TRUE, color = "red", linetype = "dashed") + 
    geom_abline(intercept = 0, slope = 1,
                linetype = "dashed", color = "darkgrey") +
    coord_cartesian(xlim = c(-2, 0.5), ylim = c(-2, 0.5)) +
    labs(
      title = paste0("Burst frequency (Dataset: ", dataset_name, ")"),
      subtitle = paste0(comp[1], " vs ", comp[2],
                        "\nPCC = ", round(PCC_freq, 2),
                        "   R² = ", round(summary(fitlm_freq)$r.squared, 2)),
      x = paste0("log10(burst_freq_", comp[1], ")"),
      y = paste0("log10(burst_freq_", comp[2], ")")
    ) +
    theme_classic()

  x.exp <- paste("exp", e1, l1, r1, sep = "_")
  y.exp <- paste("exp", e2, l2, r2, sep = "_")

  df_exp <- data.frame(
    x = log10(cast[[x.exp]]),
    y = log10(cast[[y.exp]])
  )
  df_exp <- df_exp[is.finite(df_exp$x) & is.finite(df_exp$y), ]
  if (nrow(df_exp) < 2) return(NULL)

  fitlm_exp <- lm(y ~ x, data = df_exp)
  PCC_exp <- cor(df_exp$x, df_exp$y, method = "pearson", use = "complete.obs")

  exp_plot <- ggplot(df_exp, aes(x = x, y = y)) +
    geom_point(alpha = 0.2) +
    #geom_abline(intercept = coef(fitlm_exp)[1], slope = coef(fitlm_exp)[2],
    #            linetype = "dashed", color = "red") +
    geom_smooth(method = "lm", formula = y ~ x, se = TRUE, color = "red", linetype = "dashed") + 
    geom_abline(intercept = 0, slope = 1,
                linetype = "dashed", color = "darkgrey") +
    coord_cartesian(xlim = c(-1.5, 1.5), ylim = c(-1.5, 1.5)) +
    labs(
      title = paste0("Expression (Dataset: ", dataset_name, ")"),
      subtitle = paste0(comp[1], " vs ", comp[2],
                        "\nPCC = ", round(PCC_exp, 2),
                        "   R² = ", round(summary(fitlm_exp)$r.squared, 2)),
      x = paste0("log10(exp_", comp[1], ")"),
      y = paste0("log10(exp_", comp[2], ")")
    ) +
    theme_classic()

  list(
    freq = freq,
    size = size,
    exp = exp_plot
  )
}

# Function to process and save plots for each dataset
process_and_save_plots <- function(data, name, comparisons) {
  test <- lapply(comparisons, function(comp) {
    replicate_comparison(comp, data, dataset_name = name)
  })

  pdf_filename <- paste0(out_dir,"Comparisons_", name, ".pdf")
  pdf(pdf_filename, width = 4, height = 4)

  for (res in test) {
    if (!is.null(res)) {
      print(res$freq)
      print(res$size)
      print(res$exp)
    }
  }

  dev.off()
  return(test)
}



```

```{r}
results <- Map(function(data, name) {
  message(paste("Processing dataset:", name))
  process_and_save_plots(data, name, comparisons)
},
data_list,
names(data_list))


for (dname in names(data_list)) {
  cat("## Dataset:", dname, "\n\n")
  
  for (cname in names(comparisons)) {
    cat("### Comparison:", cname, "\n\n")
    
    plots <- replicate_comparison(
      comparisons[[cname]],
      data_list[[dname]],
      dataset_name = dname
    )
    
    if (!is.null(plots)) {
      print(plots$freq)
      print(plots$size)
      print(plots$exp)
    }
  }
}

```



```{r, echo = FALSE}

make_correlation_heatmap <- function(dt, feature, dataset_name, sample_reps = NULL) {
  dt <- as.data.table(dt)
  
  if (!"sample" %in% names(dt)) stop("No 'sample' column found in data")
  
  # Just pull Seq_ID, sample, and the numerical feature
  wide_dt <- dt[, .(
    seq_ID,
    sample_label = sample, 
    value = as.numeric(get(feature))
  )]
  
  # FILTER STEP: Keep only rows where 'sample_label' is in sample_reps
  if (!is.null(sample_reps)) {
    wide_dt <- wide_dt[sample_label %in% sample_reps]
  }
  
  # Check if we still have data after filtering
  if (nrow(wide_dt) == 0) {
    warning(paste("No data left after filtering for sample_reps in dataset:", dataset_name))
    return(NULL)
  }
  
  # Reshape to wide format using the sample names
  wide_dt <- dcast(wide_dt, seq_ID ~ sample_label, value.var = "value")
  
  mat <- as.matrix(wide_dt[, -1, with = FALSE])
  storage.mode(mat) <- "numeric"
  mat[mat <= 0] <- NA
  mat <- log10(mat)
  
  mat <- mat[, colSums(is.finite(mat)) >= 2, drop = FALSE]
  if (ncol(mat) < 2) return(NULL)
  
  cormat <- cor(mat, use = "pairwise.complete.obs", method = "pearson")
  cormat_tbl <- as.data.frame(as.table(cormat))
  colnames(cormat_tbl) <- c("Sample1", "Sample2", "Correlation")
  
  ggplot(cormat_tbl, aes(Sample1, Sample2, fill = Correlation)) +
    geom_tile(color = "white") +
    scale_fill_gradient2(
      low = "#4575b4", mid = "white", high = "#d73027",
      midpoint = 0.5, limits = c(0, 1), na.value = "grey90"
    ) +
    geom_text(aes(label = ifelse(is.na(Correlation), "NA", sprintf("%.2f", Correlation))), size = 3) +
    labs(title = paste0("Dataset: ", dataset_name, " | Feature: ", feature), x = "", y = "") +
    theme_minimal(base_size = 10) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          panel.grid = element_blank())
}

```

```{r}
features <- c("exp", "bfreq", "bsize")
all_heatmaps <- list()

# Define your vector of samples to include
# e.g., sample_reps <- c("libA_exp1_rep1", "libB_exp1_rep2", ...) 
# Make sure this matches the formatting paste(sample, experiment, rep, sep = "_")

for (dname in names(data_list)) {
  cat("Processing dataset:", dname, "\n")
  dt <- data_list[[dname]]
  
  for (f in features) {
    # Generate the plot, passing the sample_reps vector
    p <- make_correlation_heatmap(dt, f, dname, sample_reps = sample_reps)
    
    if (!is.null(p)) {
      heatmap_name <- paste(dname, f, sep = "_")
      all_heatmaps[[heatmap_name]] <- p
      
      # Save the plot automatically
      file_path <- paste0(out_dir, "Heatmap_", heatmap_name, ".pdf")
      ggsave(filename = file_path, plot = p, width = 5, height = 5, units = "in")
    }
  }
}

# Print them locally/to the console 
for (nm in names(all_heatmaps)) {
  print(all_heatmaps[[nm]])
}

```


##looking at examples 

```{r, echo = FALSE}
diff_exp_sub2 %>% filter() %>% 
  ggplot() + 
  geom_histogram(aes(x = log2FoldChange_RT2),bins = 50) + 
  theme_classic(base_size = 15)

STAP_annotated <- STAP_annotated %>% 
  mutate(log2_class = ifelse(between(log2FoldChange, 5, 7.5), "low", 
                      ifelse(between(log2FoldChange, 7.5, 10), "mid", 
                      ifelse(between(log2FoldChange, 10, 12.5), "high", "too_low"))))

#similar exp
oligo_low <- unique(STAP_annotated[which(STAP_annotated$log2_class == "low" & STAP_annotated$condition == "RT2" & STAP_annotated$filter1 == TRUE & STAP_annotated$filter2 == TRUE & STAP_annotated$full_prec2 == TRUE),]$seq_ID)
oligo_mid <- unique(STAP_annotated[which(STAP_annotated$log2_class == "mid" & STAP_annotated$condition == "RT2" & STAP_annotated$filter1 == TRUE & STAP_annotated$filter2 == TRUE & STAP_annotated$full_prec2 == TRUE),]$seq_ID)
oligo_high <- unique(STAP_annotated[which(STAP_annotated$log2_class == "high" & STAP_annotated$condition == "RT2" & STAP_annotated$filter1 == TRUE & STAP_annotated$filter2 == TRUE & STAP_annotated$full_prec2 == TRUE),]$seq_ID)

#diff_exp_long <- merge(diff_exp_long, new_STAP, by.x = "seq_ID", by.y = "oligo_id", all.x = TRUE)

#ftz chr3R_2689979_2690112_+ (in data set!)
#gt chrX_2322929_2323062_-

oligo_TATA <- c("chr2R_8592612_8592745_+", "chr3L_21375536_21375669_+", "chr3R_14061849_14061982_+", "chr3R_9916429_9916562_+")
#chr2R_8592612_8592745_+ = CG42663
# chr3L_21375536_21375669_+ = rgn 
# chr3R_14061849_14061982_+ = repo
# chr3R_9916429_9916562_+ = FBgn0038198 Npc2b

#more examples from high 
#Inr DPE
Inr_DPE <- c("chrX_16615373_16615506_-", "chr3R_9809016_9809149_-", "chr2R_8592612_8592745_+")

#DPE
DPE <- c("chrX_18428631_18428764_+", "chrX_16615373_16615506_-", "chr3R_9809016_9809149_-")

#Inr Ebox
c("chrX_1930426_1930559_-")

#Inr
c("chrX_13298723_13298856_+")
#TATAInr
c("chr3R_9916429_9916562_+")

#TATA examples from high 
TATA <- c("chrX_9556722_9556855_-", "chrX_6173234_6173367_-", "chrX_14970176_14970309_-", "chr3R_9916429_9916562_+")

FISH <- c("chr2L_15762688_15762821_-", "chr2L_1737402_1737535_+", "chr2L_21629861_21629994_-", "chr2R_8592612_8592745_+", "chr3L_21375536_21375669_+", "chr3R_14061849_14061982_+", "Dscp_promoter_synthetic_+")

#write.table(STAP_annotated,"/groups/stark/lorbeer/Burst_params_DG/STAP_March2025_pools/STAP_data_annotated_log2FC.txt", sep = "\t", row.names = FALSE, quote = FALSE)

```

## Reyna fits

```{r, echo = FALSE, fig.width = 12,warning = FALSE, message = FALSE}

#Given two sequences, each one with:
#mu (expression)
#r parameter
#UMI counts
#Which ca be respectively annotated into the following objects:
#mu_n1, r_n1, x1 (for sequence 1)
#mu_n2, r_n2, x2 (for sequence 2)

prom1 <- "chr2R_8592612_8592745_+"
prom2 <- TATA[1]

#Function for computing the densities of the 0 truncate negative binomial:
d0T_NB <- function(x, mu, size){
  dnbinom(x, mu=mu, size=size)/(1-dnbinom(0, mu=mu, size=size))
}

x1 <- count_data %>%
  filter(sample == "libFL008_E28_RT2_r2" & seq_ID == prom1) %>%
  pull(N)
x2 <- count_data %>%
  filter(sample == "libFL008_E28_RT2_r2" & seq_ID == prom2) %>%
  pull(N)


mu_n1 <- STAP_annotated %>% filter(sample == "libFL008_E28_RT2_rep2" & seq_ID == prom1) %>% pull(exp)
r_n1 <- STAP_annotated %>% filter(sample == "libFL008_E28_RT2_rep2" & seq_ID == prom1) %>% pull(bfreq)


mu_n2 <- STAP_annotated %>% filter(sample == "libFL008_E28_RT2_rep2" & seq_ID == prom2) %>% pull(exp)
r_n2 <- STAP_annotated %>% filter(sample == "libFL008_E28_RT2_rep2" & seq_ID == prom2) %>% pull(bfreq)


xr <- 1:max(c(x1, x2))

d1 <- data.frame(
  x = xr,
  density = d0T_NB(xr, mu = mu_n1, size = r_n1),
  sample = prom1
)

d2 <- data.frame(
  x = xr,
  density = d0T_NB(xr, mu = mu_n2, size = r_n2),
  sample = prom2
)

densities <- rbind(d1, d2)

# build a tidy frame 
obs <- bind_rows(
  data.frame(N = x1, sample = prom1),
  data.frame(N = x2, sample = prom2)
)

p <- ggplot(obs, aes(x = N, fill = sample)) +
  geom_histogram(aes(y = after_stat(density)), 
                 position = "identity", alpha = 0.4, binwidth = 1) +
  geom_line(data = densities, aes(x = x, y = density, colour = sample)) +
  scale_fill_manual(values = c("goldenrod2", "darkcyan")) +
  scale_colour_manual(values = c("goldenrod2", "darkcyan")) +
    xlim(0,50) + 
  ylim(0,0.35) + 
  labs(x = "transcript/plasmid", y = "Relative frequency",
       title = "Observed vs. 0-truncated NB fit in rep2") +
  theme_classic(base_size = 15)

x3 <- count_data %>%
  filter(sample == "libFL008_E28_RT2_r3" & seq_ID == prom1) %>%
  pull(N)
x4 <- count_data %>%
  filter(sample == "libFL008_E28_RT2_r3" & seq_ID == prom2) %>%
  pull(N)

mu_n3 <- STAP_annotated %>% filter(sample == "libFL008_E28_RT2_rep3" & seq_ID == prom1) %>% pull(exp)
r_n3 <- STAP_annotated %>% filter(sample == "libFL008_E28_RT2_rep3" & seq_ID == prom1) %>% pull(bfreq)

mu_n4 <- STAP_annotated %>% filter(sample == "libFL008_E28_RT2_rep3" & seq_ID == prom2) %>% pull(exp)
r_n4 <- STAP_annotated %>% filter(sample == "libFL008_E28_RT2_rep3" & seq_ID == prom2) %>% pull(bfreq)

xr2 <- 1:max(c(x1, x2))

d3 <- data.frame(
  x = xr,
  density = d0T_NB(xr, mu = mu_n3, size = r_n3),
  sample = prom1
)

d4 <- data.frame(
  x = xr,
  density = d0T_NB(xr, mu = mu_n4, size = r_n4),
  sample = prom2
)

densities2 <- rbind(d3, d4)

# build a tidy frame 
obs2 <- bind_rows(
  data.frame(N = x3, sample = prom1),
  data.frame(N = x4, sample = prom2)
)
q<- ggplot(obs2, aes(x = N, fill = sample)) +
  geom_histogram(aes(y = after_stat(density)), 
                 position = "identity", alpha = 0.4, binwidth = 1) +
  geom_line(data = densities2, aes(x = x, y = density, colour = sample)) +
  scale_fill_manual(values = c("goldenrod2", "darkcyan")) +
  scale_colour_manual(values = c("goldenrod2", "darkcyan")) +
  labs(x = "transcript/plasmid", y = "Relative frequency",
       title = "Observed vs. 0-truncated NB fit in rep3") +
  xlim(0,50) + 
  ylim(0,0.35) + 
  theme_classic(base_size = 15)

grid.arrange(p, q)


#comparison of same 
d5 <- data.frame(
  x = xr,
  density = d0T_NB(xr, mu = mu_n1, size = r_n1),
  sample = "rep2"
)

d6 <- data.frame(
  x = xr,
  density = d0T_NB(xr, mu = mu_n2, size = r_n2),
  sample = "rep2"
)
d7 <- data.frame(
  x = xr,
  density = d0T_NB(xr, mu = mu_n3, size = r_n3),
  sample = "rep3"
)

d8 <- data.frame(
  x = xr,
  density = d0T_NB(xr, mu = mu_n4, size = r_n4),
  sample = "rep3"
)

densities3 <- rbind(d5, d7)
densities4 <- rbind(d6, d8)

# build a tidy frame 
obs3 <- bind_rows(
  data.frame(N = x1, sample = "rep2"),
  data.frame(N = x3, sample = "rep3")
)

obs4 <- bind_rows(
  data.frame(N = x2, sample = "rep2"),
  data.frame(N = x4, sample = "rep3")
)

p1 <- ggplot(obs3, aes(x = N, fill = sample)) +
  geom_histogram(aes(y = after_stat(density)), 
                 position = "identity", alpha = 0.4, binwidth = 1) +
  geom_line(data = densities3, aes(x = x, y = density, colour = sample)) +
  scale_fill_manual(values = c("goldenrod2", "goldenrod4")) +
  scale_colour_manual(values = c("goldenrod2", "goldenrod4")) +
    xlim(0,50) + 
  ylim(0,0.35) + 
  labs(x = "transcript/plasmid", y = "Relative frequency",
       title = "chr2R_8592612_8592745_+ DPE") +
  theme_classic(base_size = 15)

q1 <- ggplot(obs4, aes(x = N, fill = sample)) +
  geom_histogram(aes(y = after_stat(density)), 
                 position = "identity", alpha = 0.4, binwidth = 1) +
  geom_line(data = densities4, aes(x = x, y = density, colour = sample)) +
  scale_fill_manual(values = c("cyan2", "darkcyan")) +
  scale_colour_manual(values = c("cyan2", "darkcyan")) +
  labs(x = "transcript/plasmid", y = "Relative frequency",
       title = "chrX_9556722_9556855_- TATA") +
  xlim(0,50) + 
  ylim(0,0.35) + 
  theme_classic(base_size = 15)

grid.arrange(p,p1, q, q1, layout_matrix = cbind(c(1,3), c(1,3), c(2,4)))

# Save to PDF
# ggsave(
#   filename = paste0(out_dir, "Histogram_comparison_TATA-noTATA_E28_v1.pdf"),
#   plot = grid.arrange(p,p1, q, q1, layout_matrix = cbind(c(1,3), c(1,3), c(2,4))),
#   width = 12, height = 8, units = "in"  # adjust size as needed
# )

```
#make the plots for more sequences and more samples
```{r, echo = FALSE}

plot_nb_fit_facets <- function(count_data, stap_data, seq_ids, samples, 
                              sample_labels = NULL) {
  
  # Default labels if not provided
  if (is.null(sample_labels)) {
    sample_labels <- samples
  }
  
  # Check inputs
  if (length(sample_labels) != length(samples)) {
    stop("sample_labels must have same length as samples")
  }
  
  # Function for computing the densities of the 0 truncate negative binomial
  d0T_NB <- function(x, mu, size){
    dnbinom(x, mu=mu, size=size)/(1-dnbinom(0, mu=mu, size=size))
  }
  
  # Storage for all plots
  plot_list <- list()
  
  for (seq_id in seq_ids) {
    
    # Storage for this sequence
    all_obs <- list()
    all_densities <- list()
    all_params <- list()
    
    # Process each sample
    for (i in seq_along(samples)) {
      
      # Extract UMI counts
      x <- count_data %>%
        filter(sample == samples[i] & seq_ID == seq_id) %>%
        pull(N)
      
      # Skip if no data
      if (length(x) == 0) {
        warning(paste("No data found for", seq_id, "in sample", samples[i]))
        next
      }
      
      # Extract parameters from STAP_annotated
      # Convert sample name format (r2 -> rep2, etc.)
      stap_sample <- gsub("_r(\\d+)$", "_rep\\1", samples[i])
      
      mu_n <- stap_data %>% 
        filter(sample == stap_sample & seq_ID == seq_id) %>% 
        pull(exp)
      r_n <- stap_data %>% 
        filter(sample == stap_sample & seq_ID == seq_id) %>% 
        pull(bfreq)
      
      # Skip if parameters not found
      if (length(mu_n) == 0 || length(r_n) == 0) {
        warning(paste("Parameters not found for", seq_id, "in sample", samples[i]))
        next
      }
      
      # Store observed data
      all_obs[[i]] <- data.frame(
        N = x, 
        sample = sample_labels[i],
        sample_order = i  # for consistent ordering
      )
      
      # Create density curve
      xr <- 1:max(x)
      all_densities[[i]] <- data.frame(
        x = xr,
        density = d0T_NB(xr, mu = mu_n, size = r_n),
        sample = sample_labels[i],
        sample_order = i
      )
      
      # Store parameters for display
      all_params[[i]] <- data.frame(
        sample = sample_labels[i],
        sample_order = i,
        mu = round(mu_n, 3),
        r = round(r_n, 3),
        param_label = paste0("μ = ", round(mu_n, 3), "\nr = ", round(r_n, 3))
      )
    }
    
    # Skip if no valid data found
    if (length(all_obs) == 0) {
      warning(paste("No valid data found for", seq_id))
      next
    }
    
    # Combine all data
    obs <- bind_rows(all_obs)
    densities <- bind_rows(all_densities)
    params <- bind_rows(all_params)
    
    # Determine global x-axis range for consistent scaling
    x_max <- max(obs$N)
    
    # Extend density curves to common range if needed
    densities_extended <- list()
    for (i in seq_along(all_densities)) {
      if (nrow(all_densities[[i]]) > 0) {
        current_max <- max(all_densities[[i]]$x)
        if (current_max < x_max) {
          # Extend with zeros
          extension <- data.frame(
            x = (current_max + 1):x_max,
            density = 0,
            sample = all_densities[[i]]$sample[1],
            sample_order = all_densities[[i]]$sample_order[1]
          )
          densities_extended[[i]] <- rbind(all_densities[[i]], extension)
        } else {
          densities_extended[[i]] <- all_densities[[i]]
        }
      }
    }
    densities <- bind_rows(densities_extended)
    
    # Create colors (cycle through if more samples than colors)
    colors <- c('#1f77b4')
    sample_colors <- colors[((1:length(unique(obs$sample))) - 1) %% length(colors) + 1]
    names(sample_colors) <- unique(obs$sample[order(obs$sample_order)])
    
    # Create plot with facets
    p <- ggplot(obs, aes(x = N)) +
      geom_histogram(aes(y = after_stat(density), fill = sample), 
                     alpha = 0.4, binwidth = 1, #fill = "#1f77b4"
                     ) +
      geom_line(data = densities, aes(x = x, y = density, color = sample)) +
      geom_text(data = params, aes(label = param_label), 
                x = Inf, y = Inf, hjust = 1.1, vjust = 1.1, size = 3) +
      xlim(0,25) + 
      ylim(0,0.4) + 
      facet_wrap(~ factor(sample, levels = unique(sample[order(sample_order)])), 
                 scales = "fixed") +
      scale_fill_manual(values = sample_colors) +
      scale_color_manual(values = sample_colors) +
      labs(x = "Transcript/plasmid", y = "Relative frequency",
           title = paste("Observed vs. 0-truncated NB fit:", seq_id),
           fill = "Sample", color = "Sample") +
      theme_classic(base_size = 12) +
      theme(strip.background = element_rect(fill = "lightgray"))
    
    plot_list[[seq_id]] <- p
  }
  
  # Return single plot if only one seq_ID, otherwise return list
  if (length(plot_list) == 1) {
    return(plot_list[[1]])
  } else {
    return(plot_list)
  }
}




```


```{r}
plot_nb_fit_facets(count_data, STAP_annotated, 
                   seq_ids = "Dscp_promoter_synthetic_+",
                   samples = c("libFL008_E28_RT2_r2", "libFL008_E28_RT2_r3", "libFL008_E28_RT2_pooledby2"),
                   sample_labels = c( "rep2", "rep3", "pooledby2"))

```


```{r}
plots <- plot_nb_fit_facets(count_data, STAP_annotated, 
                           seq_ids = TATA,
                           samples = c("libFL008_E28_RT2_pooledby2"),
                           sample_labels = c("pooledby2"))

# Display individual plots
plots
#plots[["chr2R_8592612_8592745_+"]]

```




```{r}

plot_nb_fit_facets_motif(count_data, STAP_annotated,
                   seq_ids=c("Dscp_promoter_synthetic_+",oligo_TATA),
                   samples=c("libFL008_E28_RT2_r1","libFL008_E28_RT2_r2","libFL008_E28_RT2_r3"),
                   sample_labels=c("rep1","rep2","rep3"),
                   color_by="sample")


plot_nb_fit_facets_motif(count_data, STAP_annotated,
                   seq_ids=c("Dscp_promoter_synthetic_+",oligo_TATA),
                   samples=c("libFL008_E28_RT2_r1","libFL008_E28_RT2_r2","libFL008_E28_RT2_r3"),
                   color_by="TATAbox")

plot_nb_fit_facets_motif(count_data, STAP_annotated,
                   seq_ids=c("Dscp_promoter_synthetic_+",FISH),
                   samples=c("libFL008_E28_RT2_r1","libFL008_E28_RT2_r2","libFL008_E28_RT2_r3"),
                   color_by="TATAbox")



```



```{r, echo = FALSE}
# count_data <- merge(count_data, new_STAP, by.x = "seq_ID", by.y = "oligo_id", all.x = TRUE)
# 
# count_data %>% filter(enhancer == "RT2" & rep %in% c("rep2", "rep3")) %>%
#   filter(seq_ID %in% oligo_low[c(6:7,9,14, 24)]) %>%
#   ggplot() +
#   geom_histogram(aes(x = N), binwidth = 1,position = "identity", fill = "lightgrey",
#                  alpha = 0.9) +
#   #scale_color_manual(values = c("#3B9AB2", "#03596D")) +
#   #scale_color_manual(values = c("darkgrey", "black")) +
#   xlim(0,55)+
#   #ylim(0,100) +
#   facet_grid(rep~ seq_ID) +
#   theme_classic(base_size = 15)
# oligo_low[1:7]

count_data %>% filter(enhancer == "RT2" & rep %in% c("rep2", "rep3")) %>%
  filter(seq_ID %in% oligo_mid[c(11,14,17)]) %>%
  ggplot() +
  geom_histogram(aes(x = N), binwidth = 1,position = "identity", fill = "lightgrey",
                 alpha = 0.9) +
  #scale_color_manual(values = c("#3B9AB2", "#03596D")) +
  #scale_color_manual(values = c("darkgrey", "black")) +
  xlim(0,55)+
  #ylim(0,100) +
    labs(title = "Histograms of mid expression examples")+ 
  facet_grid(rep~ seq_ID) +
  theme_classic(base_size = 15)
oligo_mid[1:7]

count_data %>% filter(enhancer == "RT2" & rep %in% c("rep2", "rep3")) %>%
  filter(seq_ID %in% oligo_high[c(3,5,6, 10)]) %>%
  ggplot() +
  geom_histogram(aes(x = N), binwidth = 1,position = "identity", fill = "lightgrey",
                 alpha = 0.9) +
  #scale_color_manual(values = c("#3B9AB2", "#03596D")) +
  #scale_color_manual(values = c("darkgrey", "black")) +
  xlim(0,55)+
  #ylim(0,100) +
    labs(title = "Histograms of high expression examples")+ 
  facet_grid(rep~ seq_ID) +
  theme_classic(base_size = 15)
oligo_high[1:7]
# 



low <- c(#"chr2L_10993783_10993916_-" "chr2L_11285706_11285839_+" "chr2L_12370533_12370666_-" 
  "chr2L_13060071_13060204_-","chr2L_13375384_13375517_+", "chr2L_16816598_16816731_+", "chr2L_21656707_21656840_-", "chr2L_7402839_7402972_+")
mid <- c(#"chr2L_10199045_10199178_-" "chr2L_10770801_10770934_+" "chr2L_10801041_10801174_-" 
  #"chr2L_11081330_11081463_-", #"chr2L_11136335_11136468_+" 
  #"chr2L_11149117_11149250_-", "chr2L_1163815_1163948_-", 
  "chr2L_13550076_13550209_+", "chr2L_13847859_13847992_+", #"chr2L_13971350_13971483_+", 
  "chr2L_14417280_14417413_-")
high <- c(#"Dscp_promoter_synthetic_+", 
  #"chr2L_14001181_14001314_-",  
  "chr2L_15762688_15762821_-",#"chr2L_15893748_15893881_-", 
  "chr2L_1737402_1737535_+",
  "chr2L_18987209_18987342_-"#, "chr2L_20115909_20116042_-"
  ,"chr2L_21629861_21629994_-"
)
low_new <- STAP_annotated %>% filter(condition == "RT2" & full_prec2 == TRUE & filter2 == TRUE & high_mu == TRUE & rep == "pooledby2" & log2_class == "low") %>% pull(seq_ID)

low <- c(#"chr2L_10993783_10993916_-" "chr2L_11285706_11285839_+" "chr2L_12370533_12370666_-" 
  "chr2L_13060071_13060204_-","chr2L_16816598_16816731_+","chr2L_7497293_7497426_-" )

count_data %>% filter(enhancer == "RT2" & rep %in% c("rep2", "rep3")) %>%
  filter(seq_ID %in% low) %>%
  ggplot() +
  geom_histogram(aes(x = N), binwidth = 1,position = "identity", fill = "lightgrey",
                 alpha = 0.9) +
  labs(title = "Histograms of low expression examples")+ 
  #scale_color_manual(values = c("#3B9AB2", "#03596D")) +
  #scale_color_manual(values = c("darkgrey", "black")) +
  xlim(0,55)+
  #ylim(0,100) +
  facet_grid(rep~ seq_ID) +
  theme_classic(base_size = 15)

```


```{r, echo = FALSE}

plot_nb_fit_facets(count_data, STAP_annotated, 
                   seq_ids = low,
                   samples = c("libFL008_E28_RT2_pooledby2"),
                   sample_labels = c("pooledby2"))

plot_nb_fit_facets(count_data, STAP_annotated, 
                   seq_ids = mid,
                   samples = c("libFL008_E28_RT2_pooledby2"),
                   sample_labels = c("pooledby2"))

plot_nb_fit_facets(count_data, STAP_annotated, 
                   seq_ids = c(high, TATA[1]),
                   samples = c("libFL008_E28_RT2_pooledby2"),
                   sample_labels = c("pooledby2"))

plot_nb_fit_facets(count_data, STAP_annotated, 
                   seq_ids = c(FISH),
                   samples = c("libFL008_E28_RT2_pooledby2"),
                   sample_labels = c("pooledby2"))

```


```{r}
plots2 <- plot_nb_fit_facets(count_data, STAP_annotated, 
                           seq_ids = c(low,mid,high),
                           samples = c("libFL008_E28_RT2_r2", 
                                     "libFL008_E28_RT2_r3"),
                           sample_labels = c("rep2", "rep3"))

# Display individual plots
plots2

plots2 <- plot_nb_fit_facets(
  count_data, STAP_annotated, 
  seq_ids = c(low, mid, high, TATA[1], Inr_DPE[3], FISH),
  samples = c("libFL008_E28_RT2_pooledby2"),
  sample_labels = c("pool")
)

#print plots
if (is.list(plots2)) {
  for (i in seq_along(plots2)) {
    p <- plots2[[i]]
    if (!is.null(p) && inherits(p, "ggplot")) {
      nm <- names(plots2)[i]
      fname <- if (!is.null(nm) && nzchar(nm)) {
        paste0("nb_fit_facets_pool_ylim", nm, ".pdf")
      } else {
        paste0("nb_fit_facets_pool_ylim", i, ".pdf")
      }
      grDevices::cairo_pdf(paste0(out_dir,fname), width = 3, height = 2)
      print(p)
      dev.off()
    }
  }
}
head(unique(STAP_annotated[seq_ID %in% c(low, mid, high, TATA[1], Inr_DPE[3]),.(seq_ID, main_mots)]), 11)
```


```{r, echo = FALSE}

fitlm = lm((log2FoldChange_RT2)~(log2FoldChange_zfh1) , data = diff_exp_sub2)
PCC <- cor((diff_exp_sub2$log2FoldChange_zfh1), diff_exp_sub2$log2FoldChange_RT2, method = 'pearson', use = "complete.obs")

library(ggrepel)

p_scatter <- diff_exp_sub2 %>%
  mutate(selected_oligos = ifelse(oligo_id %in% low, "low", 
                          ifelse(oligo_id %in% mid, "mid", 
                          ifelse(oligo_id %in% c(high, TATA[1], Inr_DPE[3]), "high", "not_selected")))) %>%
  ggplot() + 
  # base layer: faint background
  geom_point(aes(x = log2FoldChange_zfh1, y = log2FoldChange_RT2, color = selected_oligos), 
             alpha = 0.3, size = 2) + 
  
  # overlay: highlighted low/mid/high
  geom_point(data = . %>% filter(selected_oligos %in% c("low","mid","high")),
             aes(x = log2FoldChange_zfh1, y = log2FoldChange_RT2, color = selected_oligos),
             alpha = 1, size = 4) +
  
  # labels: oligo IDs for highlighted only
  geom_text_repel(data = . %>% filter(selected_oligos %in% c("low","mid","high")),
                  aes(x = log2FoldChange_zfh1, 
                      y = log2FoldChange_RT2, 
                      label = oligo_id,
                      color = selected_oligos),
                  size = 4, show.legend = FALSE) +
  
  geom_abline(aes(intercept = fitlm$coefficients[1], slope = fitlm$coefficients[2]), 
              linetype = "dashed") + 
  scale_color_manual(values = c("low"='#00a9da', "mid"='#1f77b4', "high"='#023e64', "not_selected"='grey32')) + 
  ggtitle("BARe-seq library activity vs. STAP-seq E28") + 
  labs(title = paste0("PCC = ", round(PCC, digits = 2)), 
       subtitle = paste0("R^2 = ", round(summary(fitlm)$r.squared, digits = 2))) + 
  ylab("log2FC BARe-seq promoter activity") + 
  xlab("log2FC STAP-seq promoter activity") + 
  xlim(-3, 15) + 
  ylim(-3, 15) + 
  coord_fixed() + 
  theme_classic(base_size = 15)

# Display in the Rmd output
p_scatter

# Save to PDF (adjust size as needed)
ggsave("BAre_vs_STAP_labeled_oligos.pdf", plot = p_scatter, width = 7, height = 7, units = "in")


```





#how do the parameters look? 
```{r}


p <- STAP_annotated %>% filter(condition == "RT2" & high_mu == TRUE & KS_pval == TRUE) %>%
  ggplot() + 
  geom_histogram(aes(x = log10(bfreq)), binwidth = 0.1,  fill = "#F2994A", alpha = 0.8) + 
      labs(title = "high_mu and KS_pval")+ 
  xlim(-2,1.5) + 
  theme_classic(base_size = 15)

q <- STAP_annotated %>% filter(condition == "RT2" & high_mu == TRUE & KS_pval == TRUE) %>%
  ggplot() + 
  geom_histogram(aes(x = log10(bsize)), binwidth = 0.1, fill = "#F2994A", alpha = 0.8) + 
  xlim(-2,1.5) + 
  theme_classic(base_size = 15)
q1 <- STAP_annotated %>% filter(condition == "RT2" & high_mu == TRUE & KS_pval == TRUE) %>%
  ggplot() + 
  geom_histogram(aes(x = log10(exp)), binwidth = 0.1, fill = "#F2994A", alpha = 0.8) + 
  xlim(-2,1.5) + 
  theme_classic(base_size = 15)

grid.arrange(p,q,q1, ncol = 1)

ggsave(paste0(out_dir,"BAre_E28_parameter_histograms.pdf"), plot = grid.arrange(p,q,q1, ncol = 1), width = 3, height = 8, units = "in")

```

```{r}

#plotting burst size vs burst frequency
scatter_df <- STAP_annotated %>% 
  filter(condition == "RT2", full_prec2 == TRUE, filter2 == TRUE, high_mu == TRUE, rep == "pooledby2") %>%
  mutate(
    selected_oligos = ifelse(
      seq_ID %in% low, "low",
      ifelse(
        seq_ID %in% mid, "mid",
        ifelse(seq_ID %in% c(high, TATA[1], Inr_DPE[3]), "high", "not_selected")
      )
    )
  )

scatter2 <- ggplot(scatter_df) +
  # base/background points
  geom_point(aes(x = log10(bsize), y = log10(bfreq)),
             color = "gray12", alpha = 0.4, size = 2) +
  
  # overlay highlighted: low/mid/high
  geom_point(
    data = subset(scatter_df, selected_oligos %in% c("low","mid","high")),
    aes(x = log10(bsize), y = log10(bfreq), color = selected_oligos),
    alpha = 1, size = 4
  ) +
  
  # labels for selected oligos
  geom_text_repel(
    data = subset(scatter_df, selected_oligos %in% c("low","mid","high")),
    aes(x = log10(bsize), y = log10(bfreq), label = seq_ID, color = selected_oligos),
    size = 4, show.legend = FALSE
  ) +
  
  scale_color_manual(values = c(
    low = '#00a9da', mid = '#1f77b4', high = '#023e64', not_selected = 'grey32'
  )) +
  xlim(-2.5, 2.5) +
  theme_classic(base_size = 15)

scatter2
ggsave("bsize_bfreq_higjlighted.pdf", plot = scatter2, width = 7, height = 7, units = "in")

```



