---
title: "Figure 2 & Supplemental Figure 3 — promoter core-motif analysis"
author: "FL"
date: "`r Sys.Date()`"
output: html_document
---

<!--
================================================================================
FIGURE / PANEL MAP  (which chunk makes which panel)
================================================================================
This script reproduces the promoter core-promoter-motif analyses.

FIGURE 2
  2A  Illustration of the library + motif set .............. (no code)
  2B  TATA box vs DPE vs other: burst size / frequency /
      mean expression boxplots with ANOVA ................... chunk "FIGURE 2B"
  2C  Motif PWM score vs burst size / frequency
      (TATA box, DPE) ....................................... chunk "FIGURE 2C"
                                                              *** needs external
                                                              files, see below ***
  2D  Regression-coefficient heatmap, pooled samples,
      multivariate model .................................... chunk "FIGURE 2D"
  2E  Meta-analysis: pooled effect on burst size vs
      burst frequency ...................................... chunk "FIGURE 2E"

SUPPLEMENTAL FIGURE 3
  S3A Motif-frequency donut + motif-combination UpSet ....... chunk "SUPPL. FIG. 3A"
  S3B TATA box boxplots split by expression bin ............. chunk "SUPPL. FIG. 3B"
  S3C log2FC (DSCP) histogram ............................... not in this script
  S3D Luciferase validation ................................. (no code)
  S3E BARe-seq vs STAP-seq scatter (CG4822, Pld) ............ see Fig. 1 script
  S3F Replicate correlation heatmaps ....................... see Fig. 1 script
  S3G Regression-coefficient heatmap, all samples ........... chunk "SUPPL. FIG. 3G"

  Additional exploratory chunks (univariate heatmaps, median-difference /
  log2FC heatmaps, and per-motif expression-bin boxplots for DPE / INR /
  E-box / Ohler6 / Ohler1) are kept and clearly labelled "EXPLORATORY".

================================================================================
MISSING / EXTERNAL INPUTS  --  please decide how to handle these
================================================================================
Everything EXCEPT Figure 2C is driven by the published supplemental tables
(Table4, Table5) that are read in the data-loading chunk below.

Figure 2C (motif PWM score vs burst parameters) additionally needs raw inputs
that are NOT part of the supplemental tables and were NOT provided. The chunk
that produces Fig. 2C is therefore set to `eval = FALSE`; supply these files,
set it to `eval = TRUE`, and verify before use:

  1. CORE-PROMOTER MOTIF PWM DATABASE
       original path : /groups/stark/vloubiere/motifs_db/CP_motifs_PWM.rds
       expected here : data/CP_motifs_PWM.rds
       R object      : list with element $Pwms_perc (a TFBSTools PWM list)

  2. PROMOTER / OLIGO SEQUENCES (with linker)
       original path : /groups/stark/lorbeer/STAP-TWIST/
                       Franzi_STAP_STARR_TWIST_seqonly-v2.tsv
       expected here : data/Franzi_STAP_STARR_TWIST_seqonly-v2.tsv
       columns used  : ID_TWIST, Sequence_with_linker

  3. PRIVATE R PACKAGE "vlite" (V. Loubiere)
       original call : devtools::load_all("/groups/stark/vloubiere/vlite/")
       Only used for the `vl_par` plotting helper in a commented-out diagnostic
       block. Not required for the panels; the active code uses TFBSTools /
       Biostrings / motifmatchr instead.

  4. MISSING CODE STEP (not a file): the original script is missing the
     motif-scanning call that defines `motif_hits`. A reconstructed
     motifmatchr::matchMotifs(...) call is provided in the Fig. 2C chunk and is
     clearly flagged --- it MUST be checked against the original analysis.

All other paths were internal cluster paths and have been replaced with the
relative locations ./Tables (supplemental tables) and ./output/Fig2 (outputs).
================================================================================
-->

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE
)

require(data.table)
require(ggplot2)
require(broom)
require(purrr)
require(metafor)
library(dplyr)
library(tidyr)
require(readxl)
require(ggpubr)       # stat_compare_means
require(gridExtra)    # grid.arrange
require(patchwork)    # wrap_plots
require(scales)       # squish
require(ComplexUpset)
require(TFBSTools)
library(Biostrings)
```

## Load pre-processed data

Both inputs are published supplemental tables (expected in ./Tables):

* `Table5_promoter_burst_parameters.xlsx` (sheet "STAP"): per-promoter burst
  parameters (`bfreq`, `bsize`, `exp`, `log2FoldChange`) plus the pre-computed
  quality flags `high_mu` and `KS_pval`, for every sample.
* `Table4_promoter_library_annotations.xlsx` (sheet "STAP"): promoter
  annotations including the binary core-promoter-motif indicator columns
  (TATAbox, DPE, INR, Ebox, Ohler1, Ohler6, ...) and the STAP-seq reference.

```{r load_data}
out_dir <- "output/Fig2/"
dir.create(out_dir, recursive = TRUE, showWarnings = FALSE)

anno         <- read_xlsx("/groups/stark/lorbeer/BARE-seq_data_prep/Tables/Table4_promoter_library_annotations.xlsx", sheet = "STAP")
burst_params <- read_xlsx("/groups/stark/lorbeer/BARE-seq_data_prep/Tables/Table5_promoter_burst_parameters.xlsx",    sheet = "STAP")

# `param` = burst parameters + promoter annotations, keyed on seq_ID.
param <- as.data.table(burst_params)

# Split the publication sample name into its components:
#   lib_experiment_condition_rep   e.g.  libFL001 / STAP / zfh1 / pool
param[, c("lib", "experiment", "condition", "rep") :=
        tstrsplit(sample, "_", keep = c(1:4))]

# Add promoter annotations (motif columns, gene names, classification, ...).
param <- merge(param, as.data.table(anno), by = "seq_ID", all.x = TRUE)

# `high_mu` is stored as text in Table5; make it a proper logical. The quality
# filter used throughout is:  high_mu == TRUE  &  KS_pval > 0.01
param[, high_mu := (high_mu == "TRUE")]

# Publication sample names (grouped by condition, pool then two replicates).
pooled_samples <- c(
  "libFL001_STAP_zfh1_pool",
  "libFL002_STAP_CG4822_pool",
  "libFL003_STAP_Pld_pool"
)

sample_names <- c(
  "libFL001_STAP_zfh1_pool",   "libFL001_STAP_zfh1_replicate1",   "libFL001_STAP_zfh1_replicate2",
  "libFL002_STAP_CG4822_pool", "libFL002_STAP_CG4822_replicate1", "libFL002_STAP_CG4822_replicate2",
  "libFL003_STAP_Pld_pool",    "libFL003_STAP_Pld_replicate1",    "libFL003_STAP_Pld_replicate2"
)
```

## Create filtered per-sample list

Keep, for each sample, the promoters that pass the `high_mu` and `KS_pval`
quality thresholds.

```{r filtered_list}
param_list <- list()
for (s in sample_names) {
  param_list[[s]] <- as.data.frame(param[sample == s & (high_mu & KS_pval > 0.01)])
}
```

# Helper functions

1. `sig_stars`: significance labels.
2. `na_block_for_motifs`: placeholder result tables filled with NA when a model
   cannot be fit, to avoid downstream problems (needed only for non-pooled data).

```{r helpers}
sig_stars <- function(p) {
  case_when(
    is.na(p)  ~ NA_character_,
    p < 0.001 ~ "***",
    p < 0.01  ~ "**",
    p < 0.05  ~ "*",
    TRUE      ~ "ns"
  )
}

na_block_for_motifs <- function(motifs, outcome, model_label, n_used = NA_integer_) {
  tibble(
    Motif = motifs,
    Outcome = outcome,
    Model = model_label,
    n_used = n_used,
    n_present = NA_integer_,
    prop_present = NA_real_,
    estimate = NA_real_,
    std.error = NA_real_,
    statistic = NA_real_,
    p.value = NA_real_,
    FoldChange = NA_real_
  )
}
```

# Model fitting functions

Two functions implementing the core regression framework:

- univariate models, where each motif is tested separately, and
- multivariate models, where all selected motifs are included simultaneously.

Linear regression with `lm()`:  `log10(outcome) ~ motif1 + motif2 ... motifn`.
Both model types can be run with or without adjustment for expression via
`log10(exp)`. Output: estimated effect sizes, standard errors, p-values, and
fold-change values derived from the regression coefficients.

```{r model_functions}
fit_motif_model_one_df <- function(data, outcome, motifs, include_exp = FALSE) {
  model_label <- ifelse(include_exp, "motifs_plus_log10exp", "motifs_only")
  
  motifs_present <- intersect(motifs, colnames(data))
  motifs_missing <- setdiff(motifs, motifs_present)
  
  if (!outcome %in% colnames(data)) {
    return(na_block_for_motifs(motifs, outcome, model_label, n_used = NA_integer_))
  }
  if (include_exp && !"exp" %in% colnames(data)) {
    return(na_block_for_motifs(motifs, outcome, model_label, n_used = NA_integer_))
  }
  
  fitted_present <- purrr::map_dfr(motifs_present, function(m) {
    rhs <- m
    if (include_exp) rhs <- paste(rhs, "+ log10(exp)")
    fml <- as.formula(paste0("log10(", outcome, ") ~ ", rhs))
    
    mf <- tryCatch(model.frame(fml, data = data), error = function(e) NULL)
    n_used <- if (is.null(mf)) NA_integer_ else nrow(mf)
    
    if (is.null(mf) || n_used == 0) {
      return(
        na_block_for_motifs(m, outcome, model_label, n_used = ifelse(is.null(mf), NA_integer_, 0L)) %>%
          dplyr::mutate(
            n_present = ifelse(is.null(mf), NA_integer_, sum(mf[[m]] == 1, na.rm = TRUE)),
            prop_present = ifelse(is.null(mf), NA_real_, n_present / n_used)
          )
      )
    }
    
    n_present <- sum(mf[[m]] == 1, na.rm = TRUE)
    prop_present <- n_present / n_used
    
    model <- tryCatch(lm(fml, data = data), error = function(e) NULL)
    if (is.null(model)) {
      return(
        na_block_for_motifs(m, outcome, model_label, n_used = n_used) %>%
          dplyr::mutate(n_present = n_present, prop_present = prop_present)
      )
    }
    
    broom::tidy(model) %>%
      dplyr::filter(term == m) %>%
      dplyr::transmute(
        Motif = m,
        Outcome = outcome,
        Model = model_label,
        n_used = n_used,
        n_present = n_present,
        prop_present = prop_present,
        estimate = estimate,
        std.error = std.error,
        statistic = statistic,
        p.value = p.value,
        FoldChange = 10^estimate
      )
  })
  
  missing_part <- na_block_for_motifs(motifs_missing, outcome, model_label, n_used = NA_integer_) %>%
    dplyr::mutate(n_present = NA_integer_, prop_present = NA_real_)
  
  dplyr::bind_rows(fitted_present, missing_part) %>%
    dplyr::mutate(Motif = factor(Motif, levels = motifs)) %>%
    dplyr::arrange(Motif) %>%
    dplyr::mutate(Motif = as.character(Motif))
}

fit_motif_model_multiv_df <- function(data, outcome, motifs, include_exp = FALSE) {
  model_label <- ifelse(include_exp, "motifs_plus_log10exp", "motifs_only")
  
  motifs_present <- intersect(motifs, colnames(data))
  motifs_missing <- setdiff(motifs, motifs_present)
  
  if (!outcome %in% colnames(data)) {
    return(na_block_for_motifs(motifs, outcome, model_label, n_used = NA_integer_))
  }
  if (include_exp && !"exp" %in% colnames(data)) {
    return(na_block_for_motifs(motifs, outcome, model_label, n_used = NA_integer_))
  }
  if (length(motifs_present) == 0) {
    return(na_block_for_motifs(motifs, outcome, model_label, n_used = NA_integer_))
  }
  
  rhs <- paste(motifs_present, collapse = " + ")
  if (include_exp) rhs <- paste(rhs, "+ log10(exp)")
  fml <- as.formula(paste0("log10(", outcome, ") ~ ", rhs))
  
  mf <- tryCatch(model.frame(fml, data = data), error = function(e) NULL)
  n_used <- if (is.null(mf)) NA_integer_ else nrow(mf)
  
  if (is.null(mf) || n_used == 0) {
    out <- na_block_for_motifs(motifs_present, outcome, model_label, n_used = ifelse(is.null(mf), NA_integer_, 0L))
    return(dplyr::bind_rows(
      out,
      na_block_for_motifs(motifs_missing, outcome, model_label, n_used = NA_integer_)
    ))
  }
  
  presence <- tibble::tibble(
    Motif = motifs_present,
    n_present = purrr::map_int(motifs_present, ~ sum(mf[[.x]] == 1, na.rm = TRUE)),
    prop_present = n_present / n_used
  )
  
  model <- tryCatch(lm(fml, data = data), error = function(e) NULL)
  if (is.null(model)) {
    out <- na_block_for_motifs(motifs_present, outcome, model_label, n_used = n_used)
    out <- dplyr::left_join(out, presence, by = "Motif")
    return(dplyr::bind_rows(
      out,
      na_block_for_motifs(motifs_missing, outcome, model_label, n_used = NA_integer_)
    ))
  }
  
  coefs <- broom::tidy(model) %>%
    dplyr::filter(term %in% motifs_present) %>%
    dplyr::transmute(
      Motif = term,
      Outcome = outcome,
      Model = model_label,
      n_used = n_used,
      estimate = estimate,
      std.error = std.error,
      statistic = statistic,
      p.value = p.value,
      FoldChange = 10^estimate
    ) %>%
    dplyr::left_join(presence, by = "Motif")
  
  missing_part <- na_block_for_motifs(motifs_missing, outcome, model_label, n_used = NA_integer_) %>%
    dplyr::mutate(n_present = NA_integer_, prop_present = NA_real_)
  
  dplyr::bind_rows(coefs, missing_part) %>%
    dplyr::mutate(Motif = factor(Motif, levels = motifs)) %>%
    dplyr::arrange(Motif) %>%
    dplyr::mutate(Motif = as.character(Motif))
}

fit_motif_models_list_both <- function(data_list, outcomes, motifs, include_exp = FALSE) {
  purrr::imap_dfr(data_list, function(df, nm) {
    purrr::map_dfr(outcomes, function(outcome) {
      
      multiv <- fit_motif_model_multiv_df(
        data = df, outcome = outcome, motifs = motifs, include_exp = include_exp
      ) %>% dplyr::mutate(Analysis = "multivariate")
      
      univ <- fit_motif_model_one_df(
        data = df, outcome = outcome, motifs = motifs, include_exp = include_exp
      ) %>% dplyr::mutate(Analysis = "univariate")
      
      dplyr::bind_rows(multiv, univ) %>%
        dplyr::mutate(
          sample_name = nm,
          sig = sig_stars(p.value)
        )
    })
  }) %>%
    dplyr::select(sample_name, Outcome, Analysis, Model, Motif, sig, dplyr::everything())
}
```

# Define motifs and run analyses

Create the summary table of regression analyses for the selected motifs. The
reduced motif set (`motifs_10`) is the one shown in the figures.

```{r run_models}
motifs <- c("Ohler1", "DRE", "TATAbox", "INR", "Ebox", "Ohler6", "Ohler7",
            "DPE", "MTE", "TCT", "BREu", "BREd", "DCE1", "DCE2", "DCE3")

# Reduced set used for the published panels.
motifs_10 <- c("DPE", "INR", "Ebox", "Ohler6", "Ohler1", "TATAbox")
print(motifs_10)

outcomes <- c("bsize", "bfreq", "exp")

results_both <- dplyr::bind_rows(
  fit_motif_models_list_both(param_list, outcomes, motifs_10, include_exp = FALSE),
  fit_motif_models_list_both(param_list, outcomes, motifs_10, include_exp = TRUE)
)

head(results_both)

# Optional manual export:
# write.table(
#   results_both,
#   paste0(out_dir, "linear_regression_results_STAP_KS_mu.txt"),
#   quote = FALSE, row.names = FALSE, sep = "\t"
# )
```

# Heatmap plotting function

Visualizes motif-associated effect estimates across samples. Tile colour is the
direction/magnitude of the estimated effect; significance stars show the raw
p-value category. Motifs and samples can be reordered, and tiles for motifs
occurring too infrequently in a sample can be suppressed.

```{r heatmap_function}
sym_limits <- function(x, midpoint = 0) {
  x <- x[is.finite(x)]
  if (length(x) == 0) return(c(-1, 1))
  max_abs <- max(abs(x - midpoint))
  c(midpoint - max_abs, midpoint + max_abs)
}

make_heatmap_model <- function(results, model_name,
                    motifs_order = NULL,
                    samples_order = NULL,
                    min_occurrence = 0,
                    star_size = 3,
                    analysis = NULL) {
  
  plot_df <- results %>%
    dplyr::filter(Model == model_name, Outcome %in% c("bsize", "bfreq", "exp"))
  
  analysis_label <- if (is.null(analysis)) "both (univariate + multivariable)" else analysis
  
  if (!is.null(analysis)) {
    plot_df <- plot_df %>% dplyr::filter(Analysis == analysis)
  }
  
  if (!is.null(samples_order)) {
    plot_df <- plot_df %>%
      dplyr::filter(sample_name %in% samples_order) %>%
      dplyr::mutate(sample_name = factor(sample_name, levels = samples_order))
  } else {
    plot_df <- plot_df %>% dplyr::mutate(sample_name = factor(sample_name))
  }
  
  plot_df <- plot_df %>%
    dplyr::mutate(
      value = estimate,
      value = ifelse(!is.na(n_present) & n_present < min_occurrence, NA_real_, value),
      Sig_raw_plot = ifelse(sig %in% c("*", "**", "***"), sig, "")
    )
  
  if (!is.null(motifs_order)) {
    plot_df <- plot_df %>% dplyr::mutate(Motif = factor(Motif, levels = motifs_order))
  } else {
    plot_df <- plot_df %>% dplyr::mutate(Motif = factor(Motif))
  }
  
  lims <- sym_limits(plot_df$value, midpoint = 0)
  
  ggplot2::ggplot(plot_df, ggplot2::aes(x = sample_name, y = Motif)) +
    ggplot2::geom_tile(ggplot2::aes(fill = value)) +
    ggplot2::geom_text(ggplot2::aes(label = Sig_raw_plot), size = star_size, na.rm = TRUE) +
    ggplot2::facet_grid(. ~ Outcome) +
    ggplot2::scale_fill_gradient2(
      low = "blue", mid = "white", high = "red",
      midpoint = 0,
      limits = lims,
      na.value = "grey80"
    ) +
    ggplot2::labs(
      title = paste0("Motif effects (", analysis_label, "): ", model_name),
      subtitle = paste0(
        "Stars = raw p-value; tiles shown only if n_present >= ", min_occurrence,
        if (!is.null(samples_order)) paste0("; samples ordered (n=", length(samples_order), ")") else ""
      ),
      x = "Sample",
      y = "Motif",
      fill = "estimate"
    ) +
    ggplot2::theme_minimal(base_size = 11) +
    ggplot2::theme(
      panel.grid = ggplot2::element_blank(),
      axis.text.x = ggplot2::element_text(angle = 45, hjust = 1, vjust = 1),
      strip.text = ggplot2::element_text(face = "bold")
    )
}

min_occurrence <- 5
motifs_order   <- motifs_10
samples_order  <- pooled_samples
all_samples    <- sample_names
```

# FIGURE 2D — regression-coefficient heatmap (pooled samples)

Multivariate ("motifs_only") regression coefficients per motif for the three
pooled samples (zfh1, CG4822, Pld). The multivariate panel is Figure 2D; the
univariate version is kept as an exploratory companion.

```{r fig2d_heatmap_pooled}
q_multiv <- make_heatmap_model(
  results_both, "motifs_only",
  motifs_order = motifs_order,
  samples_order = samples_order,
  min_occurrence = min_occurrence,
  analysis = "multivariate"
)

q_univ <- make_heatmap_model(
  results_both, "motifs_only",
  motifs_order = motifs_order,
  samples_order = samples_order,
  min_occurrence = min_occurrence,
  analysis = "univariate"
)

q_multiv   # <- Figure 2D
q_univ     # exploratory (univariate)

# Optional manual PDF export:
# pdf(paste0(out_dir, "heatmap_multivariate_pooled.pdf"), width = 10, height = 5)
# print(q_multiv); dev.off()
# pdf(paste0(out_dir, "heatmap_univariate_pooled.pdf"),  width = 10, height = 5)
# print(q_univ);   dev.off()
```

# SUPPL. FIG. 3G — regression-coefficient heatmap (all samples)

Same multivariate model, shown for all nine samples (pool + two replicates per
condition).

```{r suppl3g_heatmap_all}
p_multiv <- make_heatmap_model(
  results_both, "motifs_only",
  motifs_order = motifs_order,
  samples_order = all_samples,
  min_occurrence = min_occurrence,
  analysis = "multivariate"
)

p_univ <- make_heatmap_model(
  results_both, "motifs_only",
  motifs_order = motifs_order,
  samples_order = all_samples,
  min_occurrence = min_occurrence,
  analysis = "univariate"
)

p_multiv   # <- Supplemental Figure 3G
p_univ     # exploratory (univariate)

# Optional manual PDF export:
# pdf(paste0(out_dir, "heatmap_multivariate_all_samples.pdf"), width = 14, height = 5)
# print(p_multiv); dev.off()
# pdf(paste0(out_dir, "heatmap_univariate_all_samples.pdf"),   width = 14, height = 5)
# print(p_univ);   dev.off()
```

# EXPLORATORY — median-difference / log2 fold-change heatmaps

For each motif, compares the median burst parameter between promoters that
carry the motif and those that do not (Wilcoxon test), expressed as a log2
fold-change. Not a published panel, but a useful non-parametric cross-check of
the regression heatmaps above.

```{r log2fc_function}
analyze_param_list_log2FC <- function(param_list,
                    motifs_order,
                    samples_order = NULL) {
  stopifnot(is.list(param_list))
  
  if (is.null(names(param_list)) || any(names(param_list) == "")) {
    names(param_list) <- paste0("sample_", seq_along(param_list))
  }
  
  if (!is.null(samples_order)) {
    keep <- intersect(samples_order, names(param_list))
    param_list <- param_list[keep]
    param_list <- param_list[samples_order[samples_order %in% names(param_list)]]
  }
  
  add_significance <- function(p) {
    dplyr::case_when(
      is.na(p)  ~ "",
      p < 0.001 ~ "***",
      p < 0.01  ~ "**",
      p < 0.05  ~ "*",
      TRUE      ~ ""
    )
  }
  
  sym_limits <- function(x, midpoint = 0) {
    x <- x[is.finite(x)]
    if (length(x) == 0) return(c(-1, 1))
    max_abs <- max(abs(x - midpoint), na.rm = TRUE)
    c(midpoint - max_abs, midpoint + max_abs)
  }
  
  results_df <- purrr::imap_dfr(param_list, function(df, sample_name) {
    df <- data.table::as.data.table(df)
    
    purrr::map_dfr(motifs_order, function(m) {
      if (!m %in% names(df)) {
        return(tibble::tibble(
          sample_name = sample_name,
          motif = m,
          measurement = c("bfreq", "bsize", "exp"),
          value_0 = NA_real_, value_1 = NA_real_, diff = NA_real_, log2FC = NA_real_,
          pvalue = NA_real_, Nobservations_0 = NA_integer_, Nobservations_1 = NA_integer_
        ))
      }
      
      dt_m <- df[, .(bfreq, bsize, exp, motif = get(m))]
      
      n0 <- sum(dt_m$motif == 0, na.rm = TRUE)
      n1 <- sum(dt_m$motif > 0, na.rm = TRUE)
      
      one_measure <- function(y) {
        med0 <- dt_m[motif == 0, median(get(y), na.rm = TRUE)]
        med1 <- dt_m[motif > 0, median(get(y), na.rm = TRUE)]
        wt <- suppressWarnings(stats::wilcox.test(get(y) ~ (motif > 0), data = dt_m))
        
        tibble::tibble(
          sample_name = sample_name,
          motif = m,
          measurement = y,
          value_0 = med0,
          value_1 = med1,
          diff = med1 - med0,
          log2FC = log2(med1 / med0),
          pvalue = wt$p.value,
          Nobservations_0 = n0,
          Nobservations_1 = n1
        )
      }
      
      dplyr::bind_rows(one_measure("bfreq"), one_measure("bsize"), one_measure("exp"))
    })
  }) %>%
    dplyr::mutate(
      significance = add_significance(pvalue),
      motif = factor(motif, levels = motifs_order),
      sample_name = factor(sample_name, levels = names(param_list)),
      measurement = dplyr::recode(
        measurement,
        bfreq = "Burst Frequency",
        bsize = "Burst Size",
        exp   = "Mean expression"
      )
    )
  
  lims <- sym_limits(results_df$log2FC, midpoint = 0)
  
  p <- ggplot2::ggplot(results_df, ggplot2::aes(x = sample_name, y = motif, fill = log2FC)) +
    ggplot2::geom_tile() +
    ggplot2::geom_text(ggplot2::aes(label = significance), size = 3) +
    ggplot2::facet_grid(. ~ measurement) +
    ggplot2::scale_fill_gradient2(
      limits = lims,
      low = "blue", mid = "white", high = "red",
      midpoint = 0,
      oob = scales::squish,
      na.value = "grey90",
      name = "log2FC"
    ) +
    ggplot2::theme_minimal() +
    ggplot2::theme(
      panel.grid = ggplot2::element_blank(),
      axis.text.x = ggplot2::element_text(angle = 45, hjust = 1),
      axis.title.x = ggplot2::element_blank(),
      axis.title.y = ggplot2::element_blank(),
      strip.text = ggplot2::element_text(face = "bold")
    )
  
  list(table = results_df, plot = p)
}
```

```{r log2fc_plots}
res_pooled <- analyze_param_list_log2FC(
  param_list,
  motifs_order = motifs_10,
  samples_order = pooled_samples
)
pooled_table <- res_pooled$table
head(pooled_table)
res_pooled$plot

res_all <- analyze_param_list_log2FC(param_list, motifs_order = motifs_10)
all_table <- res_all$table
head(all_table)
res_all$plot

# Optional manual PDF export:
# pdf(paste0(out_dir, "log2FC_heatmap_pooled.pdf"),      width = 12, height = 5)
# print(res_pooled$plot); dev.off()
# pdf(paste0(out_dir, "log2FC_heatmap_all_samples.pdf"), width = 12, height = 5)
# print(res_all$plot);    dev.off()
```

# Meta-analysis combining datasets (input for Figure 2E)

Fixed-effects, inverse-variance meta-analysis of the per-sample multivariate
regression coefficients (Model == "motifs_only", no `log10(exp)` adjustment),
restricted to the three pooled samples. Pooled estimates summarise the overall
direction and magnitude of each motif's association with burst size and burst
frequency on the log10 scale.

```{r meta_analysis}
min_k <- 3

meta_input <- results_both %>%
  filter(sample_name %in% pooled_samples[1:3] & Analysis == "multivariate" & Model == "motifs_only") %>%
  filter(
    is.finite(estimate),
    is.finite(std.error),
    std.error > 0,
    !is.na(sample_name)
  )

combos <- meta_input %>%
  distinct(sample_name, Motif, Outcome) %>%
  dplyr::count(Motif, Outcome, name = "k") %>%
  filter(k >= min_k)

pooled_list <- vector("list", nrow(combos))

for (i in seq_len(nrow(combos))) {
  subset_data <- meta_input %>%
    filter(Motif == combos$Motif[i], Outcome == combos$Outcome[i]) %>%
    distinct(sample_name, .keep_all = TRUE)
  
  meta_obj <- rma(
    yi = subset_data$estimate,
    sei = subset_data$std.error,
    method = "FE"
  )
  
  pooled_list[[i]] <- data.frame(
    Motif = combos$Motif[i],
    Outcome = combos$Outcome[i],
    k = nrow(subset_data),
    pooled_estimate = as.numeric(meta_obj$beta[1]),
    pooled_SE = as.numeric(meta_obj$se),
    pooled_pValue = as.numeric(meta_obj$pval),
    pooled_zValue = as.numeric(meta_obj$zval),
    pooled_FoldChange = 10^as.numeric(meta_obj$beta[1]),
    pooled_FC_low  = 10^as.numeric(meta_obj$ci.lb),
    pooled_FC_high = 10^as.numeric(meta_obj$ci.ub),
    tau2 = as.numeric(meta_obj$tau2),
    I2 = as.numeric(meta_obj$I2)
  )
}

pooled_results <- bind_rows(pooled_list) %>%
  mutate(
    Sig_raw = case_when(
      pooled_pValue < 0.001 ~ "***",
      pooled_pValue < 0.01  ~ "**",
      pooled_pValue < 0.05  ~ "*",
      TRUE ~ "ns"
    )
  )

head(pooled_results)
```

# Reshape pooled results for Figure 2E

Reshape to wide format so burst-size and burst-frequency effects for the same
motif sit in one row; add approximate 95% CIs and a Both/size-only/freq-only/
Neither significance class.

```{r meta_reshape}
df_wide <- pooled_results %>%
  filter(Outcome %in% c("bsize", "bfreq")) %>%
  select(Motif, Outcome,
         pooled_estimate, pooled_SE, pooled_pValue,
         pooled_FoldChange, pooled_FC_low, pooled_FC_high,
         k, tau2, I2) %>%
  pivot_wider(
    names_from = Outcome,
    values_from = c(pooled_estimate, pooled_SE, pooled_pValue,
                    pooled_FoldChange, pooled_FC_low, pooled_FC_high,
                    k, tau2, I2),
    names_sep = "_"
  ) %>%
  filter(!is.na(pooled_FoldChange_bsize), !is.na(pooled_FoldChange_bfreq))

df_wide <- df_wide %>%
  mutate(
    est_low_bsize  = pooled_estimate_bsize - 1.96 * pooled_SE_bsize,
    est_high_bsize = pooled_estimate_bsize + 1.96 * pooled_SE_bsize,
    est_low_bfreq  = pooled_estimate_bfreq - 1.96 * pooled_SE_bfreq,
    est_high_bfreq = pooled_estimate_bfreq + 1.96 * pooled_SE_bfreq
  )

df_wide <- df_wide %>%
  mutate(
    Significance = case_when(
      pooled_pValue_bsize < 0.05 & pooled_pValue_bfreq < 0.05  ~ "Both",
      pooled_pValue_bsize < 0.05 & pooled_pValue_bfreq >= 0.05 ~ "bsize only",
      pooled_pValue_bsize >= 0.05 & pooled_pValue_bfreq < 0.05 ~ "bfreq only",
      TRUE ~ "Neither"
    )
  )

head(df_wide)
```

# FIGURE 2E — meta-analysis scatter

Each point is one motif; axes are the pooled regression coefficients on the
log10 scale (x = burst size, y = burst frequency). Error bars are the
approximate 95% CIs.

```{r fig2e_meta_plot}
p_meta <- ggplot(
  df_wide,
  aes(
    x = pooled_estimate_bsize,
    y = pooled_estimate_bfreq,
    color = Significance,
    label = Motif
  )
) +
  geom_errorbarh(
    aes(xmin = est_low_bsize, xmax = est_high_bsize),
    height = 0, alpha = 0.4
  ) +
  geom_errorbar(
    aes(ymin = est_low_bfreq, ymax = est_high_bfreq),
    width = 0, alpha = 0.4
  ) +
  geom_point(size = 3) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray40") +
  geom_hline(yintercept = 0, linetype = "dotted") +
  geom_vline(xintercept = 0, linetype = "dotted") +
  theme_classic(base_size = 14) +
  scale_color_manual(values = c(
    "Both" = "deepskyblue",
    "bsize only" = "cyan3",
    "bfreq only" = "royalblue3",
    "Neither" = "gray60"
  )) +
  labs(
    x = expression("Pooled effect on log"[10]*"(bsize)  (beta)"),
    y = expression("Pooled effect on log"[10]*"(bfreq)  (beta)"),
    title = "Meta-analysis: Effect of motif presence on burst parameters",
    subtitle = paste0("Points = pooled betas; error bars = 95% CI; min_k = ", min_k)
  )

if (requireNamespace("ggrepel", quietly = TRUE)) {
  p_meta <- p_meta + ggrepel::geom_text_repel(size = 3, max.overlaps = 30)
} else {
  p_meta <- p_meta + geom_text(size = 3, vjust = -0.5, check_overlap = TRUE)
}

p_meta   # <- Figure 2E

# Optional manual PDF export:
# pdf(paste0(out_dir, "meta_pooled_analysis_estimates_KS_mu.pdf"), width = 6, height = 5)
# print(p_meta); dev.off()
```

# SUPPL. FIG. 3A — motif-content donut + UpSet

Uses the full promoter-library annotation (Table4). `main_mots` concatenates
the motifs present in each promoter; the donut counts single-motif occurrences
and the UpSet plot shows the most frequent motif combinations.

```{r suppl3a_prepare}
TF_motifs <- motifs_10
names(TF_motifs) <- motifs_10

# Full promoter library annotation = Table4 (read in the data-loading chunk).
new_STAP <- as.data.table(anno)

# Build a "-"-separated string of the motifs present in each promoter.
new_STAP$main_mots <- ""
for (x in names(TF_motifs)) {
  column <- new_STAP[[x]]
  new_STAP$main_mots <- paste0(new_STAP$main_mots, ifelse(column == 0, "", paste0(x, "-")))
}
```

```{r suppl3a_donut_upset}
# --- Motif-frequency donut ----------------------------------------------------
TF_freq <- as.data.frame(table(unlist(strsplit(new_STAP$main_mots, "-"))))
TF_freq <- TF_freq[TF_freq$Var1 != "", ]

# Fix drawing order and colours.
desired_order <- rev(motifs_10)
TF_freq$Var1 <- factor(TF_freq$Var1, levels = desired_order)

motif_cols <- c(
  "TATAbox" = "darkcyan",
  "DPE"     = "goldenrod2",
  "Ebox"    = "coral3",
  "INR"     = "orchid4",
  "Ohler6"  = "lightblue",
  "Ohler1"  = "darkblue"
)

motif_plot <- ggplot(TF_freq, aes(x = 3, y = Freq, fill = Var1)) +
  geom_col() +
  coord_polar(theta = "y") +
  xlim(c(1.5, 3.5)) +
  scale_fill_manual(values = motif_cols) +
  theme_void()

motif_plot   # <- Supplemental Figure 3A (donut)

pdf(paste0(out_dir, "promoter_motif_content.pdf"), width = 6, height = 5)
print(motif_plot)
dev.off()

print(TF_freq)
#write.table(TF_freq, paste0(out_dir, "promoter_motif_count.txt"),
#            sep = "\t", quote = FALSE, col.names = FALSE)

# --- Motif-combination UpSet --------------------------------------------------
#Using ComplexUpset package 

motif_sets <- motifs_10

# ComplexUpset works best with logical set columns.
new_STAP_upset <- new_STAP %>%
  mutate(across(all_of(motif_sets), ~ as.logical(.)))

up <- upset(
  new_STAP_upset,
  intersect = motif_sets,
  name = "Motif combinations",
  min_size = 2,
  width_ratio = 0.15
)

up   # <- Supplemental Figure 3A (UpSet)

# pdf(paste0(out_dir, "promoter_motif_content_upset.pdf"), width = 6, height = 5)
# print(up)
# dev.off()
```

# FIGURE 2B — TATA box vs DPE vs other

Boxplots of burst size / frequency / mean expression for promoters classified
as TATA-box-only, DPE-only or none (promoters with both are excluded), for the
zfh1 pooled sample. ANOVA p-values are shown; pairwise t-tests are added as
significance brackets.

```{r fig2b_tata_dpe}
scatter.plots <- list()

dat.sub <- param %>%
  filter(KS_pval > 0.01 & high_mu == TRUE & !is.na(TATAbox) &
         condition == "zfh1" & rep == "pool") %>%
  mutate(
    promoter_type = case_when(
      TATAbox == 1 & DPE == 1 ~ "both",
      TATAbox == 1 & DPE == 0 ~ "TATAbox",
      TATAbox == 0 & DPE == 1 ~ "DPE",
      TATAbox == 0 & DPE == 0 ~ "none"
    )
  ) %>%
  filter(promoter_type != "both")

dat.sub$promoter_type <- factor(dat.sub$promoter_type, levels = c("TATAbox", "DPE", "both", "none"))

my_comparisons <- list(c("TATAbox", "DPE"), c("TATAbox", "none"), c("DPE", "none"))

# --- Burst size ---
p_val1 <- summary(aov(log10(bsize) ~ promoter_type, dat.sub))[[1]][["Pr(>F)"]][1]
scatter.plots[["burst size"]] <- dat.sub %>%
  ggplot(aes(x = promoter_type, y = log10(bsize), color = promoter_type)) +
  geom_boxplot(outlier.shape = NA) +
  scale_color_manual(values = c("darkcyan", "goldenrod2", "gray32", "gray12")) +
  facet_wrap(~lib) +
  ggtitle("burst size", subtitle = paste0("ANOVA p = ", format(p_val1, scientific = TRUE, digits = 3))) +
  theme_classic() +
  theme(text = element_text(size = 15), legend.position = "none") +
  stat_compare_means(comparisons = my_comparisons, method = "t.test", label = "p.signif")

# --- Burst frequency ---
p_val2 <- summary(aov(log10(bfreq) ~ promoter_type, dat.sub))[[1]][["Pr(>F)"]][1]
scatter.plots[["burst frequency"]] <- dat.sub %>%
  ggplot(aes(x = promoter_type, y = log10(bfreq), color = promoter_type)) +
  geom_boxplot(outlier.shape = NA) +
  scale_color_manual(values = c("darkcyan", "goldenrod2", "gray32", "gray12")) +
  facet_wrap(~lib) +
  ggtitle("burst frequency", subtitle = paste0("ANOVA p = ", format(p_val2, scientific = TRUE, digits = 3))) +
  theme_classic() +
  theme(text = element_text(size = 15), legend.position = "none") +
  stat_compare_means(comparisons = my_comparisons, method = "t.test", label = "p.signif")

# --- Mean expression ---
p_val3 <- summary(aov(log10(exp) ~ promoter_type, dat.sub))[[1]][["Pr(>F)"]][1]
scatter.plots[["exp"]] <- dat.sub %>%
  ggplot(aes(x = promoter_type, y = log10(exp), color = promoter_type)) +
  geom_boxplot(outlier.shape = NA) +
  scale_color_manual(values = c("darkcyan", "goldenrod2", "gray32", "gray12")) +
  facet_wrap(~lib) +
  ggtitle("exp", subtitle = paste0("ANOVA p = ", format(p_val3, scientific = TRUE, digits = 3))) +
  theme_classic() +
  theme(text = element_text(size = 15), legend.position = "none") +
  stat_compare_means(comparisons = my_comparisons, method = "t.test", label = "p.signif")

wrap_plots(scatter.plots)   # <- Figure 2B

pdf(paste0(out_dir, "DPE_TATA_comparison_zfh1_pool.pdf"), width = 7, height = 5)
print(wrap_plots(scatter.plots))
dev.off()
```

# FIGURE 2C — motif PWM score vs burst parameters

*** THIS CHUNK IS DISABLED (`eval = FALSE`). ***

It needs external inputs that are NOT in the supplemental tables (see the
"MISSING / EXTERNAL INPUTS" block at the top of this file):

  * data/CP_motifs_PWM.rds                          (core-promoter motif PWMs)
  * data/Franzi_STAP_STARR_TWIST_seqonly-v2.tsv     (promoter/oligo sequences)

It also contains a motif-scanning step that was MISSING from the original
script (the line that defines `motif_hits`). A reconstructed
`motifmatchr::matchMotifs(...)` call is provided and clearly flagged --- verify
it against the original analysis before enabling this chunk.

To run Fig. 2C: place the two files under ./data, confirm/adjust the
`motif_hits` step, then set `eval = TRUE`.

```{r fig2c_motif_score, eval = FALSE}
require(TFBSTools)
library(Biostrings)
library(motifmatchr)

# --- Import core-promoter motif PWMs -----------------------------------------
#this file is not provided but needs to be generated 
pwm <- readRDS("/groups/stark/vloubiere/motifs_db/CP_motifs_PWM.rds")

motif_names <- sapply(pwm$Pwms_perc, TFBSTools::name)

# Extract the raw numeric percentage matrices from the S4 objects.
raw_percentage_matrices <- lapply(pwm$Pwms_perc, function(x) TFBSTools::Matrix(x))

# Convert percentages to integer "counts" (TFBSTools/motifmatchr prefer counts).
count_matrices <- lapply(raw_percentage_matrices, function(x) round(x * 100))

# Build a PFMatrixList (position frequency matrices).
pfm_list <- do.call(PFMatrixList, lapply(seq_along(count_matrices), function(i) {
  orig_names <- names(pwm$Pwms_perc)
  m_name <- if (!is.null(orig_names)) orig_names[i] else paste0("Motif_", i)
  PFMatrix(name = m_name, profileMatrix = count_matrices[[i]])
}))

# Convert to log-odds PWMs (format required by motifmatchr).
fixed_pwms <- toPWM(pfm_list, type = "log2probratio", pseudocounts = 0.8)

# --- Import promoter sequences (with linker) into a DNAStringSet -------------
#these sequences also need to be generated using genomic coordinates 
seq <- fread("/groups/stark/lorbeer/STAP-TWIST/Franzi_STAP_STARR_TWIST_seqonly-v2.tsv")[grep("STAP", ID_TWIST)]
seqV <- substr(seq$Sequence_with_linker, 59, 59 + 132)
names(seqV) <- seq$ID_TWIST
dna_seqs <- DNAStringSet(seqV)

# -----------------------------------------------------------------------------
# NOTE: the original script is MISSING the motif-scanning call that produces
# `motif_hits`. Downstream code expects a per-motif list of hit ranges carrying
# a `score` column and a `group` index into dna_seqs, which is what
# motifmatchr::matchMotifs(..., out = "positions") returns. The line below is a
# RECONSTRUCTION and must be verified against the original analysis.
# -----------------------------------------------------------------------------
motif_hits <- motifmatchr::matchMotifs(
  fixed_pwms, dna_seqs,
  out = "positions",
  bg = "even",
  p.cutoff = 5e-4
)

# --- Unpack the per-motif hit list into a single data.table ------------------
res_list <- lapply(seq_along(motif_hits), function(i) {
  hits <- motif_hits[[i]]
  if (length(hits) > 0) {
    dt <- as.data.table(as.data.frame(hits))
    dt[, motif := motif_names[i]]
    return(dt)
  }
})
bed <- rbindlist(res_list, fill = TRUE)

# Map the sequence index back to the sequence name and clean it up.
bed[, seqlvls := names(dna_seqs)[group]]
bed[, seqlvls := gsub("_STAP-v2$", "", seqlvls)]

if ("strand" %in% names(bed)) bed <- bed[strand == "+"]

# Motif centre position and the max score per sequence/motif.
bed[, pos := (start + end) / 2]
bed[, seqlvls := gsub("_STAP-v2$", "", as.character(seqlvls))]
bed <- bed[, .SD[which.max(score)], .(seqlvls, motif)]

# Wide score and position matrices (one row per sequence).
score_mat <- dcast(bed, seqlvls ~ motif, value.var = "score", fill = 0)
pos_mat   <- dcast(bed, seqlvls ~ motif, value.var = "pos")

head(score_mat)

# --- Join the burst parameters of the zfh1 pooled sample ---------------------
param_28 <- param[sample == "libFL001_STAP_zfh1_pool" & (high_mu & KS_pval > 0.01)]
sub.score <- merge(score_mat, param_28[, .(seqlvls = seq_ID, bfreq, bsize, log2FoldChange)],
                   by = "seqlvls", sort = FALSE)
sub.pos <- pos_mat[sub.score$seqlvls, on = "seqlvls"]
stopifnot(identical(sub.score$seqlvls, sub.pos$seqlvls))

# --- Score vs burst size / frequency, all hits (coloured by position) --------
pdf(paste0(out_dir, "motif_score_vs_size_freq.pdf"), height = 3.5, width = 7.5)
for (motif in c("TATAbox", "DPE")) {
  pos.ok <- if (motif == "TATAbox") between(sub.pos[[motif]], 35, 45)
            else if (motif == "DPE") between(sub.pos[[motif]], 80, 100)
            else stop("Please define expected position for motif")
  
  dt_plot <- data.table(
    score = sub.score[[motif]],
    freq  = log10(sub.score$bfreq),
    size  = log10(sub.score$bsize),
    pos_ok = factor(pos.ok, levels = c("FALSE", "TRUE"))
  )
  
  cor_freq <- cor.test(dt_plot$score, dt_plot$freq, method = "pearson")
  anno_freq <- sprintf("PCC = %.2f\np-val = %.2e", cor_freq$estimate, cor_freq$p.value)
  p_freq <- ggplot(dt_plot, aes(x = score, y = freq)) +
    geom_point(aes(color = pos_ok), alpha = 0.3, size = 1) +
    geom_smooth(method = "lm", color = "blue", se = FALSE, linewidth = 0.8) +
    scale_color_manual(values = c("TRUE" = "red", "FALSE" = "black")) +
    annotate("text", x = -Inf, y = -Inf, label = anno_freq,
             vjust = -0.5, hjust = -0.1, fontface = "bold", size = 3.5) +
    labs(x = "Motif score", y = "Burst Frequency (log10)", title = motif) +
    theme_classic() +
    theme(text = element_text(size = 15), legend.position = "none")
  
  cor_size <- cor.test(dt_plot$score, dt_plot$size, method = "pearson")
  anno_size <- sprintf("PCC = %.2f\np-val = %.2e", cor_size$estimate, cor_size$p.value)
  p_size <- ggplot(dt_plot, aes(x = score, y = size)) +
    geom_point(aes(color = pos_ok), alpha = 0.3, size = 1) +
    geom_smooth(method = "lm", color = "blue", se = FALSE, linewidth = 0.8) +
    scale_color_manual(values = c("TRUE" = "red", "FALSE" = "black")) +
    annotate("text", x = -Inf, y = -Inf, label = anno_size,
             vjust = -0.5, hjust = -0.1, fontface = "bold", size = 3.5) +
    labs(x = "Motif score", y = "Burst Size (log10)", title = motif) +
    theme_classic() +
    theme(text = element_text(size = 15), legend.position = "none")
  
  grid.arrange(p_freq, p_size, ncol = 2)
}
dev.off()

# --- Figure 2C: same, restricted to correctly-positioned hits ----------------
pdf(paste0(out_dir, "motif_score_vs_size_freq_pos_ok.pdf"), height = 3.5, width = 7.5)
for (motif in c("TATAbox", "DPE")) {
  pos.ok <- if (motif == "TATAbox") between(sub.pos[[motif]], 35, 45)
            else if (motif == "DPE") between(sub.pos[[motif]], 80, 100)
            else stop("Please define expected position for motif")
  
  dt_plot <- data.table(
    score = sub.score[[motif]],
    freq  = log10(sub.score$bfreq),
    size  = log10(sub.score$bsize),
    pos_ok = pos.ok
  )[pos_ok == TRUE]
  
  pt_color <- if (motif == "TATAbox") "darkcyan" else "goldenrod2"
  
  cor_freq <- cor.test(dt_plot$score, dt_plot$freq, method = "pearson")
  anno_freq <- sprintf("PCC = %.2f\np-val = %.2e", cor_freq$estimate, cor_freq$p.value)
  p_freq <- ggplot(dt_plot, aes(x = score, y = freq)) +
    geom_point(color = pt_color, alpha = 0.5, size = 2) +
    geom_smooth(method = "lm", color = "black", se = FALSE, linewidth = 0.8) +
    annotate("text", x = -Inf, y = -Inf, label = anno_freq,
             vjust = -0.5, hjust = -0.1, fontface = "bold", size = 3.5) +
    labs(x = "Motif score", y = "Burst frequency (log10)", title = motif) +
    theme_classic() +
    theme(text = element_text(size = 15), legend.position = "none")
  
  cor_size <- cor.test(dt_plot$score, dt_plot$size, method = "pearson")
  anno_size <- sprintf("PCC = %.2f\np-val = %.2e", cor_size$estimate, cor_size$p.value)
  p_size <- ggplot(dt_plot, aes(x = score, y = size)) +
    geom_point(color = pt_color, alpha = 0.5, size = 2) +
    geom_smooth(method = "lm", color = "black", se = FALSE, linewidth = 0.8) +
    annotate("text", x = -Inf, y = -Inf, label = anno_size,
             vjust = -0.5, hjust = -0.1, fontface = "bold", size = 3.5) +
    labs(x = "Motif score", y = "Burst size (log10)", title = motif) +
    theme_classic() +
    theme(text = element_text(size = 15), legend.position = "none")
  
  grid.arrange(p_freq, p_size, ncol = 2)
}
dev.off()
```

# SUPPL. FIG. 3B — TATA box boxplots split by expression bin

For the zfh1 pooled sample, promoters are binned by mean expression
(low / medium / high) and TATA-box-positive vs -negative promoters are compared
within each bin (Wilcoxon test).

```{r suppl3b_expression_bins_setup}
library(ggpubr)

STAP_annotated <- dat.sub %>%
  filter(between(exp, 0.1, 5)) %>%
  mutate(
    expression_range = ifelse(between(exp, 1.5, 5), "high",
                       ifelse(between(exp, 0.3, 1.5), "medium", "low")),
    expression_range = factor(expression_range, levels = c("low", "medium", "high"))
  )
```

```{r suppl3b_tata}
df_base <- STAP_annotated %>% mutate(TATAbox = as.factor(TATAbox))

p_bsize <- ggplot(df_base, aes(x = TATAbox, y = log10(bsize))) +
  geom_boxplot(aes(color = TATAbox), alpha = 0.5, outlier.shape = NA) +
  facet_grid(sample ~ expression_range) +
  scale_color_manual(values = c("0" = "grey32", "1" = "darkcyan")) +
  stat_compare_means(method = "wilcox.test", label = "p.format", comparisons = list(c("0", "1"))) +
  theme_classic() + theme(text = element_text(size = 15), legend.position = "none")

p_bfreq <- ggplot(df_base, aes(x = TATAbox, y = log10(bfreq))) +
  geom_boxplot(aes(color = TATAbox), alpha = 0.5, outlier.shape = NA) +
  facet_grid(sample ~ expression_range) +
  scale_color_manual(values = c("0" = "grey32", "1" = "darkcyan")) +
  stat_compare_means(method = "wilcox.test", label = "p.format", comparisons = list(c("0", "1"))) +
  theme_classic() + theme(text = element_text(size = 15), legend.position = "none")

p_exp <- ggplot(df_base, aes(x = TATAbox, y = log10(exp))) +
  geom_boxplot(aes(color = TATAbox), alpha = 0.5, outlier.shape = NA) +
  facet_grid(sample ~ expression_range) +
  scale_color_manual(values = c("0" = "grey32", "1" = "darkcyan")) +
  stat_compare_means(method = "wilcox.test", label = "p.format", comparisons = list(c("0", "1"))) +
  theme_classic() + theme(text = element_text(size = 15), legend.position = "none")

grid.arrange(p_bsize, p_bfreq, p_exp, ncol = 3)   # <- Supplemental Figure 3B
# ggsave(paste0(out_dir, "TATA_boxplots_expressionbins_zfh1.pdf"),
#        plot = grid.arrange(p_bsize, p_bfreq, p_exp, ncol = 3),
#        width = 8, height = 3, units = "in")
```

# EXPLORATORY — expression-bin boxplots for the other motifs

The same low/medium/high expression-bin comparison as Supplemental Figure 3B,
repeated for DPE, INR, E-box, Ohler6 and Ohler1. Kept as exploratory support.

```{r explore_dpe}
df_base <- STAP_annotated %>% mutate(DPE = as.factor(DPE))

p_bsize <- ggplot(df_base, aes(x = DPE, y = log10(bsize))) +
  geom_boxplot(aes(color = DPE), alpha = 0.5, outlier.shape = NA) +
  facet_grid(sample ~ expression_range) +
  scale_color_manual(values = c("0" = "grey32", "1" = "goldenrod2")) +
  stat_compare_means(method = "wilcox.test", label = "p.format", comparisons = list(c("0", "1"))) +
  theme_classic() + theme(text = element_text(size = 15), legend.position = "none")

p_bfreq <- ggplot(df_base, aes(x = DPE, y = log10(bfreq))) +
  geom_boxplot(aes(color = DPE), alpha = 0.5, outlier.shape = NA) +
  facet_grid(sample ~ expression_range) +
  scale_color_manual(values = c("0" = "grey32", "1" = "goldenrod2")) +
  stat_compare_means(method = "wilcox.test", label = "p.format", comparisons = list(c("0", "1"))) +
  theme_classic() + theme(text = element_text(size = 15), legend.position = "none")

p_exp <- ggplot(df_base, aes(x = DPE, y = log10(exp))) +
  geom_boxplot(aes(color = DPE), alpha = 0.5, outlier.shape = NA) +
  facet_grid(sample ~ expression_range) +
  scale_color_manual(values = c("0" = "grey32", "1" = "goldenrod2")) +
  stat_compare_means(method = "wilcox.test", label = "p.format", comparisons = list(c("0", "1"))) +
  theme_classic() + theme(text = element_text(size = 15), legend.position = "none")

grid.arrange(p_bsize, p_bfreq, p_exp, ncol = 3)
```

```{r explore_inr}
df_base <- STAP_annotated %>% mutate(INR = as.factor(INR))

p_bsize <- ggplot(df_base, aes(x = INR, y = log10(bsize))) +
  geom_boxplot(aes(color = INR), alpha = 0.5, outlier.shape = NA) +
  facet_grid(sample ~ expression_range) +
  scale_color_manual(values = c("0" = "grey32", "1" = "orchid4")) +
  stat_compare_means(method = "wilcox.test", label = "p.format", comparisons = list(c("0", "1"))) +
  theme_classic() + theme(text = element_text(size = 15), legend.position = "none")

p_bfreq <- ggplot(df_base, aes(x = INR, y = log10(bfreq))) +
  geom_boxplot(aes(color = INR), alpha = 0.5, outlier.shape = NA) +
  facet_grid(sample ~ expression_range) +
  scale_color_manual(values = c("0" = "grey32", "1" = "orchid4")) +
  stat_compare_means(method = "wilcox.test", label = "p.format", comparisons = list(c("0", "1"))) +
  theme_classic() + theme(text = element_text(size = 15), legend.position = "none")

p_exp <- ggplot(df_base, aes(x = INR, y = log10(exp))) +
  geom_boxplot(aes(color = INR), alpha = 0.5, outlier.shape = NA) +
  facet_grid(sample ~ expression_range) +
  scale_color_manual(values = c("0" = "grey32", "1" = "orchid4")) +
  stat_compare_means(method = "wilcox.test", label = "p.format", comparisons = list(c("0", "1"))) +
  theme_classic() + theme(text = element_text(size = 15), legend.position = "none")

grid.arrange(p_bsize, p_bfreq, p_exp, ncol = 3)
```

```{r explore_ebox}
df_base <- STAP_annotated %>% mutate(Ebox = as.factor(Ebox))

p_bsize <- ggplot(df_base, aes(x = Ebox, y = log10(bsize))) +
  geom_boxplot(aes(color = Ebox), alpha = 0.5, outlier.shape = NA) +
  facet_grid(sample ~ expression_range) +
  scale_color_manual(values = c("0" = "grey32", "1" = "coral3")) +
  stat_compare_means(method = "wilcox.test", label = "p.format", comparisons = list(c("0", "1"))) +
  theme_classic() + theme(text = element_text(size = 15), legend.position = "none")

p_bfreq <- ggplot(df_base, aes(x = Ebox, y = log10(bfreq))) +
  geom_boxplot(aes(color = Ebox), alpha = 0.5, outlier.shape = NA) +
  facet_grid(sample ~ expression_range) +
  scale_color_manual(values = c("0" = "grey32", "1" = "coral3")) +
  stat_compare_means(method = "wilcox.test", label = "p.format", comparisons = list(c("0", "1"))) +
  theme_classic() + theme(text = element_text(size = 15), legend.position = "none")

p_exp <- ggplot(df_base, aes(x = Ebox, y = log10(exp))) +
  geom_boxplot(aes(color = Ebox), alpha = 0.5, outlier.shape = NA) +
  facet_grid(sample ~ expression_range) +
  scale_color_manual(values = c("0" = "grey32", "1" = "coral3")) +
  stat_compare_means(method = "wilcox.test", label = "p.format", comparisons = list(c("0", "1"))) +
  theme_classic() + theme(text = element_text(size = 15), legend.position = "none")

grid.arrange(p_bsize, p_bfreq, p_exp, ncol = 3)
```

```{r explore_ohler6}
df_base <- STAP_annotated %>% mutate(Ohler6 = as.factor(Ohler6))

p_bsize <- ggplot(df_base, aes(x = Ohler6, y = log10(bsize))) +
  geom_boxplot(aes(color = Ohler6), alpha = 0.5, outlier.shape = NA) +
  facet_grid(sample ~ expression_range) +
  scale_color_manual(values = c("0" = "grey32", "1" = "darkblue")) +
  stat_compare_means(method = "wilcox.test", label = "p.format", comparisons = list(c("0", "1"))) +
  theme_classic() + theme(text = element_text(size = 15), legend.position = "none")

p_bfreq <- ggplot(df_base, aes(x = Ohler6, y = log10(bfreq))) +
  geom_boxplot(aes(color = Ohler6), alpha = 0.5, outlier.shape = NA) +
  facet_grid(sample ~ expression_range) +
  scale_color_manual(values = c("0" = "grey32", "1" = "darkblue")) +
  stat_compare_means(method = "wilcox.test", label = "p.format", comparisons = list(c("0", "1"))) +
  theme_classic() + theme(text = element_text(size = 15), legend.position = "none")

p_exp <- ggplot(df_base, aes(x = Ohler6, y = log10(exp))) +
  geom_boxplot(aes(color = Ohler6), alpha = 0.5, outlier.shape = NA) +
  facet_grid(sample ~ expression_range) +
  scale_color_manual(values = c("0" = "grey32", "1" = "darkblue")) +
  stat_compare_means(method = "wilcox.test", label = "p.format", comparisons = list(c("0", "1"))) +
  theme_classic() + theme(text = element_text(size = 15), legend.position = "none")

grid.arrange(p_bsize, p_bfreq, p_exp, ncol = 3)
```

```{r explore_ohler1}
df_base <- STAP_annotated %>% mutate(Ohler1 = as.factor(Ohler1))

p_bsize <- ggplot(df_base, aes(x = Ohler1, y = log10(bsize))) +
  geom_boxplot(aes(color = Ohler1), alpha = 0.5, outlier.shape = NA) +
  facet_grid(sample ~ expression_range) +
  scale_color_manual(values = c("0" = "grey32", "1" = "lightblue")) +
  stat_compare_means(method = "wilcox.test", label = "p.format", comparisons = list(c("0", "1"))) +
  theme_classic() + theme(text = element_text(size = 15), legend.position = "none")

p_bfreq <- ggplot(df_base, aes(x = Ohler1, y = log10(bfreq))) +
  geom_boxplot(aes(color = Ohler1), alpha = 0.5, outlier.shape = NA) +
  facet_grid(sample ~ expression_range) +
  scale_color_manual(values = c("0" = "grey32", "1" = "lightblue")) +
  stat_compare_means(method = "wilcox.test", label = "p.format", comparisons = list(c("0", "1"))) +
  theme_classic() + theme(text = element_text(size = 15), legend.position = "none")

p_exp <- ggplot(df_base, aes(x = Ohler1, y = log10(exp))) +
  geom_boxplot(aes(color = Ohler1), alpha = 0.5, outlier.shape = NA) +
  facet_grid(sample ~ expression_range) +
  scale_color_manual(values = c("0" = "grey32", "1" = "lightblue")) +
  stat_compare_means(method = "wilcox.test", label = "p.format", comparisons = list(c("0", "1"))) +
  theme_classic() + theme(text = element_text(size = 15), legend.position = "none")

grid.arrange(p_bsize, p_bfreq, p_exp, ncol = 3)
```
