---
title: "Figure 3 — global burst-parameter trends: promoters vs enhancers"
author: "FL"
date: "`r Sys.Date()`"
output: html_document
---

<!--
================================================================================
FIGURE / PANEL MAP  (which chunk makes which panel)
================================================================================
This script compares how promoters (STAP-seq) and enhancers (STARR-seq) shape
burst parameters (burst size, burst frequency) as a function of mean expression.

FIGURE 3
  3A  Promoter library: burst size vs mean expression (top) and
      burst frequency vs mean expression (bottom); coloured by promoter
      condition (zfh1 / CG4822 / Pld) ...................... chunk "FIGURE 3A"
  3B  Enhancer library: same layout, coloured by enhancer
      condition (DSCP / Rpl35 / Rps15A) ................... chunk "FIGURE 3B"
  3C  Slope bar plots across all 6 samples:
      slope of burst size ~ expression (top) and
      slope of burst frequency ~ expression (bottom),
      annotated with R2 and p-value ....................... chunk "FIGURE 3C"
Supplemental FIGURE 4
  4D viridis mean colored scatter 
  
  Exploratory / QC chunks that are NOT publication panels are kept and clearly
  labelled "EXPLORATORY" (all-samples-overlaid
  fit plots; regression-summary tables). Every fit is generated for BOTH plot
  orientations (expression on x-axis and expression on y-axis) so either can be
  used downstream -- see the `exp_on_x` argument of `make_fit_plot()` /
  `summarize_lm_direction()`.

  NOTE: no plots are written to disk in this version -- every figure is only
  displayed inline (per request). To save a figure, wrap it in the usual
  pdf(...) / print(...) / dev.off() block yourself.

All data comes from the published supplemental tables (Tables 5 and 7), read in
the data-loading chunk below. All former internal cluster paths (data .rds
files, the mu-threshold Excel sheet, the output directory) have been removed.
The mu-threshold merge/recompute is no longer needed because Tables 5 and 7
already ship the quality flags `high_mu` and `KS_pval` used for filtering.
================================================================================
-->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

library(data.table)
library(dplyr)
library(ggplot2)
library(readxl)
library(tibble)
library(scales)
library(forcats)
library(viridis)
```

## Sample definitions and colour scheme

```{r sample-definitions}
# Published sample names (pooled replicates).
# Promoters were assayed by STAP-seq, enhancers by STARR-seq.
samples_promoter <- c(
  "libFL001_STAP_zfh1_pool",
  "libFL002_STAP_CG4822_pool",
  "libFL003_STAP_Pld_pool"
)

samples_enhancer <- c(
  "libFL004_STARR_DSCP_pool",
  "libFL005_STARR_Rpl35_pool",
  "libFL006_STARR_Rps15A_pool"
)

samples_all <- c(samples_promoter, samples_enhancer)

# Colour scheme used for the publication figure:
#   promoters  -> yellow/orange shades
#   enhancers  -> green/teal shades
palette_all <- c(
  "libFL001_STAP_zfh1_pool"    = "#F2994A",
  "libFL002_STAP_CG4822_pool"  = "#F7D154",
  "libFL003_STAP_Pld_pool"     = "#B87400",
  "libFL004_STARR_DSCP_pool"   = "#66C2A5",
  "libFL005_STARR_Rpl35_pool"  = "#009E73",
  "libFL006_STARR_Rps15A_pool" = "#005D5D"
)

palette_promoter <- palette_all[samples_promoter]
palette_enhancer <- palette_all[samples_enhancer]
```

## Loading and merging burst-parameter data

We build one combined data set `dat_base` containing the burst parameters
(burst size `bsize`, burst frequency `bfreq`, mean expression `exp`) for all
samples, then keep only high-quality promoters/enhancers
(`high_mu == TRUE & KS_pval > 0.01`).

```{r load-data}
# ---- Promoters (STAP-seq): published Table 5, sheet "STAP" ----
burst_STAP <- as.data.table(
  read_xlsx("/groups/stark/lorbeer/BARE-seq_data_prep/Tables/Table5_promoter_burst_parameters.xlsx", sheet = "STAP")
)

# ---- Enhancers (STARR-seq): published Table 7, sheet "STARR" ----
burst_STARR <- as.data.table(
  read_xlsx("/groups/stark/lorbeer/BARE-seq_data_prep/Tables/Table7_enhancer_burst_parameters.xlsx", sheet = "STARR")
)

# ---- Combine ----
# Both tables have the same schema (seq_ID, sample, bfreq, bsize, exp,
# log2FoldChange, high_mu, KS_pval), so we stack them directly.
dat <- rbind(burst_STAP, burst_STARR)

# Parse the published sample name into its components:
#   <library>_<assay>_<condition>_<rep>   e.g. libFL001_STAP_zfh1_pool
dat[, c("library", "assay", "condition", "rep") :=
      tstrsplit(sample, "_", fixed = TRUE)]

# Quality filter (Tables 5 and 7 already ship high_mu and KS_pval).
dat_base <- dat %>%
  filter(
    KS_pval > 0.01,
    high_mu == TRUE,
    sample %in% samples_all
  )
```

## EXPLORATORY — mean-coloured burst-size vs burst-frequency scatter

```{r exploratory-mean-color}
mean_color <- dat_base %>%
  filter(sample %in% samples_all) %>%
  mutate(sample = fct_relevel(sample, samples_all)) %>%
  ggplot() +
  geom_point(aes(y = log10(bsize), x = log10(bfreq), color = log10(exp))) +
  scale_color_viridis(
    alpha = 1, begin = 0, end = 1, direction = 1,
    discrete = FALSE, option = "D", aesthetics = "color"
  ) +
  facet_wrap(~sample) +
  theme_classic(base_size = 15)
mean_color
```

## Plotting and regression helper functions

`make_fit_plot()` and `summarize_lm_direction()` both take `exp_on_x`, so the
relationship can be plotted / modelled with expression on either axis.

```{r helper-functions}
# Keep only rows usable on a log10 scale for the requested parameter, and lock
# the sample factor order so colours/legends stay consistent.
prep_plot_data <- function(df, param, sample_levels) {
  df %>%
    filter(sample %in% sample_levels, exp > 0, .data[[param]] > 0) %>%
    mutate(sample = factor(sample, levels = sample_levels))
}

# Scatter + linear fit of a burst parameter against mean expression.
#   exp_on_x = TRUE  -> x = log10(exp),   y = log10(param)
#   exp_on_x = FALSE -> x = log10(param), y = log10(exp)
make_fit_plot <- function(df, param, exp_on_x = FALSE, show_points = FALSE,
                          sample_levels, palette_vals,
                          xlim = NULL, ylim = NULL, title_text = NULL) {

  df_plot <- prep_plot_data(df, param, sample_levels)

  # Guard: if none of the requested samples are present (e.g. enhancer data
  # not supplied), return an informative placeholder instead of erroring.
  if (nrow(df_plot) == 0) {
    return(
      ggplot() + theme_void() +
        annotate("text", x = 0, y = 0,
                 label = "No data for the selected samples\n(see MISSING INPUT in header)")
    )
  }

  if (exp_on_x) {
    p <- ggplot(df_plot, aes(x = log10(exp), y = log10(.data[[param]])))
  } else {
    p <- ggplot(df_plot, aes(x = log10(.data[[param]]), y = log10(exp)))
  }

  if (show_points) {
    p <- p + geom_point(aes(color = sample), alpha = 0.25)
  }

  p +
    geom_smooth(
      aes(color = sample, fill = sample),
      method = "lm", formula = y ~ x, se = TRUE, alpha = 0.2
    ) +
    scale_color_manual(values = palette_vals) +
    scale_fill_manual(values = palette_vals) +
    coord_cartesian(xlim = xlim, ylim = ylim) +
    theme_classic(base_size = 15) +
    theme(legend.position = "none") +
    labs(
      title = title_text,
      x = if (exp_on_x) "log10(mean expression)" else paste0("log10(", param, ")"),
      y = if (exp_on_x) paste0("log10(", param, ")") else "log10(mean expression)"
    )
}

# Per-sample linear-model summary (slope, SE, p, R2, ...) in either direction.
summarize_lm_direction <- function(df, param, exp_on_x = TRUE, sample_levels) {
  df_model <- prep_plot_data(df, param, sample_levels) %>%
    mutate(
      x = if (exp_on_x) log10(exp) else log10(.data[[param]]),
      y = if (exp_on_x) log10(.data[[param]]) else log10(exp)
    ) %>%
    filter(is.finite(x), is.finite(y))

  df_model %>%
    group_by(sample) %>%
    do({
      m <- lm(y ~ x, data = .)
      tibble(
        intercept    = coef(m)[["(Intercept)"]],
        slope        = coef(m)[["x"]],
        intercept_se = coef(summary(m))[1, "Std. Error"],
        slope_se     = coef(summary(m))[2, "Std. Error"],
        slope_p      = coef(summary(m))[2, "Pr(>|t|)"],
        r2           = summary(m)$r.squared,
        adj_r2       = summary(m)$adj.r.squared,
        sigma        = summary(m)$sigma,
        nobs         = length(m$fitted.values)
      )
    }) %>%
    ungroup() %>%
    mutate(sample = factor(sample, levels = sample_levels)) %>%
    arrange(sample)
}

# Combined "R2=.., p=.." annotation used on the slope bar plots.
fmt_label <- function(p, r2) {
  p_str <- dplyr::case_when(
    is.na(p) ~ "p = NA",
    p < 1e-4 ~ "p < 0.0001",
    p < 1e-3 ~ "p < 0.001",
    TRUE     ~ paste0("p = ", scales::number(p, accuracy = 0.001))
  )
  sprintf("R2=%.2f, %s", r2, p_str)
}

# Bar plot of per-sample regression slopes +/- SE, annotated with R2 and p.
plot_slopes <- function(stats_df, palette_vals, title_text, y_label) {
  if (nrow(stats_df) == 0) {
    return(
      ggplot() + theme_void() +
        annotate("text", x = 0, y = 0,
                 label = "No data for the selected samples\n(see MISSING INPUT in header)")
    )
  }
  ymax <- max(stats_df$slope + stats_df$slope_se, na.rm = TRUE)
  ymin <- min(stats_df$slope - stats_df$slope_se, na.rm = TRUE)
  label_offset <- 0.05 * max(ymax - ymin, 0.1)

  ggplot(stats_df, aes(x = sample, y = slope, fill = sample)) +
    geom_col(width = 0.8) +
    geom_errorbar(aes(ymin = slope - slope_se, ymax = slope + slope_se), width = 0.2) +
    geom_text(
      aes(y = slope + slope_se + label_offset, label = fmt_label(slope_p, r2)),
      size = 3.5, vjust = 0
    ) +
    scale_fill_manual(values = palette_vals, guide = "none") +
    theme_classic(base_size = 15) +
    theme(axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1)) +
    labs(title = title_text, x = "Sample", y = y_label)
}
```

## EXPLORATORY — regression summary tables (both orientations)

```{r exploratory-lm-tables}
bfreq_on_exp_stats <- summarize_lm_direction(dat_base, "bfreq", exp_on_x = TRUE,  sample_levels = samples_all)
bsize_on_exp_stats <- summarize_lm_direction(dat_base, "bsize", exp_on_x = TRUE,  sample_levels = samples_all)
exp_on_bfreq_stats <- summarize_lm_direction(dat_base, "bfreq", exp_on_x = FALSE, sample_levels = samples_all)
exp_on_bsize_stats <- summarize_lm_direction(dat_base, "bsize", exp_on_x = FALSE, sample_levels = samples_all)

print("bfreq ~ exp"); bfreq_on_exp_stats
print("bsize ~ exp"); bsize_on_exp_stats
print("exp ~ bfreq"); exp_on_bfreq_stats
print("exp ~ bsize"); exp_on_bsize_stats
```

## EXPLORATORY — all samples overlaid (promoters + enhancers together)

```{r exploratory-all-overlaid}
all_bfreq_expY_nopoints <- make_fit_plot(
  dat_base, "bfreq", exp_on_x = FALSE, show_points = FALSE,
  sample_levels = samples_all, palette_vals = palette_all,
  xlim = c(-2, 0.5), ylim = c(-1.5, 1.25),
  title_text = "All samples: log10(exp) vs log10(bfreq)"
)
all_bfreq_expX_points <- make_fit_plot(
  dat_base, "bfreq", exp_on_x = TRUE, show_points = TRUE,
  sample_levels = samples_all, palette_vals = palette_all,
  xlim = c(-1.5, 1.25), ylim = c(-2, 0.5),
  title_text = "All samples: log10(bfreq) vs log10(exp) + points"
)
all_bsize_expX_points <- make_fit_plot(
  dat_base, "bsize", exp_on_x = TRUE, show_points = TRUE,
  sample_levels = samples_all, palette_vals = palette_all,
  xlim = c(-1.25, 1.25), ylim = c(-1.5, 1.5),
  title_text = "All samples: log10(bsize) vs log10(exp) + points"
)

all_bfreq_expY_nopoints
all_bfreq_expX_points
all_bsize_expX_points
```

## FIGURE 3A — promoter library (burst size / frequency vs expression)

```{r FIGURE 3A}
promoter_bsize_expX_points <- make_fit_plot(
  dat_base, "bsize", exp_on_x = TRUE, show_points = TRUE,
  sample_levels = samples_promoter, palette_vals = palette_promoter,
  xlim = c(-1.25, 1.25), ylim = c(-1.5, 1.5),
  title_text = "Promoters: log10(bsize) vs log10(exp)"
)

promoter_bfreq_expX_points <- make_fit_plot(
  dat_base, "bfreq", exp_on_x = TRUE, show_points = TRUE,
  sample_levels = samples_promoter, palette_vals = palette_promoter,
  xlim = c(-1.5, 1.25), ylim = c(-2, 0.5),
  title_text = "Promoters: log10(bfreq) vs log10(exp)"
)

# Panel 3A (top = burst size, bottom = burst frequency)
promoter_bsize_expX_points
promoter_bfreq_expX_points
```

## FIGURE 3B — enhancer library (burst size / frequency vs expression)

```{r FIGURE 3B}
enhancer_bsize_expX_points <- make_fit_plot(
  dat_base, "bsize", exp_on_x = TRUE, show_points = TRUE,
  sample_levels = samples_enhancer, palette_vals = palette_enhancer,
  xlim = c(-1.25, 1.25), ylim = c(-1.5, 1.5),
  title_text = "Enhancers: log10(bsize) vs log10(exp)"
)

enhancer_bfreq_expX_points <- make_fit_plot(
  dat_base, "bfreq", exp_on_x = TRUE, show_points = TRUE,
  sample_levels = samples_enhancer, palette_vals = palette_enhancer,
  xlim = c(-1.5, 1.25), ylim = c(-2, 0.5),
  title_text = "Enhancers: log10(bfreq) vs log10(exp)"
)

# Panel 3B (top = burst size, bottom = burst frequency)
enhancer_bsize_expX_points
enhancer_bfreq_expX_points
```

## FIGURE 3C — regression slopes across all samples

```{r FIGURE 3C}
bsize_stats_expX_all <- summarize_lm_direction(
  dat_base, "bsize", exp_on_x = TRUE, sample_levels = samples_all
)
bfreq_stats_expX_all <- summarize_lm_direction(
  dat_base, "bfreq", exp_on_x = TRUE, sample_levels = samples_all
)

slope_bsize_expX_all <- plot_slopes(
  bsize_stats_expX_all, palette_vals = palette_all,
  title_text = "Slope of log10(bsize) ~ log10(exp)",
  y_label = "Slope: burst size on expression"
)
slope_bfreq_expX_all <- plot_slopes(
  bfreq_stats_expX_all, palette_vals = palette_all,
  title_text = "Slope of log10(bfreq) ~ log10(exp)",
  y_label = "Slope: burst frequency on expression"
)

# Panel 3C (top = slope of burst size, bottom = slope of burst frequency)
slope_bsize_expX_all
slope_bfreq_expX_all
```
