---
title: "Figure 4 & Supplemental Figure 4 — enhancer motif analysis and BARe-seq validation"
author: "FL"
date: "`r Sys.Date()`"
output: html_document
---

<!--
================================================================================
FIGURE / PANEL MAP  (which chunk makes which panel)
================================================================================
This combined script produces Figure 4 (enhancer motif regression) and
Supplemental Figure 4 (correlation heatmaps + BARe-seq validation).

FIGURE 4
  4A  Enhancer library + motif-frequency donut (ETS, GATA, AP1, SREBP, Trl,
      twist, Ohler6, Ohler7, Ohler1, Dref) ................. chunk "FIGURE 4A"
  4B  Regression-coefficient heatmap (motif x promoter), pooled samples,
      multivariate model, for burst size / frequency / mean
      expression .......................................... chunk "FIGURE 4B"
  4C  Illustration of the per-motif effect analysis ........ (no code)
  4D  Per-motif burst-size delta scatter + boxplots ........ (NOT in this script)
  4E  Per-motif burst-frequency delta scatter .............. (NOT in this script)

SUPPLEMENTAL FIGURE 4
  S4A Log2 promoter-responsiveness scatter ................. (different dataset, excluded)
  S4B BARe-seq vs STARR-seq log2FC scatter (3 promoters) .. chunk "SUPPL. FIG. 4B"
  S4C Mean expression histograms with mu threshold ......... chunk "SUPPL. FIG. 4C"
  S4D Viridis burst-size vs frequency, log10(exp) .......... (in Fig3_promoter_enhancer_clean.Rmd)
  S4E Replicate-level correlation heatmaps ................. chunk "SUPPL. FIG. 4E"
  S4F Pooled-sample correlation heatmaps ................... chunk "SUPPL. FIG. 4F"

  Additional exploratory chunks (NOT publication panels) are labelled "EXPLORATORY":
  per-motif-instance boxplots, presence/absence median heatmaps, meta-analysis scatter.

  NOTE: no plots are written to disk in this version -- every figure is only
  displayed inline. To save a figure, wrap it in a pdf(...) / print(...) /
  dev.off() block writing to a relative directory such as ./output/Fig4.

================================================================================
DATA SOURCES
================================================================================
All inputs are the published supplemental tables -- no cluster files needed.

  * Enhancer burst parameters : Tables/Table7_enhancer_burst_parameters.xlsx
                                (sheet "STARR"): seq_ID, sample, bfreq, bsize,
                                exp, log2FoldChange, high_mu, KS_pval
  * Enhancer motif counts +   : Tables/Table6_enhancer_library_annotations.xlsx
    annotations + BARe-seq      (sheet "STARR"): per-oligo integer counts for
    log2FC                      ETS, GATA, AP1, SREBP, Trl, twist, Ohler6,
                                Ohler7, Ohler1, Dref (+ enhancer_group, class,
                                dev_log2FoldChange, hk_log2FoldChange for
                                BARe-seq validation)
  * Sample name mapping +     : Tables/Table3_mu_thresholds.xlsx
    mu thresholds               (new_sample_name, sample, mu_threshold)
================================================================================
-->

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE
)

require(data.table)
require(ggplot2)
require(broom)
require(purrr)
require(metafor)
library(dplyr)
library(tidyr)
require(readxl)
#require(tibble)
require(ggrepel)
#require(scales)
```

## Sample definitions

```{r sample-definitions}
# Published enhancer (STARR-seq) sample names.
pooled_samples <- c(
  "libFL004_STARR_DSCP_pool",
  "libFL005_STARR_Rpl35_pool",
  "libFL006_STARR_Rps15A_pool"
)

replicate_samples <- c(
  "libFL004_STARR_DSCP_replicate1", "libFL004_STARR_DSCP_replicate2",
  "libFL005_STARR_Rpl35_replicate1", "libFL005_STARR_Rpl35_replicate2",
  "libFL006_STARR_Rps15A_replicate1", "libFL006_STARR_Rps15A_replicate2"
)

# All samples analysed (pools + replicates).
sample_names <- c(pooled_samples, replicate_samples)

# Motif columns used throughout (integer counts in Table 6).
cols <- c("ETS", "GATA", "AP1", "SREBP", "Trl", "twist",
          "Ohler6", "Ohler7", "Ohler1", "Dref")
```

## Load and merge data

Enhancer burst parameters (Table 7) are joined to the per-oligo motif counts and
annotations (Table 6) on `seq_ID`, then split per sample and filtered to
high-quality oligos (`high_mu == TRUE & KS_pval > 0.01`).

```{r load-data}
# Load Table 7: STARR burst parameters
burst <- as.data.table(
  read_xlsx("/groups/stark/lorbeer/BARE-seq_data_prep/Tables/Table7_enhancer_burst_parameters.xlsx", sheet = "STARR")
)

# Load Table 6: enhancer annotations + motif counts + BARe-seq log2FC
anno <- as.data.table(
  read_xlsx("/groups/stark/lorbeer/BARE-seq_data_prep/Tables/Table6_enhancer_library_annotations.xlsx", sheet = "STARR")
)

# Load Table 3: mu thresholds and sample name mapping
mu_thresholds_raw <- read_excel("/groups/stark/lorbeer/BARE-seq_data_prep/Tables/Table3_mu_thresholds.xlsx")
mu_thresholds <- as.data.table(mu_thresholds_raw)

# Keep motif counts + a few annotation columns; join on seq_ID.
anno_keep <- unique(anno[, c("seq_ID", "enhancer_group", "class", cols), with = FALSE])
param <- merge(burst, anno_keep, by = "seq_ID", all.x = TRUE)

# Merge mu thresholds into param (using new_sample_name)
mu_map <- mu_thresholds[, .(sample = sample, mu_threshold)]
param <- merge(param, mu_map, by = "sample", all.x = TRUE)

# Coerce high_mu to logical (ships as text "TRUE"/"FALSE" in supplemental table)
param[, high_mu := (high_mu == "TRUE" | high_mu == TRUE)]

# Parse the published sample name: <library>_<assay>_<condition>_<rep>
param[, c("library", "assay", "condition", "rep") :=
        tstrsplit(sample, "_", fixed = TRUE)]

head(param)

# Read into a per-sample list and apply the quality filter.
param_list <- list()
for (s in sample_names) {
  param_list[[s]] <- as.data.frame(param[sample == s & (high_mu & KS_pval > 0.01)])
}
```

## Build BARe-seq validation table (for Suppl. Fig. 4B)

Merge BARe-seq log2FC (from anno) with STARR-seq log2FC (from param) to compare
the two assays.

```{r build-bareseq-validation}
# Extract pooled samples log2FC from param and reshape wide
log2fc_wide <- param[sample %in% pooled_samples, .(seq_ID, sample, log2FoldChange)]
log2fc_wide <- dcast(log2fc_wide, seq_ID ~ sample, value.var = "log2FoldChange")
setnames(log2fc_wide, 
         old = c("libFL004_STARR_DSCP_pool", "libFL005_STARR_Rpl35_pool", "libFL006_STARR_Rps15A_pool"),
         new = c("log2FoldChange_DSCP", "log2FoldChange_Rpl35", "log2FoldChange_Rps15A"))

# Merge with anno to get BARe-seq log2FC
diff_exp <- merge(anno[, .(seq_ID, ID, Group, enhancer_group, dev_log2FoldChange, hk_log2FoldChange)], 
                  log2fc_wide, 
                  by = "seq_ID", 
                  all.x = TRUE)

# For backward compatibility with original variable names in plots:
diff_exp[, log2FoldChange_rg := log2FoldChange_DSCP]
diff_exp[, log2FoldChange_p309 := log2FoldChange_Rpl35]
diff_exp[, log2FoldChange_p319 := log2FoldChange_Rps15A]

head(diff_exp)
```

## Helper functions for motif regression

Motif count predictors are upper-clipped before modelling: the clipping
threshold is the 95th percentile within each motif column, constrained to lie
between 2 and 3. Values above the threshold are replaced by the threshold to
reduce leverage from rare high-count observations while retaining information.

```{r helper-functions}
# Generate clipped motif counts (few observations at the high end).
clip_motifs <- function(dat, cols,
                        q = 0.95,
                        min_cutoff = 2,
                        max_cutoff = 3,
                        suffix = "_clip") {
  dat <- as.data.table(dat)  # safe if it's already a data.table
  out_cols <- paste0(cols, suffix)

  dat[, (out_cols) := lapply(.SD, function(x) {
    cutoff <- quantile(x, q, na.rm = TRUE)
    cutoff <- fcase(
      cutoff < min_cutoff, min_cutoff,
      cutoff > max_cutoff, max_cutoff,
      default = cutoff
    )
    fifelse(x > cutoff, cutoff, x)
  }), .SDcols = cols]

  dat
}

param_list <- lapply(param_list, clip_motifs, cols = cols)
cols_clip <- paste0(cols, "_clip")

sig_stars <- function(p) {
  case_when(
    is.na(p)  ~ NA_character_,
    p < 0.001 ~ "***",
    p < 0.01  ~ "**",
    p < 0.05  ~ "*",
    TRUE      ~ "ns"
  )
}

na_block_for_motifs <- function(motifs, outcome, model_label, n_used = NA_integer_) {
  tibble(
    Motif = motifs,
    Outcome = outcome,
    Model = model_label,
    n_used = n_used,
    n_present = NA_integer_,
    prop_present = NA_real_,
    estimate = NA_real_,
    std.error = NA_real_,
    statistic = NA_real_,
    p.value = NA_real_,
    FoldChange = NA_real_
  )
}
```

## EXPLORATORY — burst parameters by motif-instance count

```{r exploratory-motif-boxplots}
df <- param_list$libFL004_STARR_DSCP_pool

plot_df <- df %>%
  as.data.frame() %>%
  pivot_longer(
    cols = ends_with("_clip"),
    names_to = "clipped_motif",
    values_to = "n_clipped_instances"
  ) %>%
  pivot_longer(
    cols = c(bsize, bfreq, exp),
    names_to = "parameter",
    values_to = "value"
  ) %>%
  mutate(
    clipped_motif = sub("_clip$", "", clipped_motif),
    n_clipped_instances = factor(n_clipped_instances)
  )

desired_order <- cols
plot_df <- plot_df %>%
  mutate(
    clipped_motif = factor(clipped_motif, levels = desired_order),
    parameter = factor(parameter, levels = c("bsize", "bfreq", "exp"))
  )

p_motifclip <- ggplot(plot_df, aes(x = n_clipped_instances, y = log10(value))) +
  geom_boxplot(outlier.shape = NA) +
  facet_grid(clipped_motif ~ parameter, scales = "free_y") +
  labs(
    x = "Number of motif instances",
    y = "parameter (log10)",
    title = "bsize, bfreq, and exp by motif instance"
  ) +
  theme_classic(base_size = 15) +
  theme(
    strip.background = element_rect(fill = "grey90"),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
p_motifclip
```

## Linear-model functions

In the enhancer analysis clipped motif counts are used as numeric predictors;
regression coefficients represent the expected change in log10(outcome) per
one-unit increase in clipped motif count.

```{r lm-functions}
fit_motif_model_one_df <- function(data, outcome, motifs, include_exp = FALSE) {
  model_label <- ifelse(include_exp, "motifs_plus_log10exp", "motifs_only")

  motifs_present <- intersect(motifs, colnames(data))
  motifs_missing <- setdiff(motifs, motifs_present)

  if (!outcome %in% colnames(data)) {
    return(na_block_for_motifs(motifs, outcome, model_label, n_used = NA_integer_))
  }
  if (include_exp && !"exp" %in% colnames(data)) {
    return(na_block_for_motifs(motifs, outcome, model_label, n_used = NA_integer_))
  }

  fitted_present <- purrr::map_dfr(motifs_present, function(m) {
    rhs <- m
    if (include_exp) rhs <- paste(rhs, "+ log10(exp)")
    fml <- as.formula(paste0("log10(", outcome, ") ~ ", rhs))

    mf <- tryCatch(model.frame(fml, data = data), error = function(e) NULL)
    n_used <- if (is.null(mf)) NA_integer_ else nrow(mf)

    if (is.null(mf) || n_used == 0) {
      return(
        na_block_for_motifs(m, outcome, model_label, n_used = ifelse(is.null(mf), NA_integer_, 0L)) %>%
          dplyr::mutate(
            n_present = ifelse(is.null(mf), NA_integer_, sum(mf[[m]] > 0, na.rm = TRUE)),
            prop_present = ifelse(is.null(mf), NA_real_, n_present / n_used)
          )
      )
    }

    n_present <- sum(mf[[m]] == 1, na.rm = TRUE)
    prop_present <- n_present / n_used

    model <- tryCatch(lm(fml, data = data), error = function(e) NULL)
    if (is.null(model)) {
      return(
        na_block_for_motifs(m, outcome, model_label, n_used = n_used) %>%
          dplyr::mutate(n_present = n_present, prop_present = prop_present)
      )
    }

    broom::tidy(model) %>%
      dplyr::filter(term == m) %>%
      dplyr::transmute(
        Motif = m,
        Outcome = outcome,
        Model = model_label,
        n_used = n_used,
        n_present = n_present,
        prop_present = prop_present,
        estimate = estimate,
        std.error = std.error,
        statistic = statistic,
        p.value = p.value,
        FoldChange = 10^estimate
      )
  })

  missing_part <- na_block_for_motifs(motifs_missing, outcome, model_label, n_used = NA_integer_) %>%
    dplyr::mutate(n_present = NA_integer_, prop_present = NA_real_)

  dplyr::bind_rows(fitted_present, missing_part) %>%
    dplyr::mutate(Motif = factor(Motif, levels = motifs)) %>%
    dplyr::arrange(Motif) %>%
    dplyr::mutate(Motif = as.character(Motif))
}

fit_motif_model_multiv_df <- function(data, outcome, motifs, include_exp = FALSE) {
  model_label <- ifelse(include_exp, "motifs_plus_log10exp", "motifs_only")

  motifs_present <- intersect(motifs, colnames(data))
  motifs_missing <- setdiff(motifs, motifs_present)

  if (!outcome %in% colnames(data)) {
    return(na_block_for_motifs(motifs, outcome, model_label, n_used = NA_integer_))
  }
  if (include_exp && !"exp" %in% colnames(data)) {
    return(na_block_for_motifs(motifs, outcome, model_label, n_used = NA_integer_))
  }
  if (length(motifs_present) == 0) {
    return(na_block_for_motifs(motifs, outcome, model_label, n_used = NA_integer_))
  }

  rhs <- paste(motifs_present, collapse = " + ")
  if (include_exp) rhs <- paste(rhs, "+ log10(exp)")
  fml <- as.formula(paste0("log10(", outcome, ") ~ ", rhs))

  mf <- tryCatch(model.frame(fml, data = data), error = function(e) NULL)
  n_used <- if (is.null(mf)) NA_integer_ else nrow(mf)

  if (is.null(mf) || n_used == 0) {
    out <- na_block_for_motifs(motifs_present, outcome, model_label, n_used = ifelse(is.null(mf), NA_integer_, 0L))
    return(dplyr::bind_rows(
      out,
      na_block_for_motifs(motifs_missing, outcome, model_label, n_used = NA_integer_)
    ))
  }

  presence <- tibble::tibble(
    Motif = motifs_present,
    n_present = purrr::map_int(motifs_present, ~ sum(mf[[.x]] == 1, na.rm = TRUE)),
    prop_present = n_present / n_used
  )

  model <- tryCatch(lm(fml, data = data), error = function(e) NULL)
  if (is.null(model)) {
    out <- na_block_for_motifs(motifs_present, outcome, model_label, n_used = n_used)
    out <- dplyr::left_join(out, presence, by = "Motif")
    return(dplyr::bind_rows(
      out,
      na_block_for_motifs(motifs_missing, outcome, model_label, n_used = NA_integer_)
    ))
  }

  coefs <- broom::tidy(model) %>%
    dplyr::filter(term %in% motifs_present) %>%
    dplyr::transmute(
      Motif = term,
      Outcome = outcome,
      Model = model_label,
      n_used = n_used,
      estimate = estimate,
      std.error = std.error,
      statistic = statistic,
      p.value = p.value,
      FoldChange = 10^estimate
    ) %>%
    dplyr::left_join(presence, by = "Motif")

  missing_part <- na_block_for_motifs(motifs_missing, outcome, model_label, n_used = NA_integer_) %>%
    dplyr::mutate(n_present = NA_integer_, prop_present = NA_real_)

  dplyr::bind_rows(coefs, missing_part) %>%
    dplyr::mutate(Motif = factor(Motif, levels = motifs)) %>%
    dplyr::arrange(Motif) %>%
    dplyr::mutate(Motif = as.character(Motif))
}
```

## Run the linear models (univariate + multivariate, with / without expression)

Meaning of coefficients: association with clipped motif count -- a dose/count
effect.

```{r run-lm}
motifs_10 <- cols_clip

fit_motif_models_list_both <- function(data_list, outcomes, motifs, include_exp = FALSE) {
  purrr::imap_dfr(data_list, function(df, nm) {
    purrr::map_dfr(outcomes, function(outcome) {
      multiv <- fit_motif_model_multiv_df(
        data = df, outcome = outcome, motifs = motifs, include_exp = include_exp
      ) %>% dplyr::mutate(Analysis = "multivariate")

      univ <- fit_motif_model_one_df(
        data = df, outcome = outcome, motifs = motifs, include_exp = include_exp
      ) %>% dplyr::mutate(Analysis = "univariate")

      dplyr::bind_rows(multiv, univ) %>%
        dplyr::mutate(
          sample_name = nm,
          sig = sig_stars(p.value)
        )
    })
  }) %>%
    dplyr::select(sample_name, Outcome, Analysis, Model, Motif, sig, dplyr::everything())
}

outcomes <- c("bsize", "bfreq", "exp")

results_both <- dplyr::bind_rows(
  fit_motif_models_list_both(param_list, outcomes, motifs_10, include_exp = FALSE),
  fit_motif_models_list_both(param_list, outcomes, motifs_10, include_exp = TRUE)
)

print(results_both)
```

## Heatmap function (regression coefficients)

```{r heatmap-function}
sym_limits <- function(x, midpoint = 0) {
  x <- x[is.finite(x)]
  if (length(x) == 0) return(c(-1, 1))
  max_abs <- max(abs(x - midpoint))
  c(midpoint - max_abs, midpoint + max_abs)
}

make_heatmap_model <- function(results, model_name,
                              motifs_order = NULL,
                              samples_order = NULL,
                              min_occurrence = 0,
                              star_size = 3,
                              analysis = NULL) {

  plot_df <- results %>%
    dplyr::filter(Model == model_name, Outcome %in% c("bsize", "bfreq", "exp"))

  analysis_label <- if (is.null(analysis)) "both (univariate + multivariable)" else analysis

  if (!is.null(analysis)) {
    plot_df <- plot_df %>% dplyr::filter(Analysis == analysis)
  }

  if (!is.null(samples_order)) {
    plot_df <- plot_df %>%
      dplyr::filter(sample_name %in% samples_order) %>%
      dplyr::mutate(sample_name = factor(sample_name, levels = samples_order))
  } else {
    plot_df <- plot_df %>% dplyr::mutate(sample_name = factor(sample_name))
  }

  plot_df <- plot_df %>%
    dplyr::mutate(
      value = estimate,
      value = ifelse(!is.na(n_present) & n_present < min_occurrence, NA_real_, value),
      Sig_raw_plot = ifelse(sig %in% c("*", "**", "***"), sig, "")
    )

  if (!is.null(motifs_order)) {
    plot_df <- plot_df %>% dplyr::mutate(Motif = factor(Motif, levels = motifs_order))
  } else {
    plot_df <- plot_df %>% dplyr::mutate(Motif = factor(Motif))
  }

  lims <- sym_limits(plot_df$value, midpoint = 0)

  ggplot2::ggplot(plot_df, ggplot2::aes(x = sample_name, y = Motif)) +
    ggplot2::geom_tile(ggplot2::aes(fill = value)) +
    ggplot2::geom_text(ggplot2::aes(label = Sig_raw_plot), size = star_size, na.rm = TRUE) +
    ggplot2::facet_grid(. ~ Outcome) +
    ggplot2::scale_fill_gradient2(
      low = "blue", mid = "white", high = "red",
      midpoint = 0,
      limits = lims,
      na.value = "grey80"
    ) +
    ggplot2::labs(
      title = paste0("Motif effects (", analysis_label, "): ", model_name),
      subtitle = paste0(
        "Stars = raw p-value; tiles shown only if n_present >= ", min_occurrence,
        if (!is.null(samples_order)) paste0("; samples ordered (n=", length(samples_order), ")") else ""
      ),
      x = "Sample",
      y = "Motif",
      fill = "estimate"
    ) +
    ggplot2::theme_minimal(base_size = 11) +
    ggplot2::theme(
      panel.grid = ggplot2::element_blank(),
      axis.text.x = ggplot2::element_text(angle = 45, hjust = 1, vjust = 1),
      strip.text = ggplot2::element_text(face = "bold")
    )
}
```

## FIGURE 4B — regression-coefficient heatmaps

Pooled samples (multivariate = the main Fig. 4B panel; univariate shown for
comparison). The all-samples / replicate version is used as a supplemental
panel.

```{r FIGURE 4B}
min_occurrence <- 5
motifs_order <- rev(motifs_10)

# ---- Pooled samples (Figure 4B) ----
q_multiv <- make_heatmap_model(results_both, "motifs_only",
                              motifs_order = motifs_order,
                              samples_order = pooled_samples,
                              min_occurrence = min_occurrence,
                              analysis = "multivariate")

q_univ <- make_heatmap_model(results_both, "motifs_only",
                            motifs_order = motifs_order,
                            samples_order = pooled_samples,
                            min_occurrence = min_occurrence,
                            analysis = "univariate")

q_multiv   # <- Figure 4B (multivariate, pooled)
q_univ

# ---- All (replicate-level) samples for the supplement ----
p_multiv <- make_heatmap_model(results_both, "motifs_only",
                              motifs_order = motifs_order,
                              samples_order = replicate_samples,
                              min_occurrence = min_occurrence,
                              analysis = "multivariate")

p_univ <- make_heatmap_model(results_both, "motifs_only",
                            motifs_order = motifs_order,
                            samples_order = replicate_samples,
                            min_occurrence = min_occurrence,
                            analysis = "univariate")

p_multiv
p_univ
```

## EXPLORATORY — presence/absence median (log2FC / Wilcoxon) analysis

Binary motif presence/absence: median difference and Wilcoxon test per motif.

```{r exploratory-median-function}
analyze_param_list_log2FC <- function(param_list,
                                      motifs_order,
                                      samples_order = NULL) {
  stopifnot(is.list(param_list))

  if (is.null(names(param_list)) || any(names(param_list) == "")) {
    names(param_list) <- paste0("sample_", seq_along(param_list))
  }

  if (!is.null(samples_order)) {
    keep <- intersect(samples_order, names(param_list))
    param_list <- param_list[keep]
    param_list <- param_list[samples_order[samples_order %in% names(param_list)]]
  }

  add_significance <- function(p) {
    dplyr::case_when(
      is.na(p)  ~ "",
      p < 0.001 ~ "***",
      p < 0.01  ~ "**",
      p < 0.05  ~ "*",
      TRUE      ~ ""
    )
  }

  sym_limits <- function(x, midpoint = 0) {
    x <- x[is.finite(x)]
    if (length(x) == 0) return(c(-1, 1))
    max_abs <- max(abs(x - midpoint), na.rm = TRUE)
    c(midpoint - max_abs, midpoint + max_abs)
  }

  results_df <- purrr::imap_dfr(param_list, function(df, sample_name) {
    df <- data.table::as.data.table(df)

    purrr::map_dfr(motifs_order, function(m) {
      if (!m %in% names(df)) {
        return(tibble::tibble(
          sample_name = sample_name,
          motif = m,
          measurement = c("bfreq", "bsize", "exp"),
          value_0 = NA_real_, value_1 = NA_real_, diff = NA_real_, log2FC = NA_real_,
          pvalue = NA_real_, Nobservations_0 = NA_integer_, Nobservations_1 = NA_integer_
        ))
      }

      dt_m <- df[, .(bfreq, bsize, exp, motif = get(m))]

      n0 <- sum(dt_m$motif == 0, na.rm = TRUE)
      n1 <- sum(dt_m$motif > 0, na.rm = TRUE)

      one_measure <- function(y) {
        med0 <- dt_m[motif == 0, median(get(y), na.rm = TRUE)]
        med1 <- dt_m[motif > 0, median(get(y), na.rm = TRUE)]
        wt <- suppressWarnings(stats::wilcox.test(get(y) ~ (motif > 0), data = dt_m))

        tibble::tibble(
          sample_name = sample_name,
          motif = m,
          measurement = y,
          value_0 = med0,
          value_1 = med1,
          diff = med1 - med0,
          log2FC = log2(med1 / med0),
          pvalue = wt$p.value,
          Nobservations_0 = n0,
          Nobservations_1 = n1
        )
      }

      dplyr::bind_rows(one_measure("bfreq"), one_measure("bsize"), one_measure("exp"))
    })
  }) %>%
    dplyr::mutate(
      significance = add_significance(pvalue),
      motif = factor(motif, levels = motifs_order),
      sample_name = factor(sample_name, levels = names(param_list)),
      measurement = dplyr::recode(
        measurement,
        bfreq = "Burst Frequency",
        bsize = "Burst Size",
        exp   = "Mean expression"
      )
    )

  lims <- sym_limits(results_df$log2FC, midpoint = 0)

  p <- ggplot2::ggplot(results_df, ggplot2::aes(x = sample_name, y = motif, fill = log2FC)) +
    ggplot2::geom_tile() +
    ggplot2::geom_text(ggplot2::aes(label = significance), size = 3) +
    ggplot2::facet_grid(. ~ measurement) +
    ggplot2::scale_fill_gradient2(
      limits = lims,
      low = "blue", mid = "white", high = "red",
      midpoint = 0,
      oob = scales::squish,
      na.value = "grey90",
      name = "log2FC"
    ) +
    ggplot2::theme_minimal() +
    ggplot2::theme(
      panel.grid = ggplot2::element_blank(),
      axis.text.x = ggplot2::element_text(angle = 45, hjust = 1),
      axis.title.x = ggplot2::element_blank(),
      axis.title.y = ggplot2::element_blank(),
      strip.text = ggplot2::element_text(face = "bold")
    )

  list(table = results_df, plot = p)
}
```

```{r exploratory-median-heatmaps}
# pooled samples only
res_pooled <- analyze_param_list_log2FC(
  param_list,
  motifs_order = motifs_10,
  samples_order = pooled_samples
)
print(res_pooled$plot)
pooled_table <- res_pooled$table

# all samples
res_all <- analyze_param_list_log2FC(param_list, motifs_order = motifs_10)
print(res_all$plot)
all_table <- res_all$table
```

## EXPLORATORY — meta-analysis across pooled samples

Fixed-effect meta-analysis of the multivariate motifs_only coefficients across
the three pooled samples (motif must contribute in >= min_k samples).

```{r exploratory-meta}
min_k <- 3

meta_input <- results_both %>%
  filter(sample_name %in% pooled_samples & Analysis == "multivariate" & Model == "motifs_only") %>%
  filter(
    is.finite(estimate),
    is.finite(std.error),
    std.error > 0,
    !is.na(sample_name)
  )

combos <- meta_input %>%
  distinct(sample_name, Motif, Outcome) %>%
  count(Motif, Outcome, name = "k") %>%
  filter(k >= min_k)

pooled_list <- vector("list", nrow(combos))

for (i in seq_len(nrow(combos))) {
  subset_data <- meta_input %>%
    filter(Motif == combos$Motif[i], Outcome == combos$Outcome[i]) %>%
    distinct(sample_name, .keep_all = TRUE)

  meta_obj <- rma(
    yi = subset_data$estimate,
    sei = subset_data$std.error,
    method = "FE"
  )

  pooled_list[[i]] <- data.frame(
    Motif = combos$Motif[i],
    Outcome = combos$Outcome[i],
    k = nrow(subset_data),
    pooled_estimate = as.numeric(meta_obj$beta[1]),
    pooled_SE = as.numeric(meta_obj$se),
    pooled_pValue = as.numeric(meta_obj$pval),
    pooled_zValue = as.numeric(meta_obj$zval),
    pooled_FoldChange = 10^as.numeric(meta_obj$beta[1]),
    pooled_FC_low  = 10^as.numeric(meta_obj$ci.lb),
    pooled_FC_high = 10^as.numeric(meta_obj$ci.ub),
    tau2 = as.numeric(meta_obj$tau2),
    I2 = as.numeric(meta_obj$I2)
  )
}

pooled_results <- bind_rows(pooled_list) %>%
  mutate(
    Sig_raw = case_when(
      pooled_pValue < 0.001 ~ "***",
      pooled_pValue < 0.01  ~ "**",
      pooled_pValue < 0.05  ~ "*",
      TRUE ~ "ns"
    )
  )

print(pooled_results)

df_wide <- pooled_results %>%
  filter(Outcome %in% c("bsize", "bfreq")) %>%
  select(Motif, Outcome,
         pooled_estimate, pooled_SE, pooled_pValue,
         pooled_FoldChange, pooled_FC_low, pooled_FC_high,
         k, tau2, I2) %>%
  pivot_wider(
    names_from = Outcome,
    values_from = c(pooled_estimate, pooled_SE, pooled_pValue,
                    pooled_FoldChange, pooled_FC_low, pooled_FC_high,
                    k, tau2, I2),
    names_sep = "_"
  ) %>%
  filter(!is.na(pooled_FoldChange_bsize), !is.na(pooled_FoldChange_bfreq)) %>%
  mutate(
    est_low_bsize  = pooled_estimate_bsize - 1.96 * pooled_SE_bsize,
    est_high_bsize = pooled_estimate_bsize + 1.96 * pooled_SE_bsize,
    est_low_bfreq  = pooled_estimate_bfreq - 1.96 * pooled_SE_bfreq,
    est_high_bfreq = pooled_estimate_bfreq + 1.96 * pooled_SE_bfreq,
    Significance = case_when(
      pooled_pValue_bsize < 0.05 & pooled_pValue_bfreq < 0.05 ~ "Both",
      pooled_pValue_bsize < 0.05 & pooled_pValue_bfreq >= 0.05 ~ "bsize only",
      pooled_pValue_bsize >= 0.05 & pooled_pValue_bfreq < 0.05 ~ "bfreq only",
      TRUE ~ "Neither"
    )
  )

print(df_wide)
```

```{r exploratory-meta-plot}
p_meta <- ggplot(df_wide,
            aes(x = pooled_estimate_bsize,
                y = pooled_estimate_bfreq,
                color = Significance,
                label = Motif)) +
  geom_errorbarh(aes(xmin = est_low_bsize, xmax = est_high_bsize),
                 height = 0, alpha = 0.4) +
  geom_errorbar(aes(ymin = est_low_bfreq, ymax = est_high_bfreq),
                width = 0, alpha = 0.4) +
  geom_point(size = 3) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray40") +
  geom_hline(yintercept = 0, linetype = "dotted") +
  geom_vline(xintercept = 0, linetype = "dotted") +
  theme_classic(base_size = 14) +
  scale_color_manual(values = c(
    "Both" = "deepskyblue",
    "bsize only" = "cyan3",
    "bfreq only" = "royalblue3",
    "Neither" = "gray60"
  )) +
  labs(
    x = expression("Pooled effect on log"[10]*"(bsize)  (beta)"),
    y = expression("Pooled effect on log"[10]*"(bfreq)  (beta)"),
    title = "Meta-analysis: Association of motif count with burst parameters",
    subtitle = paste0("Points = pooled betas; error bars = 95% CI; min_k = ", min_k)
  )

if (requireNamespace("ggrepel", quietly = TRUE)) {
  p_meta <- p_meta + ggrepel::geom_text_repel(size = 3, max.overlaps = 30)
} else {
  p_meta <- p_meta + geom_text(size = 3, vjust = -0.5, check_overlap = TRUE)
}
print(p_meta)
```

## FIGURE 4A — enhancer motif-frequency donut

Computed directly from the Table 6 motif counts: for each oligo a motif is
"present" when its count is > 0, and the donut shows how many oligos contain
each motif.

```{r FIGURE 4A}
# One row per enhancer oligo with its motif counts.
enh_lib <- unique(anno[, c("seq_ID", cols), with = FALSE])

# Count oligos containing each motif (count > 0).
TF_freq <- data.frame(
  Var1 = cols,
  Freq = vapply(cols, function(m) sum(enh_lib[[m]] > 0, na.rm = TRUE), integer(1))
)

desired_order <- cols
TF_freq$Var1 <- factor(TF_freq$Var1, levels = desired_order)

motif_cols <- c(
  "ETS" = "darkorange",
  "GATA" = "red",
  "AP1" = "orange",
  "SREBP" = "darkred",
  "Trl" = "pink",
  "twist" = "pink3",
  "Ohler6" = "lightblue",
  "Ohler7" = "dodgerblue1",
  "Ohler1" = "darkblue",
  "Dref" = "blue"
)

motif_plot <- ggplot(TF_freq, aes(x = 3, y = Freq, fill = Var1)) +
  geom_col() +
  coord_polar(theta = "y") +
  xlim(c(1.5, 3.5)) +
  scale_fill_manual(values = motif_cols) +
  theme_void()

motif_plot
print(TF_freq)
```

## SUPPL. FIG. 4B — BARe-seq vs STARR-seq validation

Three scatter plots comparing BARe-seq log2FC (independent assay) to STARR-seq
log2FC (this study), one per promoter.

```{r SUPPL. FIG. 4B, fig.width=5, fig.height=5}
# Left panel: BARe-seq (dev, DSCP promoter) vs STARR-seq (DSCP promoter)
fitlm <- lm(log2FoldChange_rg ~ dev_log2FoldChange, data = diff_exp)
PCC <- cor(diff_exp$dev_log2FoldChange, diff_exp$log2FoldChange_rg, 
           method = 'pearson', use = "complete.obs")

p_dscp <- ggplot(diff_exp) + 
  geom_point(aes(x = dev_log2FoldChange, y = log2FoldChange_rg), color = "grey34", alpha = 0.3) + 
  geom_abline(aes(intercept = fitlm$coefficients[1], slope = fitlm$coefficients[2]), 
              linetype = "dashed") + 
  ggtitle("BARe-seq vs. STARR-seq (DSCP)") + 
  labs(subtitle = bquote(R^2 ~ "=" ~ .(round(summary(fitlm)$r.squared, 2)) ~ 
                         ", PCC =" ~ .(round(PCC, 2)))) + 
  ylab("log2FC BARe-seq enhancer activity") + 
  xlab("log2FC STARR-seq enhancer activity") + 
  xlim(-3, 15) + 
  ylim(-3, 15) + 
  coord_fixed() + 
  theme_classic(base_size = 15)

p_dscp

# Middle panel: BARe-seq (hk, Rpl35 promoter) vs STARR-seq (Rpl35 promoter)
fitlm <- lm(hk_log2FoldChange ~ log2FoldChange_p309, data = diff_exp)
PCC <- cor(diff_exp$log2FoldChange_p309, diff_exp$hk_log2FoldChange, 
           method = 'pearson', use = "complete.obs")

p_309 <- ggplot(diff_exp) + 
  geom_point(aes(x = log2FoldChange_p309, y = hk_log2FoldChange), alpha = 0.3, color = "grey34") + 
  geom_abline(aes(intercept = fitlm$coefficients[1], slope = fitlm$coefficients[2]), 
              linetype = "dashed") + 
  ggtitle("BARe-seq vs. STARR-seq (Rpl35)") + 
  labs(subtitle = bquote(R^2 ~ "=" ~ .(round(summary(fitlm)$r.squared, 2)) ~ 
                         ", PCC =" ~ .(round(PCC, 2)))) + 
  xlab("log2FC STARR-seq (Rpl35) enhancer activity") + 
  ylab("log2FC BARe-seq (hk) enhancer activity") + 
  xlim(-3, 15) + 
  ylim(-3, 15) + 
  coord_fixed() + 
  theme_classic(base_size = 15)

p_309

# Right panel: BARe-seq (hk, Rps15A promoter) vs STARR-seq (Rps15A promoter)
fitlm <- lm(hk_log2FoldChange ~ log2FoldChange_p319, data = diff_exp)
PCC <- cor(diff_exp$log2FoldChange_p319, diff_exp$hk_log2FoldChange, 
           method = 'pearson', use = "complete.obs")

p_319 <- ggplot(diff_exp) + 
  geom_point(aes(x = log2FoldChange_p319, y = hk_log2FoldChange), alpha = 0.3, color = "grey34") + 
  geom_abline(aes(intercept = fitlm$coefficients[1], slope = fitlm$coefficients[2]), 
              linetype = "dashed") + 
  ggtitle("BARe-seq vs. STARR-seq (Rps15A)") + 
  labs(subtitle = bquote(R^2 ~ "=" ~ .(round(summary(fitlm)$r.squared, 2)) ~ 
                         ", PCC =" ~ .(round(PCC, 2)))) + 
  xlab("log2FC STARR-seq (Rps15A) enhancer activity") + 
  ylab("log2FC BARe-seq (hk) enhancer activity") + 
  xlim(-3, 15) + 
  ylim(-3, 15) + 
  coord_fixed() + 
  theme_classic(base_size = 15)

p_319
```

## SUPPL. FIG. 4C — mean expression histograms with mu threshold

Faceted histograms showing the distribution of mean expression (log10 scale)
for all 9 STARR-seq samples (3 pooled + 6 replicates), with the mu threshold
marked as a red dashed line.

```{r SUPPL. FIG. 4C, fig.width=6, fig.height=6}
p_mu <- param %>% 
  ggplot() + 
  geom_histogram(aes(x = log10(exp)), bins = 50) + 
  geom_vline(aes(xintercept = log10(mu_threshold)), 
             colour = "red", linetype = "dashed") +
  facet_wrap(~sample, ncol = 3) + 
  theme_classic(base_size = 15)

p_mu
```

## Correlation-heatmap function (Suppl. Fig. 4E/4F)

```{r correlation-function}
make_correlation_heatmap <- function(dt, feature, dataset_name, sample_reps = NULL) {
  dt <- as.data.table(dt)

  if (!"sample" %in% names(dt)) stop("No 'sample' column found in data")

  wide_dt <- dt[, .(
    seq_ID,
    sample_label = sample,
    value = as.numeric(get(feature))
  )]

  if (!is.null(sample_reps)) {
    wide_dt <- wide_dt[sample_label %in% sample_reps]
  }

  if (nrow(wide_dt) == 0) {
    warning(paste("No data left after filtering for sample_reps in dataset:", dataset_name))
    return(NULL)
  }

  wide_dt <- dcast(wide_dt, seq_ID ~ sample_label, value.var = "value")

  mat <- as.matrix(wide_dt[, -1, with = FALSE])
  storage.mode(mat) <- "numeric"
  mat[mat <= 0] <- NA
  mat <- log10(mat)

  mat <- mat[, colSums(is.finite(mat)) >= 2, drop = FALSE]
  if (ncol(mat) < 2) return(NULL)

  cormat <- cor(mat, use = "pairwise.complete.obs", method = "pearson")
  cormat_tbl <- as.data.frame(as.table(cormat))
  colnames(cormat_tbl) <- c("Sample1", "Sample2", "Correlation")

  ggplot(cormat_tbl, aes(Sample1, Sample2, fill = Correlation)) +
    geom_tile(color = "white") +
    scale_fill_gradient2(
      low = "#4575b4", mid = "white", high = "#d73027",
      midpoint = 0.5, limits = c(0, 1), na.value = "grey90"
    ) +
    geom_text(aes(label = ifelse(is.na(Correlation), "NA", sprintf("%.2f", Correlation))), size = 3) +
    labs(title = paste0(dataset_name, " | Feature: ", feature), x = "", y = "") +
    theme_minimal(base_size = 10) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          panel.grid = element_blank())
}

# High-quality enhancer oligos (same filter as the regression analysis).
cor_data <- param[KS_pval > 0.01 & high_mu == TRUE]
features <- c("exp", "bfreq", "bsize")
```

## SUPPL. FIG. 4E — replicate-level correlation heatmaps

```{r SUPPL. FIG. 4E}
for (f in features) {
  p <- make_correlation_heatmap(cor_data, f, "Replicates", sample_reps = replicate_samples)
  if (!is.null(p)) print(p)
}
```

## SUPPL. FIG. 4F — pooled-sample correlation heatmaps

```{r SUPPL. FIG. 4F}
for (f in features) {
  p <- make_correlation_heatmap(cor_data, f, "Pooled", sample_reps = pooled_samples)
  if (!is.null(p)) print(p)
}
```

```{r session-info}
sessionInfo()
```
