---
title: "Fig 4D–E & Supplemental Figure 5: Enhancer Motif-Pasting BARe-seq (libFL031)"
output: html_document
date: "`r Sys.Date()`"
---

<!--
  Panel coverage
  ─────────────────────────────────────────────────────────────────────────────
  Fig 4D       → t1_pub  : log2(rel_exp) ~ log2(rel_bsize) scatter, per motif
  Fig 4E       → t2_pub  : log2(rel_exp) ~ log2(rel_bfreq) scatter, per motif
  Suppl Fig 5A → q3_pool / q4_pool / q5_pool  (horizontal boxplots, pool4)
  Suppl Fig 5E → exp_hist / bsize_hist / bfreq_hist  (WT sequence histograms)
  Suppl Fig 5F → p / q / r  (replicate correlations, pair1 vs pair2)
  Suppl Fig 5G → slope1 / slope2  (per-motif slope bars with R² and p-value)

  Notes on panels not generated by this script
  ─────────────────────────────────────────────
  Suppl Fig 5B / 5C : regression-coefficient heatmaps — see
                       Fig4_S4_enhancer_motifs_clean.Rmd (STARR LASSO
                       regression across promoters and replicates).
  Suppl Fig 5D       : sequence-position track / motif logo — produced by an
                       external tool (MEME / FIMO or equivalent); no R code
                       in this script.

-->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                      fig.width = 10, fig.height = 8)
```

```{r libraries}
library(data.table)
library(tidyr)
library(dplyr)
library(readxl)
library(ggplot2)
library(cowplot)    # replaces ggpubr::ggarrange() — ggpubr needs Matrix ≥1.6.0
library(gridExtra)
library(broom)
library(scales)
```

## 1  Data loading

```{r load-burst-params}
# ── Table 9: burst parameters for libFL031/libFL007 motif-pasting library ────
# Replaces: E*.xlsx files from /groups/stark/.../Parameters_E57_v3/
# Note: the 'sampel' column in Table 9 has a typo; renamed to 'sample' here.
# Sample-name crosswalk (old → new) is in Tables/Table3_mu_thresholds.xlsx.
burst_params <- read_xlsx(
  "/groups/stark/lorbeer/BARE-seq_data_prep/Tables/Table9_enhancer_motifpasting_burst_parameters.xlsx"
) %>%
  rename(sample = sampel) %>%                 # fix upstream typo
  mutate(
    high_mu  = as.logical(high_mu),           # convert "TRUE"/"FALSE" strings
    KS_thres = KS_pval > 0.01,               # pass KS test → burst model valid
    # Derive replicate label from new publication sample names
    rep = case_when(
      endsWith(sample, "_pool")       ~ "pool4",
      endsWith(sample, "_replicate1") ~ "pair1",
      endsWith(sample, "_replicate2") ~ "pair2",
      grepl("_rep1$", sample)         ~ "rep1",
      grepl("_rep2$", sample)         ~ "rep2",
      grepl("_rep3$", sample)         ~ "rep3",
      grepl("_rep4$", sample)         ~ "rep4",
      TRUE                            ~ NA_character_
    ),
    condition = "DSCP"   # all libFL007 motif-pasting samples use DSCP promoter
  )

cat("Replicates in Table 9:\n")
print(table(burst_params$rep))
```

```{r load-annotations}
# ── Table 8: motif-pasting oligo annotations ─────────────────────────────────
# Replaces: annotations_twist18$pasting slot (cluster RDS file).
# Key columns:
#   sequence_motif_id – unique ID for each pasted-motif oligo (= seq_ID in Table 9)
#   original_seq_ID   – parent WT enhancer sequence
#   motif_names_simplified – simplified motif name (e.g. "ETS", "AP1", "WT")
#   motif_group       – same as motif_names_simplified but without _rev suffix
annotations_pasting <- read_xlsx(
  "/groups/stark/lorbeer/BARE-seq_data_prep/Tables/Table8_enhancer_pastedmotifs_library_annotations.xlsx"
)

cat("Table 8 dimensions:", nrow(annotations_pasting), "×", ncol(annotations_pasting), "\n")
cat("motif_group levels:", paste(sort(unique(annotations_pasting$motif_group)), collapse = ", "), "\n")
```

## 2  Motif colour palette and factor order

```{r motif-palette}
motif_order <- c(
  "WT", "ETS", "GATA", "AP1", "SREBP", "Trl", "twist", "Ohler6", "Ohler1", "Dref"
)

motif_cols <- c(
  "ETS"    = "darkorange",
  "GATA"   = "red",
  "AP1"    = "orange",
  "SREBP"  = "darkred",
  "Trl"    = "pink",
  "twist"  = "pink3",
  "Ohler6" = "lightblue",
  "Ohler1" = "darkblue",
  "Dref"   = "blue",
  "WT"     = "black"
)
```

## 3  Annotation, filtering, and relative-value computation

```{r annotate-filter}
# Merge burst parameters with motif-pasting annotations.
# by.x = "seq_ID" (Table 9) matched to by.y = "sequence_motif_id" (Table 8).
# all.x = TRUE: retain all burst-parameter rows, even those without annotation.
data_31_annotated <- merge(
  burst_params, annotations_pasting,
  by.x    = "seq_ID",
  by.y    = "sequence_motif_id",
  all.x   = TRUE
) %>%
  filter(KS_thres == TRUE & high_mu == TRUE)

# Convert to data.table for efficient grouped operations
setDT(data_31_annotated)

# Classify each sequence as forward-orientation, reverse-orientation, or WT
data_31_annotated[, fw_rev :=
  fifelse(
    motif_names_simplified == "WT", "WT",
    fifelse(grepl("_rev$", motif_names_simplified), "rev", "fw")
  )
]

# Simplified motif name (strip _rev suffix; WT stays "WT")
data_31_annotated[, motif_pasted := fcase(
  !is.na(motif_names_simplified) & grepl("WT$",  motif_names_simplified), "WT",
  !is.na(motif_names_simplified),                sub("_rev$", "", motif_names_simplified),
  default = NA_character_
)]

# Apply factor level order
data_31_annotated[, motif_pasted := factor(motif_pasted, levels = motif_order)]

# Sanity check: number of unique WT sequences per replicate
data_31_annotated[
  motif_names_simplified == "WT" & !is.na(original_seq_ID),
  .(n_unique_wt_seqs = uniqueN(original_seq_ID)),
  by = rep
]
```

```{r relative-values}
# ── Step 1: Extract WT burst parameters as per-sequence denominators ──────────
# For each (rep, condition, sample, original_seq_ID) group, take the first
# WT row — exactly one WT oligo exists per parent enhancer sequence.
# seq_ID here is the WT sequence identifier (e.g. "chrX_..._WT").
wt_denoms <- data_31_annotated[
  motif_group == "WT",
  .(
    wt_exp   = exp[1L],
    wt_bsize = bsize[1L],
    wt_bfreq = bfreq[1L]
  ),
  by = .(rep, condition, sample, original_seq_ID, seq_ID)
]

# ── Step 2: Join WT denominators back to all rows ─────────────────────────────
# After this merge, seq_ID.x = pasted-motif seq_ID, seq_ID.y = paired WT seq_ID.
data_31_annotated <- merge(
  data_31_annotated, wt_denoms,
  all.x = TRUE,
  by    = c("original_seq_ID", "rep", "condition", "sample")
)

# ── Step 3: Compute relative burst parameters (pasted-motif / WT) ────────────
# Guards: NA or zero denominators produce NA (not Inf or NaN).
data_31_annotated[
  ,
  `:=`(
    rel_exp   = ifelse(!is.na(wt_exp)   & wt_exp   != 0, exp   / wt_exp,   NA_real_),
    rel_bsize = ifelse(!is.na(wt_bsize) & wt_bsize != 0, bsize / wt_bsize, NA_real_),
    rel_bfreq = ifelse(!is.na(wt_bfreq) & wt_bfreq != 0, bfreq / wt_bfreq, NA_real_)
  )
]
```

## 4  Exploratory QC 

```{r qc-relative-pool4, fig.width=18, fig.height=5}
# Pool4 replicate: relative burst parameters vs WT absolute level, per motif
rel1 <- data_31_annotated %>% filter(rep == "pool4") %>%
  ggplot() +
  geom_point(aes(y = log2(rel_exp), x = log10(wt_exp), color = motif_pasted), size = 3) +
  scale_color_manual(values = motif_cols, drop = FALSE) +
  geom_hline(yintercept = 0) +
  geom_hline(yintercept = c(-1, 1), linetype = "dashed") +
  facet_grid(condition ~ motif_pasted) +
  theme_classic() + theme(axis.text.x = element_text(angle = 90)) +
  labs(y = "log2(rel_exp)", x = "log10(wt_exp)")

rel2 <- data_31_annotated %>% filter(rep == "pool4") %>%
  ggplot() +
  geom_point(aes(y = log2(rel_bsize), x = log10(wt_bsize), color = motif_pasted), size = 3) +
  scale_color_manual(values = motif_cols, drop = FALSE) +
  geom_hline(yintercept = 0) +
  geom_hline(yintercept = c(-1, 1), linetype = "dashed") +
  facet_grid(condition ~ motif_pasted) +
  theme_classic() + theme(axis.text.x = element_text(angle = 90)) +
  labs(y = "log2(rel_bsize)", x = "log10(wt_bsize)")

rel3 <- data_31_annotated %>% filter(rep == "pool4") %>%
  ggplot() +
  geom_point(aes(y = log2(rel_bfreq), x = log10(wt_bfreq), color = motif_pasted), size = 3) +
  scale_color_manual(values = motif_cols, drop = FALSE) +
  geom_hline(yintercept = 0) +
  geom_hline(yintercept = c(-1, 1), linetype = "dashed") +
  facet_grid(condition ~ motif_pasted) +
  theme_classic() + theme(axis.text.x = element_text(angle = 90)) +
  labs(y = "log2(rel_bfreq)", x = "log10(wt_bfreq)")

plot_grid(rel1, rel2, rel3, ncol = 1)
```

## 5  Suppl Fig 5F — Replicate correlations (pair1 vs pair2)

```{r s5f-replicate-correlations, fig.width=12, fig.height=5, fig.cap="Suppl Fig 5F"}
# Merge pair1 and pair2 on the sequence identifier (seq_ID.x = pasted-motif ID)
pair1 <- data_31_annotated[rep == "pair1"]
pair2 <- data_31_annotated[rep == "pair2"]
plot_pair <- merge(pair1, pair2, by = "seq_ID.x")

# ── Mean expression ───────────────────────────────────────────────────────────
PCC_exp <- cor(log10(plot_pair$exp.x), log10(plot_pair$exp.y),
               method = "pearson", use = "complete.obs")
p <- plot_pair %>%
  ggplot() +
  geom_point(aes(x = log10(exp.x), y = log10(exp.y)), color = "#1f77b4", alpha = 0.3) +
  labs(
    title = paste0("PCC = ", round(PCC_exp, 2)),
    x = "Mean expression (log10) replicate 1",
    y = "Mean expression (log10) replicate 2"
  ) +
  coord_fixed(ratio = 1,
    xlim = range(c(log10(plot_pair$exp.x), log10(plot_pair$exp.y)), na.rm = TRUE),
    ylim = range(c(log10(plot_pair$exp.x), log10(plot_pair$exp.y)), na.rm = TRUE)
  ) +
  theme_classic(base_size = 15)

# ── Burst frequency ───────────────────────────────────────────────────────────
PCC_bfreq <- cor(log10(plot_pair$bfreq.x), log10(plot_pair$bfreq.y),
                 method = "pearson", use = "complete.obs")
q <- plot_pair %>%
  ggplot() +
  geom_point(aes(x = log10(bfreq.x), y = log10(bfreq.y)), color = "#1f77b4", alpha = 0.3) +
  labs(
    title = paste0("PCC = ", round(PCC_bfreq, 2)),
    x = "Burst frequency (log10) replicate 1",
    y = "Burst frequency (log10) replicate 2"
  ) +
  coord_fixed(ratio = 1,
    xlim = range(c(log10(plot_pair$bfreq.x), log10(plot_pair$bfreq.y)), na.rm = TRUE),
    ylim = range(c(log10(plot_pair$bfreq.x), log10(plot_pair$bfreq.y)), na.rm = TRUE)
  ) +
  theme_classic(base_size = 15)

# ── Burst size ────────────────────────────────────────────────────────────────
PCC_bsize <- cor(log10(plot_pair$bsize.x), log10(plot_pair$bsize.y),
                 method = "pearson", use = "complete.obs")
r <- plot_pair %>%
  ggplot() +
  geom_point(aes(x = log10(bsize.x), y = log10(bsize.y)), color = "#1f77b4", alpha = 0.3) +
  labs(
    title = paste0("PCC = ", round(PCC_bsize, 2)),
    x = "Burst size (log10) replicate 1",
    y = "Burst size (log10) replicate 2"
  ) +
  coord_fixed(ratio = 1,
    xlim = range(c(log10(plot_pair$bsize.x), log10(plot_pair$bsize.y)), na.rm = TRUE),
    ylim = range(c(log10(plot_pair$bsize.x), log10(plot_pair$bsize.y)), na.rm = TRUE)
  ) +
  theme_classic(base_size = 15)

plot_grid(p, q, r, ncol = 3)
```

## 6  Suppl Fig 5E — WT sequence burst parameter histograms (pool4)

```{r s5e-wt-histograms, fig.width=10, fig.height=4, fig.cap="Suppl Fig 5E"}
wt_pool4 <- data_31_annotated %>%
  filter(rep == "pool4", !is.na(rel_exp), motif_pasted == "WT")

exp_hist <- ggplot(wt_pool4) +
  geom_histogram(aes(x = log10(exp)), bins = 10) +
  labs(x = "log10(mean expression)", title = "Mean expression") +
  theme_classic(base_size = 15)

bsize_hist <- ggplot(wt_pool4) +
  geom_histogram(aes(x = log10(bsize)), bins = 10) +
  labs(x = "log10(burst size)", title = "Burst size") +
  theme_classic(base_size = 15)

bfreq_hist <- ggplot(wt_pool4) +
  geom_histogram(aes(x = log10(bfreq)), bins = 10) +
  labs(x = "log10(burst frequency)", title = "Burst frequency") +
  theme_classic(base_size = 15)

plot_grid(exp_hist, bsize_hist, bfreq_hist, ncol = 3)
```

## 7  Summary statistics — per motif and per enhancer (pool4)

```{r summary-per-enhancer}
# Per-enhancer motif insertion summary (pool4, non-WT)
group_metrics <- data_31_annotated %>%
  filter(rep == "pool4", !is.na(rel_exp), motif_pasted != "WT") %>%
  group_by(original_seq_ID) %>%
  summarise(
    n_motif_insertions = n(),
    med_rel_bfreq      = median(rel_bfreq, na.rm = TRUE),
    med_rel_bsize      = median(rel_bsize, na.rm = TRUE),
    med_rel_exp        = median(rel_exp,   na.rm = TRUE),
    .groups = "drop"
  )

# Per-motif summary (pool4, non-WT)
group_metrics_2 <- data_31_annotated %>%
  filter(rep == "pool4", !is.na(rel_exp), motif_pasted != "WT") %>%
  group_by(motif_pasted) %>%
  summarise(
    n_motif_insertions = n(),
    med_rel_bfreq      = median(rel_bfreq, na.rm = TRUE),
    med_rel_bsize      = median(rel_bsize, na.rm = TRUE),
    med_rel_exp        = median(rel_exp,   na.rm = TRUE),
    .groups = "drop"
  )

print(group_metrics_2)
```

## 8  Paired Wilcoxon tests (pasted motif vs WT, pool4)

```{r wilcoxon-tests}
# Restrict to pool4 for all statistical tests
DT <- copy(data_31_annotated)
DT <- DT[!is.na(rep) & !is.na(motif_pasted) & rep == "pool4"]
DT[, motif_pasted := as.character(motif_pasted)]

# Melt to long format across the three burst metrics
long <- data.table::melt(
  DT,
  id.vars      = c("rep", "motif_pasted", "rel_exp", "original_seq_ID"),
  measure.vars = c("exp", "bfreq", "bsize"),
  variable.name = "metric",
  value.name    = "value",
  na.rm         = FALSE
)

# Median WT value per (rep, metric, original_seq_ID) for paired comparison
wt_map <- long[
  motif_pasted == "WT" & !is.na(original_seq_ID) & is.finite(value),
  .(wt_value = median(value)),
  by = .(rep, metric, original_seq_ID)
]

# Pair each non-WT measurement with its corresponding WT
paired_dt <- long[
  motif_pasted != "WT" & !is.na(original_seq_ID) & is.finite(value),
  .(rep, motif_pasted, metric, original_seq_ID, x = value)
][wt_map, on = .(rep, metric, original_seq_ID), nomatch = 0]

# Run paired Wilcoxon per (rep, motif, metric)
res <- paired_dt[
  ,
  {
    x <- x[is.finite(x)]
    y <- wt_value[is.finite(wt_value)]
    d <- x - y

    out <- list(
      n_pairs      = length(d),
      median_group = if (length(x)) median(x) else NA_real_,
      median_wt    = if (length(y)) median(y) else NA_real_,
      diff_median  = if (length(d)) median(d) else NA_real_,
      p_value      = NA_real_,
      test         = "wilcox_paired",
      note         = NA_character_
    )

    if (length(d) < 2) {
      out$note <- "insufficient_pairs"
    } else if (all(d == 0)) {
      out$note    <- "all_diffs_zero"
      out$p_value <- 1
    } else {
      out$p_value <- tryCatch(
        unname(wilcox.test(x, y, paired = TRUE, exact = FALSE)$p.value),
        error = function(e) {
          out$note <<- paste0("test_error: ", e$message)
          NA_real_
        }
      )
    }
    as.data.table(out)
  },
  by = .(rep, motif_pasted, metric)
]

# BH correction within each rep × metric combination
res[, p_adj_bh := p.adjust(p_value, method = "BH"), by = .(rep, metric)]

# Significance labels (using raw p-value for display, as in original)
res <- res %>%
  mutate(
    sig_label = case_when(
      p_value < 0.001 ~ "***",
      p_value < 0.01  ~ "**",
      p_value < 0.05  ~ "*",
      TRUE             ~ "ns"
    )
  )

# Re-apply factor order to result table
desired_order <- names(motif_cols)
data_31_annotated <- data_31_annotated %>%
  mutate(motif_pasted = factor(motif_pasted, levels = desired_order))
res <- res %>%
  mutate(motif_pasted = factor(motif_pasted, levels = desired_order))

print(res %>% select(motif_pasted, metric, n_pairs, diff_median, p_value, p_adj_bh, sig_label))
```

## 9  Suppl Fig 5A — Horizontal boxplots by motif (pool4)


```{r s5a-boxplots-horizontal, fig.width=10, fig.height=18, fig.cap="Suppl Fig 5A – horizontal boxplots"}
# ── S5A: Horizontal boxplots by motif, pool4 replicate ───────────────────────
x_pos_q3 <- 3.25
x_pos_q4 <- 3.25
x_pos_q5 <- 1.35

q3_pool <- data_31_annotated %>%
  filter(motif_names_simplified != "NA" & rep == "pool4") %>%
  ggplot() +
  geom_jitter(aes(y = motif_pasted, x = log2(rel_exp), color = motif_pasted),
              alpha = 0.5, position = position_jitter(width = 0, height = 0.1)) +
  geom_boxplot(aes(y = motif_pasted, x = log2(rel_exp), color = motif_pasted),
               alpha = 0.2, outlier.shape = NA, width = 0.3) +
  scale_color_manual(values = motif_cols, drop = FALSE) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  geom_text(data = res %>% filter(metric == "exp"),
            aes(y = motif_pasted, x = x_pos_q3, label = sig_label),
            size = 4, fontface = "bold", hjust = 0) +
  facet_wrap(~motif_pasted, scales = "free_y", ncol = 1) +
  theme_classic(base_size = 15) + theme(legend.position = "none") +
  labs(x = "log2(rel_exp)", y = NULL, title = "Mean expression")

q4_pool <- data_31_annotated %>%
  filter(motif_names_simplified != "NA" & rep == "pool4") %>%
  ggplot() +
  geom_jitter(aes(y = motif_pasted, x = log2(rel_bfreq), color = motif_pasted),
              alpha = 0.5, position = position_jitter(width = 0, height = 0.1)) +
  geom_boxplot(aes(y = motif_pasted, x = log2(rel_bfreq), color = motif_pasted),
               alpha = 0.2, outlier.shape = NA, width = 0.3) +
  scale_color_manual(values = motif_cols, drop = FALSE) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  geom_text(data = res %>% filter(metric == "bfreq"),
            aes(y = motif_pasted, x = x_pos_q4, label = sig_label),
            size = 4, fontface = "bold", hjust = 0) +
  facet_wrap(~motif_pasted, scales = "free_y", ncol = 1) +
  theme_classic(base_size = 15) + theme(legend.position = "none") +
  labs(x = "log2(rel_bfreq)", y = NULL, title = "Burst frequency")

q5_pool <- data_31_annotated %>%
  filter(motif_names_simplified != "NA" & rep == "pool4") %>%
  ggplot() +
  geom_jitter(aes(y = motif_pasted, x = log2(rel_bsize), color = motif_pasted),
              alpha = 0.5, position = position_jitter(width = 0, height = 0.1)) +
  geom_boxplot(aes(y = motif_pasted, x = log2(rel_bsize), color = motif_pasted),
               alpha = 0.2, outlier.shape = NA, width = 0.3) +
  scale_color_manual(values = motif_cols, drop = FALSE) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  geom_text(data = res %>% filter(metric == "bsize"),
            aes(y = motif_pasted, x = x_pos_q5, label = sig_label),
            size = 4, fontface = "bold", hjust = 0) +
  facet_wrap(~motif_pasted, scales = "free_y", ncol = 1) +
  theme_classic(base_size = 15) + theme(legend.position = "none") +
  labs(x = "log2(rel_bsize)", y = NULL, title = "Burst size")

grid.arrange(q3_pool, q4_pool, q5_pool, ncol = 3)
```

## 10  Suppl Fig 5G — Per-motif slope plots with R² (pool4)

```{r slope-data-prep}
# Data for regression: log2-transform relative values, pool4, annotated sequences
df_slopes <- data_31_annotated %>%
  filter(motif_names_simplified != "NA", rep == "pool4") %>%
  mutate(
    log2_exp   = log2(rel_exp),
    log2_bsize = log2(rel_bsize),
    log2_bfreq = log2(rel_bfreq)
  )
```

```{r slopes-bsize}
# Slopes + p-values + R² for burst size
# Y = log2(rel_exp), X = log2(rel_bsize): expression as a function of burst size
slopes_bsize <- df_slopes %>%
  group_by(condition, motif_pasted) %>%
  do({
    fit        <- lm(log2_exp ~ log2_bsize, data = .)
    tidy_fit   <- tidy(fit)   %>% filter(term == "log2_bsize")
    glance_fit <- glance(fit)
    data.frame(
      slope     = tidy_fit$estimate,
      p_value   = tidy_fit$p.value,
      std_error = tidy_fit$std.error,
      statistic = tidy_fit$statistic,
      r_squared = glance_fit$r.squared
    )
  }) %>%
  ungroup()
```

```{r slopes-bfreq}
# Slopes + p-values + R² for burst frequency
slopes_bfreq <- df_slopes %>%
  group_by(condition, motif_pasted) %>%
  do({
    fit        <- lm(log2_exp ~ log2_bfreq, data = .)
    tidy_fit   <- tidy(fit)   %>% filter(term == "log2_bfreq")
    glance_fit <- glance(fit)
    data.frame(
      slope     = tidy_fit$estimate,
      p_value   = tidy_fit$p.value,
      std_error = tidy_fit$std.error,
      statistic = tidy_fit$statistic,
      r_squared = glance_fit$r.squared
    )
  }) %>%
  ungroup()
```

```{r slope-plot-helper}
# Add 95% CI and combined R²/p-value label
prep_slopes_for_plot <- function(d) {
  d %>%
    mutate(
      ci_low  = slope - 1.96 * std_error,
      ci_high = slope + 1.96 * std_error,
      p_label = sprintf("R2=%.2f, p=%s",
                        r_squared,
                        pvalue(p_value, accuracy = 0.001))
    )
}

slopes_bsize_p <- prep_slopes_for_plot(slopes_bsize)
slopes_bfreq_p <- prep_slopes_for_plot(slopes_bfreq)
```

```{r s5g-slope-plots, fig.width=8, fig.height=18, fig.cap="Suppl Fig 5G – per-motif regression slopes"}
# ── S5G: Slope bar plots with R² and p-value ─────────────────────────────────
slope1 <- ggplot(slopes_bsize_p,
                 aes(x = motif_pasted, y = slope, fill = motif_pasted)) +
  geom_col() +
  geom_errorbar(aes(ymin = ci_low, ymax = ci_high), width = 0.2) +
  geom_text(
    aes(
      y     = ifelse(slope > 0,
                     ci_high + 0.02 * max(ci_high,       na.rm = TRUE),
                     ci_low  - 0.02 * max(abs(ci_low),   na.rm = TRUE)),
      label = p_label
    ),
    vjust = ifelse(slopes_bsize_p$slope > 0, 0, 1),
    hjust = 0.5, size = 3
  ) +
  facet_wrap(~motif_pasted, ncol = 1, scales = "free_x") +
  scale_fill_manual(values = motif_cols, drop = FALSE) +
  labs(x = NULL,
       y = "Slope (log2(rel_exp) ~ log2(rel_bsize))",
       title = "Per-motif slopes: burst size") +
  theme_classic(base_size = 15) +
  theme(legend.position = "none",
        plot.margin = margin(5.5, 5.5, 5.5, 5.5))

slope2 <- ggplot(slopes_bfreq_p,
                 aes(x = motif_pasted, y = slope, fill = motif_pasted)) +
  geom_col() +
  geom_errorbar(aes(ymin = ci_low, ymax = ci_high), width = 0.2) +
  geom_text(
    aes(
      y     = ifelse(slope > 0,
                     ci_high + 0.02 * max(ci_high,       na.rm = TRUE),
                     ci_low  - 0.02 * max(abs(ci_low),   na.rm = TRUE)),
      label = p_label
    ),
    vjust = ifelse(slopes_bfreq_p$slope > 0, 0, 1),
    hjust = 0.5, size = 3
  ) +
  scale_fill_manual(values = motif_cols, drop = FALSE) +
  facet_wrap(~motif_pasted, ncol = 1, scales = "free_x") +
  labs(x = NULL,
       y = "Slope (log2(rel_exp) ~ log2(rel_bfreq))",
       title = "Per-motif slopes: burst frequency") +
  theme_classic(base_size = 15) +
  theme(legend.position = "none",
        plot.margin = margin(5.5, 5.5, 5.5, 5.5))

plot_grid(slope1, slope2, ncol = 2)
```

## 11  Fig 4D & 4E — Scatter plots: relative expression ~ burst parameters

```{r fig4de-data-prep}
# Pool4 filtered data; WT sequences excluded from scatter (kept for context in
# boxplots but the main scatter shows pasted-motif effects only)
plot_data_initial <- data_31_annotated %>%
  filter(motif_names_simplified != "NA", rep == "pool4")

# Shared y-axis range across both panels so plots are directly comparable
y_limits <- range(log2(plot_data_initial$rel_exp), na.rm = TRUE)
```

```{r fig4de-intermediate, fig.width=6, fig.height=15}
# Intermediate version (factor levels in motif_order; pool4 already filtered)
t1_inter <- plot_data_initial %>%
  ggplot(aes(y = log2(rel_exp), x = log2(rel_bsize), color = motif_pasted)) +
  geom_point(alpha = 0.5) +
  geom_smooth(aes(group = motif_pasted, fill = motif_pasted),
              method = "lm", se = TRUE, alpha = 0.2) +
  scale_color_manual(values = motif_cols, drop = FALSE) +
  scale_fill_manual( values = motif_cols, drop = FALSE) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_vline(xintercept = 0, linetype = "dashed") +
  scale_y_continuous(limits = y_limits) +
  facet_wrap(~motif_pasted, ncol = 1) +
  theme_classic(base_size = 15) + theme(legend.position = "none") +
  labs(y = "log2(rel_exp)", x = "log2(rel_bsize)")

t2_inter <- plot_data_initial %>%
  ggplot(aes(y = log2(rel_exp), x = log2(rel_bfreq), color = motif_pasted)) +
  geom_point(alpha = 0.5) +
  geom_smooth(aes(group = motif_pasted, fill = motif_pasted),
              method = "lm", se = TRUE, alpha = 0.2) +
  scale_color_manual(values = motif_cols, drop = FALSE) +
  scale_fill_manual( values = motif_cols, drop = FALSE) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_vline(xintercept = 0, linetype = "dashed") +
  scale_y_continuous(limits = y_limits) +
  facet_wrap(~motif_pasted, ncol = 1) +
  theme_classic(base_size = 15) + theme(legend.position = "none") +
  labs(y = "log2(rel_exp)", x = "log2(rel_bfreq)")

plot_grid(t1_inter, t2_inter, ncol = 2)
```

```{r fig4de-publication, fig.width=6, fig.height=15, fig.cap="Fig 4D (left) and Fig 4E (right) – reversed factor order"}
# ── Fig 4D and 4E: publication version with reversed motif factor order ───────
plot_data <- data_31_annotated %>%
  filter(motif_names_simplified != "NA", rep == "pool4") %>%
  mutate(motif_pasted = factor(
    motif_pasted,
    levels = rev(levels(factor(motif_pasted)))  # reverse order for display
  ))

# Fig 4D: log2(rel_exp) ~ log2(rel_bsize)
t1_pub <- plot_data %>%
  ggplot(aes(y = log2(rel_exp), x = log2(rel_bsize), color = motif_pasted)) +
  geom_point(alpha = 0.5) +
  geom_smooth(aes(group = motif_pasted, fill = motif_pasted),
              method = "lm", se = TRUE, alpha = 0.2) +
  scale_color_manual(values = motif_cols, drop = FALSE) +
  scale_fill_manual( values = motif_cols, drop = FALSE) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_vline(xintercept = 0, linetype = "dashed") +
  facet_wrap(~motif_pasted, ncol = 1) +
  theme_classic(base_size = 15) + theme(legend.position = "none") +
  labs(y = "log2(rel_exp)", x = "log2(rel_bsize)", title = "Fig 4D")

# Fig 4E: log2(rel_exp) ~ log2(rel_bfreq)
t2_pub <- plot_data %>%
  ggplot(aes(y = log2(rel_exp), x = log2(rel_bfreq), color = motif_pasted)) +
  geom_point(alpha = 0.5) +
  geom_smooth(aes(group = motif_pasted, fill = motif_pasted),
              method = "lm", se = TRUE, alpha = 0.2) +
  scale_color_manual(values = motif_cols, drop = FALSE) +
  scale_fill_manual( values = motif_cols, drop = FALSE) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_vline(xintercept = 0, linetype = "dashed") +
  facet_wrap(~motif_pasted, ncol = 1) +
  theme_classic(base_size = 15) + theme(legend.position = "none") +
  labs(y = "log2(rel_exp)", x = "log2(rel_bfreq)", title = "Fig 4E")

plot_grid(t1_pub, t2_pub, ncol = 2)
```

## 12  Exploratory — bsize vs bfreq relationship

```{r exploratory-bsize-vs-bfreq, fig.width=8, fig.height=15}
data_31_annotated %>%
  filter(motif_names_simplified != "NA" & rep == "pool4") %>%
  ggplot() +
  geom_point(aes(x = log2(rel_bfreq), y = log2(rel_bsize), color = motif_pasted)) +
  scale_color_manual(values = motif_cols, drop = FALSE) +
  geom_hline(yintercept = log2(1), linetype = "dashed") +
  geom_vline(xintercept = log2(1), linetype = "dashed") +
  facet_wrap(~motif_pasted, ncol = 1) +
  theme_classic() +
  labs(x = "log2(rel_bfreq)", y = "log2(rel_bsize)")
```

## 13  Exploratory — median ± IQR summary scatter (pool4)

```{r exploratory-median-iqr, fig.width=10, fig.height=5}
# Per (rep, condition, sample, motif_pasted) medians and IQRs on log2 scale
dt_sum <- data_31_annotated[
  motif_names_simplified != "NA" & rep == "pool4",
  .(
    rel_exp_med   = median(rel_exp,   na.rm = TRUE),
    rel_exp_q25   = quantile(rel_exp,   0.25, na.rm = TRUE),
    rel_exp_q75   = quantile(rel_exp,   0.75, na.rm = TRUE),
    rel_bfreq_med = median(rel_bfreq, na.rm = TRUE),
    rel_bfreq_q25 = quantile(rel_bfreq, 0.25, na.rm = TRUE),
    rel_bfreq_q75 = quantile(rel_bfreq, 0.75, na.rm = TRUE),
    rel_bsize_med = median(rel_bsize, na.rm = TRUE),
    rel_bsize_q25 = quantile(rel_bsize, 0.25, na.rm = TRUE),
    rel_bsize_q75 = quantile(rel_bsize, 0.75, na.rm = TRUE),
    n             = .N
  ),
  by = .(rep, condition, sample, motif_pasted)
]

dt_sum[, `:=`(
  x_med       = log2(rel_exp_med),
  x_lo        = log2(rel_exp_q25),
  x_hi        = log2(rel_exp_q75),
  y_bfreq_med = log2(rel_bfreq_med),
  y_bfreq_lo  = log2(rel_bfreq_q25),
  y_bfreq_hi  = log2(rel_bfreq_q75),
  y_bsize_med = log2(rel_bsize_med),
  y_bsize_lo  = log2(rel_bsize_q25),
  y_bsize_hi  = log2(rel_bsize_q75)
)]

sum1 <- ggplot(dt_sum) +
  geom_errorbar(aes(x = x_med, ymin = y_bsize_lo, ymax = y_bsize_hi,
                    color = motif_pasted), width = 0) +
  geom_errorbarh(aes(y = y_bsize_med, xmin = x_lo, xmax = x_hi,
                     color = motif_pasted), height = 0) +
  geom_point(aes(x = x_med, y = y_bsize_med, color = motif_pasted), size = 3) +
  scale_color_manual(values = motif_cols, drop = FALSE) +
  geom_hline(yintercept = 0) + geom_vline(xintercept = 0) +
  facet_grid(rep ~ condition) +
  theme_classic() + theme(axis.text.x = element_text(angle = 45)) +
  labs(x = "log2(rel_exp) median", y = "log2(rel_bsize) median")

sum2 <- ggplot(dt_sum) +
  geom_errorbar(aes(x = x_med, ymin = y_bfreq_lo, ymax = y_bfreq_hi,
                    color = motif_pasted), width = 0) +
  geom_errorbarh(aes(y = y_bfreq_med, xmin = x_lo, xmax = x_hi,
                     color = motif_pasted), height = 0) +
  geom_point(aes(x = x_med, y = y_bfreq_med, color = motif_pasted), size = 3) +
  scale_color_manual(values = motif_cols, drop = FALSE) +
  geom_hline(yintercept = 0) + geom_vline(xintercept = 0) +
  facet_grid(rep ~ condition) +
  theme_classic() + theme(axis.text.x = element_text(angle = 45)) +
  labs(x = "log2(rel_exp) median", y = "log2(rel_bfreq) median")

grid.arrange(sum1, sum2, ncol = 2)
```


```{r session-info}
sessionInfo()
```
