---
title: "STAP_linear_regression_March2026_reducedmotifs"
author: "FL"
date: "`r Sys.Date()`"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE
)

require(data.table)
require(ggplot2)
require(broom)
require(purrr)
require(metafor)
library(dplyr)
library(tidyr)
require(readxl)

```

## Load pre-processed data
1) Table with all derived burst paramters of the STAP-SEQ (PROMOTER) experiments
2) sample-specific expression thresholds from mu_thresholds_v2_Reyna.xlsx 

```{r}

out_dir <- "/groups/stark/lorbeer/Burst_params_DG/STAP_March2025_pools/Fig2_promoter_motifs/"
#load data 
param <- fread("/groups/stark/lorbeer/Burst_params_DG/STAP_March2025_pools/STAP_data_annotated_log2FC.txt")
sample_names <- unique(as.factor(param$sample))
sample_names <- sample_names[c(1:3,7:12)]

mu_thresholds <- read_xlsx("/groups/stark/lorbeer/Burst_params_DG/mu_thresholds_v2_Reyna.xlsx")
param <- merge(param, mu_thresholds, by = c("sample", "experiment", "library"), all.x = TRUE)
param[, high_mu := exp >= mu_threshold_nofilter]
```

## Create filtered per-sample list
Filter the parameters that pass the high_mu and KS_pval thresholds
```{r}
param_list <- list()

# Read data into a list object and apply filtering criteria for each sample
for (s in sample_names) {
  # param_list[[s]] <- as.data.frame(param[sample == s & (full_prec2 & high_mu == TRUE)])
  #param_list[[s]] <- as.data.frame(param[sample == s & (bfreq.thres == TRUE & bsize.thres == TRUE & KS_pval > 0.01)])
  param_list[[s]] <- as.data.frame(param[sample == s & (high_mu & KS_pval > 0.01)])
}
```

# Helper functions
1) function sig_stars: significance labels 
2) creates placeholder result tables filled with NA values when a model cannot be fit to avoid downstream problems only necessary for non-pooled stuff 

```{r}
sig_stars <- function(p) {
  case_when(
    is.na(p)  ~ NA_character_,
    p < 0.001 ~ "***",
    p < 0.01  ~ "**",
    p < 0.05  ~ "*",
    TRUE      ~ "ns"
  )
}

na_block_for_motifs <- function(motifs, outcome, model_label, n_used = NA_integer_) {
  tibble(
    Motif = motifs,
    Outcome = outcome,
    Model = model_label,
    n_used = n_used,
    n_present = NA_integer_,
    prop_present = NA_real_,
    estimate = NA_real_,
    std.error = NA_real_,
    statistic = NA_real_,
    p.value = NA_real_,
    FoldChange = NA_real_
  )
}

```

# Model fitting functions
Two functions implementing the core regression framework used in the analysis. 

- univariate models, where each motif is tested separately, and 
- multivariate models, where all selected motifs are included simultaneously in the same regression. 

linear regression with lm()
log10(outcome) ~ motif1 + motif2... motifn

Both model types can also be run with or without adjustment for expression through log10(exp). 
Output: estimated effect sizes, standard errors, p-values, and fold-change values derived from the regression coefficients


```{r}

fit_motif_model_one_df <- function(data, outcome, motifs, include_exp = FALSE) {
  model_label <- ifelse(include_exp, "motifs_plus_log10exp", "motifs_only")
  
  motifs_present <- intersect(motifs, colnames(data))
  motifs_missing <- setdiff(motifs, motifs_present)
  
  if (!outcome %in% colnames(data)) {
    return(na_block_for_motifs(motifs, outcome, model_label, n_used = NA_integer_))
  }
  if (include_exp && !"exp" %in% colnames(data)) {
    return(na_block_for_motifs(motifs, outcome, model_label, n_used = NA_integer_))
  }
  
  fitted_present <- purrr::map_dfr(motifs_present, function(m) {
    rhs <- m
    if (include_exp) rhs <- paste(rhs, "+ log10(exp)")
    fml <- as.formula(paste0("log10(", outcome, ") ~ ", rhs))
    
    mf <- tryCatch(model.frame(fml, data = data), error = function(e) NULL)
    n_used <- if (is.null(mf)) NA_integer_ else nrow(mf)
    
    if (is.null(mf) || n_used == 0) {
      return(
        na_block_for_motifs(m, outcome, model_label, n_used = ifelse(is.null(mf), NA_integer_, 0L)) %>%
          dplyr::mutate(
            n_present = ifelse(is.null(mf), NA_integer_, sum(mf[[m]] == 1, na.rm = TRUE)),
            prop_present = ifelse(is.null(mf), NA_real_, n_present / n_used)
          )
      )
    }
    
    n_present <- sum(mf[[m]] == 1, na.rm = TRUE)
    prop_present <- n_present / n_used
    
    model <- tryCatch(lm(fml, data = data), error = function(e) NULL)
    if (is.null(model)) {
      return(
        na_block_for_motifs(m, outcome, model_label, n_used = n_used) %>%
          dplyr::mutate(n_present = n_present, prop_present = prop_present)
      )
    }
    
    broom::tidy(model) %>%
      dplyr::filter(term == m) %>%
      dplyr::transmute(
        Motif = m,
        Outcome = outcome,
        Model = model_label,
        n_used = n_used,
        n_present = n_present,
        prop_present = prop_present,
        estimate = estimate,
        std.error = std.error,
        statistic = statistic,
        p.value = p.value,
        FoldChange = 10^estimate
      )
  })
  
  missing_part <- na_block_for_motifs(motifs_missing, outcome, model_label, n_used = NA_integer_) %>%
    dplyr::mutate(n_present = NA_integer_, prop_present = NA_real_)
  
  dplyr::bind_rows(fitted_present, missing_part) %>%
    dplyr::mutate(Motif = factor(Motif, levels = motifs)) %>%
    dplyr::arrange(Motif) %>%
    dplyr::mutate(Motif = as.character(Motif))
}

fit_motif_model_multiv_df <- function(data, outcome, motifs, include_exp = FALSE) {
  model_label <- ifelse(include_exp, "motifs_plus_log10exp", "motifs_only")
  
  motifs_present <- intersect(motifs, colnames(data))
  motifs_missing <- setdiff(motifs, motifs_present)
  
  if (!outcome %in% colnames(data)) {
    return(na_block_for_motifs(motifs, outcome, model_label, n_used = NA_integer_))
  }
  if (include_exp && !"exp" %in% colnames(data)) {
    return(na_block_for_motifs(motifs, outcome, model_label, n_used = NA_integer_))
  }
  if (length(motifs_present) == 0) {
    return(na_block_for_motifs(motifs, outcome, model_label, n_used = NA_integer_))
  }
  
  rhs <- paste(motifs_present, collapse = " + ")
  if (include_exp) rhs <- paste(rhs, "+ log10(exp)")
  fml <- as.formula(paste0("log10(", outcome, ") ~ ", rhs))
  
  mf <- tryCatch(model.frame(fml, data = data), error = function(e) NULL)
  n_used <- if (is.null(mf)) NA_integer_ else nrow(mf)
  
  if (is.null(mf) || n_used == 0) {
    out <- na_block_for_motifs(motifs_present, outcome, model_label, n_used = ifelse(is.null(mf), NA_integer_, 0L))
    return(dplyr::bind_rows(
      out,
      na_block_for_motifs(motifs_missing, outcome, model_label, n_used = NA_integer_)
    ))
  }
  
  presence <- tibble::tibble(
    Motif = motifs_present,
    n_present = purrr::map_int(motifs_present, ~ sum(mf[[.x]] == 1, na.rm = TRUE)),
    prop_present = n_present / n_used
  )
  
  model <- tryCatch(lm(fml, data = data), error = function(e) NULL)
  if (is.null(model)) {
    out <- na_block_for_motifs(motifs_present, outcome, model_label, n_used = n_used)
    out <- dplyr::left_join(out, presence, by = "Motif")
    return(dplyr::bind_rows(
      out,
      na_block_for_motifs(motifs_missing, outcome, model_label, n_used = NA_integer_)
    ))
  }
  
  coefs <- broom::tidy(model) %>%
    dplyr::filter(term %in% motifs_present) %>%
    dplyr::transmute(
      Motif = term,
      Outcome = outcome,
      Model = model_label,
      n_used = n_used,
      estimate = estimate,
      std.error = std.error,
      statistic = statistic,
      p.value = p.value,
      FoldChange = 10^estimate
    ) %>%
    dplyr::left_join(presence, by = "Motif")
  
  missing_part <- na_block_for_motifs(motifs_missing, outcome, model_label, n_used = NA_integer_) %>%
    dplyr::mutate(n_present = NA_integer_, prop_present = NA_real_)
  
  dplyr::bind_rows(coefs, missing_part) %>%
    dplyr::mutate(Motif = factor(Motif, levels = motifs)) %>%
    dplyr::arrange(Motif) %>%
    dplyr::mutate(Motif = as.character(Motif))
}

fit_motif_models_list_both <- function(data_list, outcomes, motifs, include_exp = FALSE) {
  purrr::imap_dfr(data_list, function(df, nm) {
    purrr::map_dfr(outcomes, function(outcome) {
      
      multiv <- fit_motif_model_multiv_df(
        data = df, outcome = outcome, motifs = motifs, include_exp = include_exp
      ) %>% dplyr::mutate(Analysis = "multivariate")
      
      univ <- fit_motif_model_one_df(
        data = df, outcome = outcome, motifs = motifs, include_exp = include_exp
      ) %>% dplyr::mutate(Analysis = "univariate")
      
      dplyr::bind_rows(multiv, univ) %>%
        dplyr::mutate(
          sample_name = nm,
          sig = sig_stars(p.value)
        )
    })
  }) %>%
    dplyr::select(sample_name, Outcome, Analysis, Model, Motif, sig, dplyr::everything())
}

```

# Define motifs and run analyses + save results
Create summary table of analyses for selected motifs 
```{r}

motifs <- c("Ohler1", "DRE", "TATAbox", "INR", "Ebox", "Ohler6", "Ohler7",
            "DPE", "MTE", "TCT", "BREu", "BREd", "DCE1", "DCE2", "DCE3")

#motifs_10 <- c("DPE", "INR", "Ebox", "Ohler6", "DRE", "MTE", "Ohler7", "Ohler1", "TATAbox", "TCT")

motifs_10 <- c("DPE", "INR", "Ebox", "Ohler6", "Ohler1", "TATAbox")
print(motifs_10)

outcomes <- c("bsize", "bfreq", "exp")

results_both <- dplyr::bind_rows(
  fit_motif_models_list_both(param_list, outcomes, motifs_10, include_exp = FALSE),
  fit_motif_models_list_both(param_list, outcomes, motifs_10, include_exp = TRUE)
)

head(results_both)

#save 
# write.table(
#   results_both,
#   "/groups/stark/lorbeer/Burst_params_DG/STAP_March2025_pools/Fig2_promoter_motifs/linear_regression_results_STAP_KS_mu_20260330.txt",
#   quote = FALSE,
#   row.names = FALSE,
#   sep = "\t"
# )

```

# heatmap plotting functions
This section defines the plotting function used to visualize motif-associated effect estimates across samples. The heatmaps display regression coefficients for each motif and sample combination, separated by outcome. Tile color indicates the direction and magnitude of the estimated effect, while significance stars show the corresponding raw p-value category. The function also allows flexible ordering of motifs and samples and can suppress tiles for motifs that occur too infrequently in a given sample.

```{r}
sym_limits <- function(x, midpoint = 0) {
  x <- x[is.finite(x)]
  if (length(x) == 0) return(c(-1, 1))
  max_abs <- max(abs(x - midpoint))
  c(midpoint - max_abs, midpoint + max_abs)
}

make_heatmap_model <- function(results, model_name,
                               motifs_order = NULL,
                               samples_order = NULL,
                               min_occurrence = 0,
                               star_size = 3,
                               analysis = NULL) {
  
  plot_df <- results %>%
    dplyr::filter(Model == model_name, Outcome %in% c("bsize", "bfreq", "exp"))
  
  analysis_label <- if (is.null(analysis)) "both (univariate + multivariable)" else analysis
  
  if (!is.null(analysis)) {
    plot_df <- plot_df %>% dplyr::filter(Analysis == analysis)
  }
  
  if (!is.null(samples_order)) {
    plot_df <- plot_df %>%
      dplyr::filter(sample_name %in% samples_order) %>%
      dplyr::mutate(sample_name = factor(sample_name, levels = samples_order))
  } else {
    plot_df <- plot_df %>% dplyr::mutate(sample_name = factor(sample_name))
  }
  
  plot_df <- plot_df %>%
    dplyr::mutate(
      value = estimate,
      value = ifelse(!is.na(n_present) & n_present < min_occurrence, NA_real_, value),
      Sig_raw_plot = ifelse(sig %in% c("*", "**", "***"), sig, "")
    )
  
  if (!is.null(motifs_order)) {
    plot_df <- plot_df %>% dplyr::mutate(Motif = factor(Motif, levels = motifs_order))
  } else {
    plot_df <- plot_df %>% dplyr::mutate(Motif = factor(Motif))
  }
  
  lims <- sym_limits(plot_df$value, midpoint = 0)
  
  ggplot2::ggplot(plot_df, ggplot2::aes(x = sample_name, y = Motif)) +
    ggplot2::geom_tile(ggplot2::aes(fill = value)) +
    ggplot2::geom_text(ggplot2::aes(label = Sig_raw_plot), size = star_size, na.rm = TRUE) +
    ggplot2::facet_grid(. ~ Outcome) +
    ggplot2::scale_fill_gradient2(
      low = "blue", mid = "white", high = "red",
      midpoint = 0,
      limits = lims,
      na.value = "grey80"
    ) +
    ggplot2::labs(
      title = paste0("Motif effects (", analysis_label, "): ", model_name),
      subtitle = paste0(
        "Stars = raw p-value; tiles shown only if n_present ≥ ", min_occurrence,
        if (!is.null(samples_order)) paste0("; samples ordered (n=", length(samples_order), ")") else ""
      ),
      x = "Sample",
      y = "Motif",
      fill = "estimate"
    ) +
    ggplot2::theme_minimal(base_size = 11) +
    ggplot2::theme(
      panel.grid = ggplot2::element_blank(),
      axis.text.x = ggplot2::element_text(angle = 45, hjust = 1, vjust = 1),
      strip.text = ggplot2::element_text(face = "bold")
    )
}


```

# Heatmaps for pooled results

```{r}
min_occurrence <- 5
motifs_order <- motifs_10

pooled_samples <- c(
  "libFL008_E28_RT2_pooledby2",
  "libFL016_E38_e325_pooledby2",
  "libFL017_E38_e339_pooledby2"
)

all_samples <- sample_names
samples_order <- pooled_samples

q_multiv <- make_heatmap_model(
  results_both, "motifs_only",
  motifs_order = motifs_order,
  samples_order = samples_order,
  min_occurrence = min_occurrence,
  analysis = "multivariate"
)

q_univ <- make_heatmap_model(
  results_both, "motifs_only",
  motifs_order = motifs_order,
  samples_order = samples_order,
  min_occurrence = min_occurrence,
  analysis = "univariate"
)

q_multiv
q_univ

# Optional manual PDF export:
# pdf(paste0(out_dir,"heatmap_multivariate_pooled.pdf"), width = 10, height = 5)
# print(q_multiv)
# dev.off()
# 
# pdf(paste0(out_dir,"heatmap_univariate_pooled.pdf"), width = 10, height = 5)
# print(q_univ)
# dev.off()


```

# Heatmaps for all samples: supplementary informaiton

```{r}
p_multiv <- make_heatmap_model(
  results_both, "motifs_only",
  motifs_order = motifs_order,
  samples_order = all_samples,
  min_occurrence = min_occurrence,
  analysis = "multivariate"
)

p_univ <- make_heatmap_model(
  results_both, "motifs_only",
  motifs_order = motifs_order,
  samples_order = all_samples,
  min_occurrence = min_occurrence,
  analysis = "univariate"
)

p_multiv
p_univ

# Optional manual PDF export:
# pdf(paste0(out_dir,"heatmap_multivariate_all_samples.pdf"), width = 14, height = 5)
# print(p_multiv)
# dev.off()
# 
# pdf(paste0(out_dir, "heatmap_univariate_all_samples.pdf"), width = 14, height = 5)
# print(p_univ)
# dev.off()


```

# Median difference and log2 fold-change analysis
```{r}
analyze_param_list_log2FC <- function(param_list,
                                      motifs_order,
                                      samples_order = NULL) {
  stopifnot(is.list(param_list))
  
  if (is.null(names(param_list)) || any(names(param_list) == "")) {
    names(param_list) <- paste0("sample_", seq_along(param_list))
  }
  
  if (!is.null(samples_order)) {
    keep <- intersect(samples_order, names(param_list))
    param_list <- param_list[keep]
    param_list <- param_list[samples_order[samples_order %in% names(param_list)]]
  }
  
  add_significance <- function(p) {
    dplyr::case_when(
      is.na(p)  ~ "",
      p < 0.001 ~ "***",
      p < 0.01  ~ "**",
      p < 0.05  ~ "*",
      TRUE      ~ ""
    )
  }
  
  sym_limits <- function(x, midpoint = 0) {
    x <- x[is.finite(x)]
    if (length(x) == 0) return(c(-1, 1))
    max_abs <- max(abs(x - midpoint), na.rm = TRUE)
    c(midpoint - max_abs, midpoint + max_abs)
  }
  
  results_df <- purrr::imap_dfr(param_list, function(df, sample_name) {
    df <- data.table::as.data.table(df)
    
    purrr::map_dfr(motifs_order, function(m) {
      if (!m %in% names(df)) {
        return(tibble::tibble(
          sample_name = sample_name,
          motif = m,
          measurement = c("bfreq", "bsize", "exp"),
          value_0 = NA_real_, value_1 = NA_real_, diff = NA_real_, log2FC = NA_real_,
          pvalue = NA_real_, Nobservations_0 = NA_integer_, Nobservations_1 = NA_integer_
        ))
      }
      
      dt_m <- df[, .(bfreq, bsize, exp, motif = get(m))]
      
      n0 <- sum(dt_m$motif == 0, na.rm = TRUE)
      n1 <- sum(dt_m$motif > 0, na.rm = TRUE)
      
      one_measure <- function(y) {
        med0 <- dt_m[motif == 0, median(get(y), na.rm = TRUE)]
        med1 <- dt_m[motif > 0, median(get(y), na.rm = TRUE)]
        wt <- suppressWarnings(stats::wilcox.test(get(y) ~ (motif > 0), data = dt_m))
        
        tibble::tibble(
          sample_name = sample_name,
          motif = m,
          measurement = y,
          value_0 = med0,
          value_1 = med1,
          diff = med1 - med0,
          log2FC = log2(med1 / med0),
          pvalue = wt$p.value,
          Nobservations_0 = n0,
          Nobservations_1 = n1
        )
      }
      
      dplyr::bind_rows(one_measure("bfreq"), one_measure("bsize"), one_measure("exp"))
    })
  }) %>%
    dplyr::mutate(
      significance = add_significance(pvalue),
      motif = factor(motif, levels = motifs_order),
      sample_name = factor(sample_name, levels = names(param_list)),
      measurement = dplyr::recode(
        measurement,
        bfreq = "Burst Frequency",
        bsize = "Burst Size",
        exp   = "Mean expression"
      )
    )
  
  lims <- sym_limits(results_df$log2FC, midpoint = 0)
  
  p <- ggplot2::ggplot(results_df, ggplot2::aes(x = sample_name, y = motif, fill = log2FC)) +
    ggplot2::geom_tile() +
    ggplot2::geom_text(ggplot2::aes(label = significance), size = 3) +
    ggplot2::facet_grid(. ~ measurement) +
    ggplot2::scale_fill_gradient2(
      limits = lims,
      low = "blue", mid = "white", high = "red",
      midpoint = 0,
      oob = scales::squish,
      na.value = "grey90",
      name = "log2FC"
    ) +
    ggplot2::theme_minimal() +
    ggplot2::theme(
      panel.grid = ggplot2::element_blank(),
      axis.text.x = ggplot2::element_text(angle = 45, hjust = 1),
      axis.title.x = ggplot2::element_blank(),
      axis.title.y = ggplot2::element_blank(),
      strip.text = ggplot2::element_text(face = "bold")
    )
  
  list(table = results_df, plot = p)
}


```

# Log2 fold-change plots
These heatmaps display the log2 fold-change in each parameter when comparing motif-present versus motif-absent observations.

```{r}

res_pooled <- analyze_param_list_log2FC(
  param_list,
  motifs_order = motifs_10,
  samples_order = pooled_samples
)
pooled_table <- res_pooled$table
head(pooled_table)
res_pooled$plot

res_all <- analyze_param_list_log2FC(param_list, motifs_order = motifs_10)
all_table <- res_all$table

head(all_table)

res_all$plot

# Optional manual PDF export:
pdf(paste0(out_dir,"log2FC_heatmap_pooled.pdf"), width = 12, height = 5)
print(res_pooled$plot)
dev.off()

pdf(paste0(out_dir,"log2FC_heatmap_all_samples.pdf"), width = 12, height = 5)
print(res_all$plot)
dev.off()

```

# Meta-analysis combining datasets

Meta-analysis is performed on the regression output stored in "results_both".
Input is restricted to:
        samples in pooled_samples[1:3]
        Analysis == "multivariate"
        Model == "motifs_only"
Therefore, the pooled analysis uses:
        multivariate motif models
        no additional adjustment for log10(exp)
        per-sample regression coefficients estimated from

Meta-analysis here is fixed-effects meta-analysis of the per-sample regression coefficients on the log10 scale.
The pooled output includes:
- pooled_estimate: pooled regression coefficient
- pooled_SE: standard error of the pooled estimate
- pooled_zValue: test statistic for whether the pooled effect differs from 0
- pooled_pValue: p-value for the pooled effect

The meta-analysis combines the regression coefficients and standard errors obtained from the sample-specific multivariate linear models for each motif–outcome combination. Pooled estimates were calculated under a fixed-effects, inverse-variance weighting framework, such that sample-level estimates with smaller standard errors contributed more strongly to the combined effect. In the present analysis, the pooled effects represent motif-associated changes on the log10(bsize) and log10(bfreq) scales. The resulting pooled point estimates and confidence intervals summarize the overall direction and magnitude of association across samples; confidence intervals that crossed zero were interpreted as not supporting a clear non-zero pooled effect.

```{r}

min_k <- 3

meta_input <- results_both %>%
  filter(sample_name %in% pooled_samples[1:3] & Analysis == "multivariate" & Model == "motifs_only") %>% 
  filter(
    is.finite(estimate),
    is.finite(std.error),
    std.error > 0,
    !is.na(sample_name)
  )

combos <- meta_input %>%
  distinct(sample_name, Motif, Outcome) %>%
  dplyr::count(Motif, Outcome, name = "k") %>%
  filter(k >= min_k)

pooled_list <- vector("list", nrow(combos))

for (i in seq_len(nrow(combos))) {
  subset_data <- meta_input %>%
    filter(Motif == combos$Motif[i], Outcome == combos$Outcome[i]) %>%
    distinct(sample_name, .keep_all = TRUE)
  
  meta_obj <- rma(
    yi = subset_data$estimate,
    sei = subset_data$std.error,
    method = "FE"
  )
  
  pooled_list[[i]] <- data.frame(
    Motif = combos$Motif[i],
    Outcome = combos$Outcome[i],
    k = nrow(subset_data),
    pooled_estimate = as.numeric(meta_obj$beta[1]),
    pooled_SE = as.numeric(meta_obj$se),
    pooled_pValue = as.numeric(meta_obj$pval),
    pooled_zValue = as.numeric(meta_obj$zval),
    pooled_FoldChange = 10^as.numeric(meta_obj$beta[1]),
    pooled_FC_low  = 10^as.numeric(meta_obj$ci.lb),
    pooled_FC_high = 10^as.numeric(meta_obj$ci.ub),
    tau2 = as.numeric(meta_obj$tau2),
    I2 = as.numeric(meta_obj$I2)
  )
}

pooled_results <- bind_rows(pooled_list) %>%
  mutate(
    Sig_raw = case_when(
      pooled_pValue < 0.001 ~ "***",
      pooled_pValue < 0.01  ~ "**",
      pooled_pValue < 0.05  ~ "*",
      TRUE ~ "ns"
    )
  )

head(pooled_results)

```


# Reshape pooled results for plotting
The pooled meta-analysis table is reshaped to compare pooled effects on bsize and bfreq for the same motif in a single row.
Only the outcomes
- bsize
- bfreq
are retained for this comparison plot.

The following pooled quantities are selected before reshaping:
- pooled_estimate
- pooled_SE
- pooled_pValue
- pooled_FoldChange
- pooled_FC_low
- pooled_FC_high
- k
- tau2 (estimated between-study variance)
- I2 (proportion of total variability)

The pooled meta-analysis results are reshaped into a wide format to allow direct comparison between burst size and burst frequency effects for the same motif. Confidence interval bounds are calculated from the pooled estimates and standard errors, and each motif is classified according to whether it shows significant pooled association with burst size, burst frequency, both, or neither. This processed table forms the input for the final summary scatter plot.

```{r}
df_wide <- pooled_results %>%
  filter(Outcome %in% c("bsize", "bfreq")) %>%
  select(Motif, Outcome,
         pooled_estimate, pooled_SE, pooled_pValue,
         pooled_FoldChange, pooled_FC_low, pooled_FC_high,
         k, tau2, I2) %>%
  pivot_wider(
    names_from = Outcome,
    values_from = c(pooled_estimate, pooled_SE, pooled_pValue,
                    pooled_FoldChange, pooled_FC_low, pooled_FC_high,
                    k, tau2, I2),
    names_sep = "_"
  ) %>%
  filter(!is.na(pooled_FoldChange_bsize), !is.na(pooled_FoldChange_bfreq))

df_wide <- df_wide %>%
  mutate(
    est_low_bsize  = pooled_estimate_bsize - 1.96 * pooled_SE_bsize,
    est_high_bsize = pooled_estimate_bsize + 1.96 * pooled_SE_bsize,
    est_low_bfreq  = pooled_estimate_bfreq - 1.96 * pooled_SE_bfreq,
    est_high_bfreq = pooled_estimate_bfreq + 1.96 * pooled_SE_bfreq
  )

df_wide <- df_wide %>%
  mutate(
    Significance = case_when(
      pooled_pValue_bsize < 0.05 & pooled_pValue_bfreq < 0.05 ~ "Both",
      pooled_pValue_bsize < 0.05 & pooled_pValue_bfreq >= 0.05 ~ "bsize only",
      pooled_pValue_bsize >= 0.05 & pooled_pValue_bfreq < 0.05 ~ "bfreq only",
      TRUE ~ "Neither"
    )
  )

head(df_wide)
```


# Meta-analysis plot of multivariate analysis 
This plot visualizes pooled meta-analysis results jointly for:
- burst size (bsize)
- burst frequency (bfreq)

Input data are taken from the reshaped table df_wide. Each point corresponds to one motif with pooled estimates available for both outcomes. Axes represent pooled regression coefficients on the log10 scale:
- x-axis: pooled_estimate_bsize
- y-axis: pooled_estimate_bfreq

Therefore, point position summarizes the pooled conditional motif effect on:
- log10(bsize) in the horizontal direction
- log10(bfreq) in the vertical direction

Horizontal error bars represent the approximate 95% CI for the pooled burst-size effect:
- from est_low_bsize to est_high_bsize
Vertical error bars represent the approximate 95% CI for the pooled burst-frequency effect:
- from est_low_bfreq to est_high_bfreq

```{r}
p_meta <- ggplot(
  df_wide,
  aes(
    x = pooled_estimate_bsize,
    y = pooled_estimate_bfreq,
    color = Significance,
    label = Motif
  )
) +
  geom_errorbarh(
    aes(xmin = est_low_bsize, xmax = est_high_bsize),
    height = 0, alpha = 0.4
  ) +
  geom_errorbar(
    aes(ymin = est_low_bfreq, ymax = est_high_bfreq),
    width = 0, alpha = 0.4
  ) +
  geom_point(size = 3) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray40") +
  geom_hline(yintercept = 0, linetype = "dotted") +
  geom_vline(xintercept = 0, linetype = "dotted") +
  theme_classic(base_size = 14) +
  scale_color_manual(values = c(
    "Both" = "deepskyblue",
    "bsize only" = "cyan3",
    "bfreq only" = "royalblue3",
    "Neither" = "gray60"
  )) +
  labs(
    x = expression("Pooled effect on log"[10]*"(bsize)  (beta)"),
    y = expression("Pooled effect on log"[10]*"(bfreq)  (beta)"),
    title = "Meta-analysis: Effect of motif presence on burst parameters",
    subtitle = paste0("Points = pooled betas; error bars = 95% CI; min_k = ", min_k)
  )

if (requireNamespace("ggrepel", quietly = TRUE)) {
  p_meta <- p_meta + ggrepel::geom_text_repel(size = 3, max.overlaps = 30)
} else {
  p_meta <- p_meta + geom_text(size = 3, vjust = -0.5, check_overlap = TRUE)
}

p_meta

# Optional manual PDF export:
# pdf(paste0(out_dir,"meta_pooled_analysis_estimates_KS_mu.pdf"), width = 6, height = 5)
# print(p_meta)
# dev.off()


```

```{r, echo=FALSE}

TF_motifs <- motifs_10
names(TF_motifs) <- motifs_10

new_STAP <- fread("/groups/stark/lorbeer/STAP-TWIST/Franzi_STAP_TWIST_metadata-v2.txt")

new_STAP$main_mots <- ""
for (x in names(TF_motifs)){
  column <- new_STAP[[x]]  # Use [[ ]] for column extraction
  new_STAP$main_mots <- paste0(new_STAP$main_mots, ifelse(column == 0, "", paste0(x, "-")))
}

new_STAP$V2 <- lapply(strsplit(new_STAP$main_mots, "-"), unlist(unlist))


```


```{r, echo = FALSE}
#count how many lines have which motif
#https://r-graph-gallery.com/128-ring-or-donut-plot.html
#https://rforpoliticalscience.com/2020/07/26/make-wes-anderson-themed-graphs-with-wesanderson-package-in-r/ 
TF_freq <- as.data.frame(table(unlist(lapply(strsplit(new_STAP$main_mots, "-"), unlist(unlist)))))

# 2. Set the desired order natively as a factor
desired_order <- rev(motifs_10)
TF_freq$Var1 <- factor(TF_freq$Var1, levels = desired_order)

# 3. Colors (with Ohler7!)
motif_cols <- c(
  "TATAbox" = "darkcyan",
  "DPE" = "goldenrod2",
  "Ebox" = "coral3", 
  "INR" = "orchid4", 
  "Ohler6" = "lightblue",
  "Ohler1" = "darkblue"
)

# 4. Plot using geom_col() instead of geom_rect()
motif_plot <- ggplot(TF_freq, aes(x = 3, y = Freq, fill = Var1)) +
  geom_col() +                     # geom_col automatically stacks based on factor order!
  coord_polar(theta = "y") +       # Wrap the stacked bar into a circle
  xlim(c(1.5, 3.5)) +              # The xlim controls how "thick" the donut ring is
  scale_fill_manual(values = motif_cols) +
  theme_void()

motif_plot

pdf(paste0(out_dir,"libFL008_promoter_motif_content.pdf"), width = 6, height = 5)
print(motif_plot)
dev.off()




print(TF_freq)
write.table(TF_freq, paste0(out_dir, "libFL008_promoter_motif_count.txt"), sep = "\t", quote = FALSE, col.names = FALSE)

#require(ggupset)
library(ComplexUpset)
library(dplyr)

# 1. Define the subset of your columns that are the 0/1 motif indicators
motif_cols <- motifs_10

# 2. Convert those specific columns to logical (TRUE/FALSE) 
# ComplexUpset works best when the sets are logicals!
new_STAP_upset <- new_STAP %>%
  mutate(across(all_of(motif_cols), ~ as.logical(.)))

# 3. Create the plot
up <- upset(
  new_STAP_upset, 
  intersect = motif_cols, 
  name = "Motif combinations",
  min_size = 2, # Only show bars with at least 2 observations
  width_ratio = 0.15
)

up

pdf(paste0(out_dir,"libFL008_promoter_motif_content_upset.pdf"), width = 6, height = 5)
print(up)
dev.off()

```




```{r, echo = FALSE}

scatter.plots <- list()

dat.sub <- param %>% filter(KS_pval > 0.01  & high_mu == TRUE & !is.na(TATAbox) & experiment == "E28" & rep == "pooledby2") %>% mutate(
    promoter_type = case_when(
      TATAbox == 1 & DPE == 1 ~ "both",
      TATAbox == 1 & DPE == 0 ~ "TATAbox",
      TATAbox == 0 & DPE == 1 ~ "DPE",
      TATAbox == 0 & DPE == 0 ~ "none"
    )
  ) %>% filter(promoter_type != "both")

dat.sub$promoter_type <- factor(dat.sub$promoter_type, levels = c("TATAbox", "DPE", "both", "none"))

# Define the individual comparisons you want to make
my_comparisons <- list(c("TATAbox", "DPE"), c("TATAbox", "none"), c("DPE", "none"))

# --- Burst Size Plot ---
p_val1 <- summary(aov(log10(bsize) ~ promoter_type, dat.sub))[[1]][["Pr(>F)"]][1]
scatter.plots[["burst size"]] <- dat.sub %>% 
  ggplot(aes(x = promoter_type, y = log10(bsize), color = promoter_type)) + 
  geom_boxplot(outlier.shape = NA) + 
  scale_color_manual(values = c("darkcyan", "goldenrod2", "gray32", "gray12")) + 
  facet_wrap(~library) + 
  ggtitle("burst size", subtitle = paste0("ANOVA p = ", format(p_val1, scientific = TRUE, digits = 3))) +
  theme_classic() + 
  theme(text = element_text(size = 15), legend.position = "none") +
  # This single line adds the individual comparison brackets and p-values
  stat_compare_means(comparisons = my_comparisons, method = "t.test", label = "p.signif") 
  # Use label = "p.format" if you want numbers instead of asterisks (***)

# --- Burst Frequency Plot ---
p_val2 <- summary(aov(log10(bfreq) ~ promoter_type, dat.sub))[[1]][["Pr(>F)"]][1]
scatter.plots[["burst frequency"]] <- dat.sub %>% 
  ggplot(aes(x = promoter_type, y = log10(bfreq), color = promoter_type)) + 
  geom_boxplot(outlier.shape = NA) + 
  scale_color_manual(values = c("darkcyan", "goldenrod2", "gray32", "gray12")) + 
  facet_wrap(~library) + 
  ggtitle("burst frequency", subtitle = paste0("ANOVA p = ", format(p_val2, scientific = TRUE, digits = 3))) +
  theme_classic() + 
  theme(text = element_text(size = 15), legend.position = "none") +
  stat_compare_means(comparisons = my_comparisons, method = "t.test", label = "p.signif")

# --- Exp Plot ---
p_val3 <- summary(aov(log10(exp) ~ promoter_type, dat.sub))[[1]][["Pr(>F)"]][1]
scatter.plots[["exp"]] <- dat.sub %>% 
  ggplot(aes(x = promoter_type, y = log10(exp), color = promoter_type)) + 
  geom_boxplot(outlier.shape = NA) + 
  scale_color_manual(values = c("darkcyan", "goldenrod2", "gray32", "gray12")) + 
  facet_wrap(~library) + 
  ggtitle("exp", subtitle = paste0("ANOVA p = ", format(p_val3, scientific = TRUE, digits = 3))) +
  theme_classic() + 
  theme(text = element_text(size = 15), legend.position = "none") +
  stat_compare_means(comparisons = my_comparisons, method = "t.test", label = "p.signif")

wrap_plots(scatter.plots)
#grid.arrange(grobs = scatter.plots, ncol = 3)

pdf(paste0(out_dir,"DPE_TATA_comparison_E28.pdf"), width = 7, height = 5)
print(wrap_plots(scatter.plots))
dev.off()

```
```{r, echo = FALSE}

#setwd("/groups/stark/vloubiere/projects/Franzi_bursting/")
devtools::load_all("/groups/stark/vloubiere/vlite/")
require(TFBSTools)
library(Biostrings)
# Import motifs
pwm <- readRDS("/groups/stark/vloubiere/motifs_db/CP_motifs_PWM.rds")
motif_names <- sapply(pwm$Pwms_perc, TFBSTools::name)

# 1. Extract the raw numeric matrices explicitly
raw_percentage_matrices <- lapply(pwm$Pwms_perc, function(x) {
  # This safely extracts the internal numeric matrix from the S4 object
  TFBSTools::Matrix(x) 
})

# 2. Convert percentages to "counts" (e.g., multiply by 100) and round to integers
# (motifmatchr and TFBSTools prefer integer counts for base matrices)
count_matrices <- lapply(raw_percentage_matrices, function(x) round(x * 100))

# 3. Create a clean PFMatrixList (Position Frequency Matrix)
pfm_list <- do.call(PFMatrixList, lapply(seq_along(count_matrices), function(i) {
  # Try to get the original name, or make a new one
  orig_names <- names(pwm$Pwms_perc)
  m_name <- if (!is.null(orig_names)) orig_names[i] else paste0("Motif_", i)
  
  PFMatrix(name = m_name, profileMatrix = count_matrices[[i]])
}))
# 4. Convert to a proper PWMatrixList (Log-odds ratio)
# type = "log2probratio" is the standard log-odds format required by motifmatchr
fixed_pwms <- toPWM(pfm_list, type = "log2probratio", pseudocounts = 0.8)

# 1. Import sequences and store them into a DNAStringSet
seq <- fread("/groups/stark/lorbeer/STAP-TWIST/Franzi_STAP_STARR_TWIST_seqonly-v2.tsv")[grep("STAP", ID_TWIST)]
seqV <- substr(seq$Sequence_with_linker, 59, 59+132)
names(seqV) <- seq$ID_TWIST

dna_seqs <- DNAStringSet(seqV)

# 3. Convert to log odds
fixed_pwms <- toPWM(pfm_list, type = "log2probratio", pseudocounts = 0.8)

# --- 4. Unpack the GRangesList manually into plain data.tables ---
res_list <- lapply(seq_along(motif_hits), function(i) {
  hits <- motif_hits[[i]] 
  if (length(hits) > 0) {
    dt <- as.data.table(as.data.frame(hits))
    dt[, motif := motif_names[i]]
    return(dt)
  }
})

bed <- rbindlist(res_list, fill = TRUE)

# 3. Rename sequence ID column 
# If 'group' exists, it is the exact numeric index of the sequence.
bed[, seqlvls := names(dna_seqs)[group]]
bed[, seqlvls := gsub("_STAP-v2$", "", seqlvls)]

# 4. Format and Extract 
if ("strand" %in% names(bed)) bed <- bed[strand == "+"]

# Calculate the midpoint (creates the 'pos' column)
bed[, pos := (start + end) / 2]

# Clean up the sequence names string
bed[, seqlvls := gsub("_STAP-v2$", "", as.character(seqlvls))]

# Take max score per sequence/motif
bed <- bed[, .SD[which.max(score)], .(seqlvls, motif)]

# Expand out into your final matrices!
score_mat <- dcast(bed, seqlvls ~ motif, value.var = "score", fill = 0)
pos_mat   <- dcast(bed, seqlvls ~ motif, value.var = "pos")

head(score_mat)


# Sanit checks compare with Vanja  
# old <- fread("/groups/stark/lorbeer/STAP-TWIST/Franzi_STAP_TWIST_metadata-v2.txt")
# cmb <- merge(score_mat, old[, .(seqlvls= oligo_id, TATA.old= TATAbox, INR.old= INR)], by= "seqlvls", all.x= T, sort= F)

# pdf(paste0(out_dir,"define_CP_motif_center.pdf"), width = 4, height = 15)
# vl_par(mfrow= c(19, 1),
#        mai= c(.2, .9, .3, .9))
# for(motif in names(pos_mat)[-1]) {
#   # Compute density
#   dens <- density(pos_mat[[motif]][score_mat[[motif]]>0])
#   # Max position
#   c.pos <- round(dens$x[which.max(dens$y)])
#   ext <- ifelse(motif %in% c("TATAbox"), 5, 10)
#   # plot density
#   plot(dens, xlim= c(0, 133), main= paste0(motif, " n= ", sum(score_mat[[motif]]>0), "/", nrow(score_mat), " (", c.pos, "+/-", ext, ")"))
#   # Add lines around max
#   abline(v= c.pos+c(-ext, ext))
# }
# dev.off()

# Import Franzie's bursting params and add them to the score table
param_28 <- param[sample=="libFL008_E28_RT2_pooledby2" & (high_mu & KS_pval > 0.01)]
sub.score <- merge(score_mat, param_28[, .(seqlvls= seq_ID, bfreq, bsize, log2FoldChange)], by= "seqlvls", sort= F)
sub.pos <- pos_mat[sub.score$seqlvls, on= "seqlvls"]
stopifnot(identical(sub.score$seqlvls, sub.pos$seqlvls))

# 2. Open PDF
pdf(paste0(out_dir, "motif_score_vs_size_freq.pdf"), height = 3.5, width = 7.5)

for(motif in c("TATAbox", "DPE")) {
  
  # Check at position
  pos.ok <- if(motif=="TATAbox") {
    between(sub.pos[[motif]], 35, 45)
  } else if(motif=="DPE") {
    between(sub.pos[[motif]], 80, 100)
  } else {
    stop("Please define expected position for motif")
  }
  
  # Create a clean data.frame for ggplot
  dt_plot <- data.table(
    score = sub.score[[motif]],
    freq  = log10(sub.score$bfreq),
    size  = log10(sub.score$bsize),
    pos_ok = factor(pos.ok, levels = c("FALSE", "TRUE")) 
  )
  
  # --- FREQUENCY PLOT & STATS ---
  cor_freq <- cor.test(dt_plot$score, dt_plot$freq, method = "pearson")
  anno_freq <- sprintf("PCC = %.2f\np-val = %.2e", cor_freq$estimate, cor_freq$p.value)
  
  p_freq <- ggplot(dt_plot, aes(x = score, y = freq)) +
    geom_point(aes(color = pos_ok), alpha = 0.3, size = 1) +
    geom_smooth(method = "lm", color = "blue", se = FALSE, linewidth = 0.8) +
    scale_color_manual(values = c("TRUE" = "red", "FALSE" = "black")) +
    annotate("text", x = -Inf, y = -Inf, label = anno_freq, 
             vjust = -0.5, hjust = -0.1, fontface = "bold", size = 3.5) +
    labs(x = "Motif score", y = "Burst Frequency (log10)", title = motif) +
  theme_classic() + 
  theme(text = element_text(size = 15), legend.position = "none")
  
  # --- SIZE PLOT & STATS ---
  cor_size <- cor.test(dt_plot$score, dt_plot$size, method = "pearson")
  anno_size <- sprintf("PCC = %.2f\np-val = %.2e", cor_size$estimate, cor_size$p.value)
  
  p_size <- ggplot(dt_plot, aes(x = score, y = size)) +
    geom_point(aes(color = pos_ok), alpha = 0.3, size = 1) +
    geom_smooth(method = "lm", color = "blue", se = FALSE, linewidth = 0.8) +
    scale_color_manual(values = c("TRUE" = "red", "FALSE" = "black")) +
    annotate("text", x = -Inf, y = -Inf, label = anno_size, 
             vjust = -0.5, hjust = -0.1, fontface = "bold", size = 3.5) +
    labs(x = "Motif score", y = "Burst Size (log10)", title = motif) +
  theme_classic() + 
  theme(text = element_text(size = 15), legend.position = "none")
  
  # Print them side-by-side using grid.arrange
  grid.arrange(p_freq, p_size, ncol = 2)
}

dev.off()


pdf(paste0(out_dir,"motif_score_vs_size_freq_pos_ok.pdf"), height = 3.5, width = 7.5)
for(motif in c("TATAbox", "DPE")) {
  
  # Check at position
  pos.ok <- if(motif == "TATAbox") {
    between(sub.pos[[motif]], 35, 45)
  } else if(motif == "DPE") {
    between(sub.pos[[motif]], 80, 100)
  } else {
    stop("Please define expected position for motif")
  }
  
  # Create a clean data.frame and IMMEDIATELY filter for pos.ok == TRUE
  dt_plot <- data.table(
    score = sub.score[[motif]],
    freq  = log10(sub.score$bfreq),
    size  = log10(sub.score$bsize),
    pos_ok = pos.ok
  )[pos_ok == TRUE] # This strips away all non-positioned hits
  
  # Set the requested motif color
  pt_color <- if (motif == "TATAbox") "darkcyan" else "goldenrod2"
  
  # --- FREQUENCY PLOT & STATS ---
  cor_freq <- cor.test(dt_plot$score, dt_plot$freq, method = "pearson")
  anno_freq <- sprintf("PCC = %.2f\np-val = %.2e", cor_freq$estimate, cor_freq$p.value)
  
  p_freq <- ggplot(dt_plot, aes(x = score, y = freq)) +
    geom_point(color = pt_color, alpha = 0.5, size = 2) +
    geom_smooth(method = "lm", color = "black", se = FALSE, linewidth = 0.8) +
    annotate("text", x = -Inf, y = -Inf, label = anno_freq, 
             vjust = -0.5, hjust = -0.1, fontface = "bold", size = 3.5) +
    labs(x = "Motif score", y = "Burst frequency (log10)", title = motif) +
  theme_classic() + 
  theme(text = element_text(size = 15), legend.position = "none")
  
  # --- SIZE PLOT & STATS ---
  cor_size <- cor.test(dt_plot$score, dt_plot$size, method = "pearson")
  anno_size <- sprintf("PCC = %.2f\np-val = %.2e", cor_size$estimate, cor_size$p.value)
  
  p_size <- ggplot(dt_plot, aes(x = score, y = size)) +
    geom_point(color = pt_color, alpha = 0.5, size = 2) +
    geom_smooth(method = "lm", color = "black", se = FALSE, linewidth = 0.8) +
    annotate("text", x = -Inf, y = -Inf, label = anno_size, 
             vjust = -0.5, hjust = -0.1, fontface = "bold", size = 3.5) +
    labs(x = "Motif score", y = "Burst size (log10)", title = motif) +
  theme_classic() + 
  theme(text = element_text(size = 15), legend.position = "none")
  
  # Print them side-by-side
  grid.arrange(p_freq, p_size, ncol = 2)
}

dev.off()



```


```{r, echo = FALSE}
# Packages
library(dplyr)
library(ggplot2)
library(ggpubr)



# Prepare data (same as you had)
STAP_annotated <- dat.sub %>% filter(between(exp,0.1,5)) %>%
  mutate(
    expression_range = ifelse(between(exp, 1.5, 5), "high",
                       ifelse(between(exp, 0.3, 1.5), "medium", "low")),
    expression_range = factor(expression_range, levels = c("low", "medium", "high"))
  )

# Filter once for reuse
df_base <- STAP_annotated %>%
  mutate(TATAbox = as.factor(TATAbox))

# 1) bsize plot
p_bsize <- ggplot(df_base, aes(x = TATAbox, y = log10(bsize))) +
  #geom_jitter(alpha = 0.3, color = "darkgrey", width = 0.15) +
  geom_boxplot(aes(color = TATAbox), alpha = 0.5, outlier.shape = NA) +
  facet_grid(sample ~ expression_range) +
  scale_color_manual(values = c("0" = "grey32", "1" = "darkcyan")) +
  stat_compare_means(
    method = "wilcox.test",
    label = "p.format",
    comparisons = list(c("0", "1"))
  ) +
  theme_classic() + 
  theme(text = element_text(size = 15), legend.position = "none")

# 2) bfreq plot
p_bfreq <- ggplot(df_base, aes(x = TATAbox, y = log10(bfreq))) +
  #geom_jitter(alpha = 0.3, color = "darkgrey", width = 0.15) +
  geom_boxplot(aes(color = TATAbox), alpha = 0.5, outlier.shape = NA) +
  facet_grid(sample ~ expression_range) +
  scale_color_manual(values = c("0" = "grey32", "1" = "darkcyan")) +
  stat_compare_means(
    method = "wilcox.test",
    label = "p.format",
    comparisons = list(c("0", "1"))
  ) +
  theme_classic() + 
  theme(text = element_text(size = 15), legend.position = "none")

# 3) exp plot
p_exp <- ggplot(df_base, aes(x = TATAbox, y = log10(exp))) +
  #geom_jitter(alpha = 0.3, color = "darkgrey", width = 0.15) +
  geom_boxplot(aes(color = TATAbox), alpha = 0.5, outlier.shape = NA) +
  facet_grid(sample~expression_range) +
  scale_color_manual(values = c("0" = "grey32", "1" = "darkcyan")) +
  stat_compare_means(
    method = "wilcox.test",
    label = "p.format",
    comparisons = list(c("0", "1"))
  ) +
  theme_classic() + 
  theme(text = element_text(size = 15), legend.position = "none")

grid.arrange( p_bsize, p_bfreq, p_exp, ncol = 3)
#ggsave(paste0(out_dir,"TATA_boxplots_expressionbins_E28.pdf"), plot = grid.arrange( p_bsize, p_bfreq, p_exp, ncol = 3), width = 8, height = 3, units = "in")

```


```{r, echo = FALSE}

# DPE
# Filter once for reuse
df_base <- STAP_annotated %>%
  mutate(DPE = as.factor(DPE))

# 1) bsize plot
p_bsize <- ggplot(df_base, aes(x = DPE, y = log10(bsize))) +
  #geom_jitter(alpha = 0.3, color = "darkgrey", width = 0.15) +
  geom_boxplot(aes(color = DPE), alpha = 0.5, outlier.shape = NA) +
  facet_grid(sample ~ expression_range) +
  scale_color_manual(values = c("0" = "grey32", "1" = "goldenrod2")) +
  stat_compare_means(
    method = "wilcox.test",
    label = "p.format",
    comparisons = list(c("0", "1"))
  ) +
  theme_classic() + 
  theme(text = element_text(size = 15), legend.position = "none")

# 2) bfreq plot
p_bfreq <- ggplot(df_base, aes(x = DPE, y = log10(bfreq))) +
  #geom_jitter(alpha = 0.3, color = "darkgrey", width = 0.15) +
  geom_boxplot(aes(color = DPE), alpha = 0.5, outlier.shape = NA) +
  facet_grid(sample ~ expression_range) +
  scale_color_manual(values = c("0" = "grey32", "1" = "goldenrod2")) +
  stat_compare_means(
    method = "wilcox.test",
    label = "p.format",
    comparisons = list(c("0", "1"))
  ) +
  theme_classic() + 
  theme(text = element_text(size = 15), legend.position = "none")

# 3) exp plot
p_exp <- ggplot(df_base, aes(x = DPE, y = log10(exp))) +
  #geom_jitter(alpha = 0.3, color = "darkgrey", width = 0.15) +
  geom_boxplot(aes(color = DPE), alpha = 0.5, outlier.shape = NA) +
  facet_grid(sample~expression_range) +
  scale_color_manual(values = c("0" = "grey32", "1" = "goldenrod2")) +
  stat_compare_means(
    method = "wilcox.test",
    label = "p.format",
    comparisons = list(c("0", "1"))
  ) +
  theme_classic() + 
  theme(text = element_text(size = 15), legend.position = "none")


grid.arrange( p_bsize, p_bfreq, p_exp, ncol = 3)
#ggsave(paste0(out_dir,"DPE_boxplots_expressionbins_E28.pdf"), plot = grid.arrange( p_bsize, p_bfreq, p_exp, ncol = 3), width = 8, height = 3, units = "in")

```


````{r, echo = FALSE}

# INR
# Filter once for reuse
df_base <- STAP_annotated %>%
  mutate(INR = as.factor(INR))

# 1) bsize plot
p_bsize <- ggplot(df_base, aes(x = INR, y = log10(bsize))) +
  #geom_jitter(alpha = 0.3, color = "darkgrey", width = 0.15) +
  geom_boxplot(aes(color = INR), alpha = 0.5, outlier.shape = NA) +
  facet_grid(sample ~ expression_range) +
  scale_color_manual(values = c("0" = "grey32", "1" = "orchid4")) +
  stat_compare_means(
    method = "wilcox.test",
    label = "p.format",
    comparisons = list(c("0", "1"))
  ) +
  theme_classic() + 
  theme(text = element_text(size = 15), legend.position = "none")

# 2) bfreq plot
p_bfreq <- ggplot(df_base, aes(x = INR, y = log10(bfreq))) +
  #geom_jitter(alpha = 0.3, color = "darkgrey", width = 0.15) +
  geom_boxplot(aes(color = INR), alpha = 0.5, outlier.shape = NA) +
  facet_grid(sample ~ expression_range) +
  scale_color_manual(values = c("0" = "grey32", "1" = "orchid4")) +
  stat_compare_means(
    method = "wilcox.test",
    label = "p.format",
    comparisons = list(c("0", "1"))
  ) +
  theme_classic() + 
  theme(text = element_text(size = 15), legend.position = "none")

# 3) exp plot
p_exp <- ggplot(df_base, aes(x = INR, y = log10(exp))) +
  #geom_jitter(alpha = 0.3, color = "darkgrey", width = 0.15) +
  geom_boxplot(aes(color = INR), alpha = 0.5, outlier.shape = NA) +
  facet_grid(sample~expression_range) +
  scale_color_manual(values = c("0" = "grey32", "1" = "orchid4")) +
  stat_compare_means(
    method = "wilcox.test",
    label = "p.format",
    comparisons = list(c("0", "1"))
  ) +
  theme_classic() + 
  theme(text = element_text(size = 15), legend.position = "none")


grid.arrange( p_bsize, p_bfreq, p_exp, ncol = 3)
#ggsave(paste0(out_dir,"INR_boxplots_expressionbins_E28.pdf"), plot = grid.arrange( p_bsize, p_bfreq, p_exp, ncol = 3), width = 8, height = 3, units = "in")

```
````{r, echo = FALSE}

# INR
# Filter once for reuse
df_base <- STAP_annotated %>%
  mutate(Ebox = as.factor(Ebox))

# 1) bsize plot
p_bsize <- ggplot(df_base, aes(x = Ebox, y = log10(bsize))) +
  #geom_jitter(alpha = 0.3, color = "darkgrey", width = 0.15) +
  geom_boxplot(aes(color = Ebox), alpha = 0.5, outlier.shape = NA) +
  facet_grid(sample ~ expression_range) +
  scale_color_manual(values = c("0" = "grey32", "1" = "coral3")) +
  stat_compare_means(
    method = "wilcox.test",
    label = "p.format",
    comparisons = list(c("0", "1"))
  ) +
  theme_classic() + 
  theme(text = element_text(size = 15), legend.position = "none")

# 2) bfreq plot
p_bfreq <- ggplot(df_base, aes(x = Ebox, y = log10(bfreq))) +
  #geom_jitter(alpha = 0.3, color = "darkgrey", width = 0.15) +
  geom_boxplot(aes(color = Ebox), alpha = 0.5, outlier.shape = NA) +
  facet_grid(sample ~ expression_range) +
  scale_color_manual(values = c("0" = "grey32", "1" = "coral3")) +
  stat_compare_means(
    method = "wilcox.test",
    label = "p.format",
    comparisons = list(c("0", "1"))
  ) +
  theme_classic() + 
  theme(text = element_text(size = 15), legend.position = "none")

# 3) exp plot
p_exp <- ggplot(df_base, aes(x = Ebox, y = log10(exp))) +
  #geom_jitter(alpha = 0.3, color = "darkgrey", width = 0.15) +
  geom_boxplot(aes(color = Ebox), alpha = 0.5, outlier.shape = NA) +
  facet_grid(sample~expression_range) +
  scale_color_manual(values = c("0" = "grey32", "1" = "coral3")) +
  stat_compare_means(
    method = "wilcox.test",
    label = "p.format",
    comparisons = list(c("0", "1"))
  ) +
  theme_classic() + 
  theme(text = element_text(size = 15), legend.position = "none")


grid.arrange( p_bsize, p_bfreq, p_exp, ncol = 3)
#ggsave(paste0(out_dir,"Ebox_boxplots_expressionbins_E28.pdf"), plot = grid.arrange( p_bsize, p_bfreq, p_exp, ncol = 3), width = 8, height = 3, units = "in")

```
````{r, echo = FALSE}

# INR
# Filter once for reuse
df_base <- STAP_annotated %>%
  mutate(Ohler6 = as.factor(Ohler6))

# 1) bsize plot
p_bsize <- ggplot(df_base, aes(x = Ohler6, y = log10(bsize))) +
  #geom_jitter(alpha = 0.3, color = "darkgrey", width = 0.15) +
  geom_boxplot(aes(color = Ohler6), alpha = 0.5, outlier.shape = NA) +
  facet_grid(sample ~ expression_range) +
  scale_color_manual(values = c("0" = "grey32", "1" = "darkblue")) +
  stat_compare_means(
    method = "wilcox.test",
    label = "p.format",
    comparisons = list(c("0", "1"))
  ) +
  theme_classic() + 
  theme(text = element_text(size = 15), legend.position = "none")

# 2) bfreq plot
p_bfreq <- ggplot(df_base, aes(x = Ohler6, y = log10(bfreq))) +
  #geom_jitter(alpha = 0.3, color = "darkgrey", width = 0.15) +
  geom_boxplot(aes(color = Ohler6), alpha = 0.5, outlier.shape = NA) +
  facet_grid(sample ~ expression_range) +
  scale_color_manual(values = c("0" = "grey32", "1" = "darkblue")) +
  stat_compare_means(
    method = "wilcox.test",
    label = "p.format",
    comparisons = list(c("0", "1"))
  ) +
  theme_classic() + 
  theme(text = element_text(size = 15), legend.position = "none")

# 3) exp plot
p_exp <- ggplot(df_base, aes(x = Ohler6, y = log10(exp))) +
  #geom_jitter(alpha = 0.3, color = "darkgrey", width = 0.15) +
  geom_boxplot(aes(color = Ohler6), alpha = 0.5, outlier.shape = NA) +
  facet_grid(sample~expression_range) +
  scale_color_manual(values = c("0" = "grey32", "1" = "darkblue")) +
  stat_compare_means(
    method = "wilcox.test",
    label = "p.format",
    comparisons = list(c("0", "1"))
  ) +
  theme_classic() + 
  theme(text = element_text(size = 15), legend.position = "none")


grid.arrange( p_bsize, p_bfreq, p_exp, ncol = 3)
#ggsave(paste0(out_dir,"Ohler6_boxplots_expressionbins_E28.pdf"), plot = grid.arrange( p_bsize, p_bfreq, p_exp, ncol = 3), width = 8, height = 3, units = "in")

```
````{r, echo = FALSE}

# INR
# Filter once for reuse
df_base <- STAP_annotated %>%
  mutate(Ohler1 = as.factor(Ohler1))

# 1) bsize plot
p_bsize <- ggplot(df_base, aes(x = Ohler1, y = log10(bsize))) +
  #geom_jitter(alpha = 0.3, color = "darkgrey", width = 0.15) +
  geom_boxplot(aes(color = Ohler1), alpha = 0.5, outlier.shape = NA) +
  facet_grid(sample ~ expression_range) +
  scale_color_manual(values = c("0" = "grey32", "1" = "lightblue")) +
  stat_compare_means(
    method = "wilcox.test",
    label = "p.format",
    comparisons = list(c("0", "1"))
  ) +
  theme_classic() + 
  theme(text = element_text(size = 15), legend.position = "none")

# 2) bfreq plot
p_bfreq <- ggplot(df_base, aes(x = Ohler1, y = log10(bfreq))) +
  #geom_jitter(alpha = 0.3, color = "darkgrey", width = 0.15) +
  geom_boxplot(aes(color = Ohler1), alpha = 0.5, outlier.shape = NA) +
  facet_grid(sample ~ expression_range) +
  scale_color_manual(values = c("0" = "grey32", "1" = "lightblue")) +
  stat_compare_means(
    method = "wilcox.test",
    label = "p.format",
    comparisons = list(c("0", "1"))
  ) +
  theme_classic() + 
  theme(text = element_text(size = 15), legend.position = "none")

# 3) exp plot
p_exp <- ggplot(df_base, aes(x = Ohler1, y = log10(exp))) +
  #geom_jitter(alpha = 0.3, color = "darkgrey", width = 0.15) +
  geom_boxplot(aes(color = Ohler1), alpha = 0.5, outlier.shape = NA) +
  facet_grid(sample~expression_range) +
  scale_color_manual(values = c("0" = "grey32", "1" = "lightblue")) +
  stat_compare_means(
    method = "wilcox.test",
    label = "p.format",
    comparisons = list(c("0", "1"))
  ) +
  theme_classic() + 
  theme(text = element_text(size = 15), legend.position = "none")


grid.arrange( p_bsize, p_bfreq, p_exp, ncol = 3)
#ggsave(paste0(out_dir,"Ohler1_boxplots_expressionbins_E28.pdf"), plot = grid.arrange( p_bsize, p_bfreq, p_exp, ncol = 3), width = 8, height = 3, units = "in")

```






